---
title: "DApipe_HvGerm_Exp2TD"
author: "Guido Giordano"
date: "2025-09-08"
output: html_document
---
#This is the pipeline for data analysis for barley germination experiment 2TD

#Load packages
```{r}
BiocManager::install("cleave")
BiocManager::install("Mfuzz")
BiocManager::install("Biobase")
BiocManager::install("org.At.tair.db")
BiocManager::install("clusterProfiler")

library(AnnotationDbi)
library(org.At.tair.db)
library(clusterProfiler)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(ggbreak)
library(cowplot)
library(stringr)
library(kableExtra)
library(clusterProfiler)
library(mice)
library(reshape2)
library(data.table)
library(cleaver)
library(Mfuzz)
library(Biobase)
library(textshape)
library(tidyr)
library(gridExtra)
library(grid)
library(pcaMethods)
library(impute)
library(pheatmap)
library(ggh4x)

```

#Calculate the total number of MS3 spectra acquired per batch

```{r}

# Single summary file path
summary_file <- "../../../R_import/HV3TS/summary/summary.txt"

# Function to read, clean, and plot the summary file
plot_ms3_per_fraction <- function(filepath, batch_label = "HV3TS") {
  data <- read.table(filepath, header = TRUE, sep = "\t")
  
  data_clean <- data %>%
    filter(!is.na(Fraction)) %>%
    mutate(Fraction = factor(Fraction, levels = 1:48))
  
  p <- ggplot(data_clean, aes(x = Fraction, y = MS3)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_minimal() +
    labs(
      title = paste("MS3 scans", batch_label),
      x = NULL,
      y = "MS3 Identifications"
    ) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, size = 6),
      plot.title = element_text(size = 12, hjust = 0.5)
    )
  
  return(p)
}

# Generate the plot
p <- plot_ms3_per_fraction(summary_file, "HV3TS")

# Show the plot
print(p)

# Save the plot
ggsave("../../../R_export/HV3TS/ms3_ID_Hv6.png",
       p, width = 7, height = 5, dpi = 300, scale = 0.8)


```

#Check data distribution and structure 
Here using ibaq protein groups after conversion with ibaq pipeline 
```{r}

# --- Load file ---
data_file <- "../../../R_import/HV3TS/pG_ibaq/pG_BAR_b1.csv"
pG_HV3TS <- data.table::fread(data_file)

# --- Remove Contaminants, Reverse, or OIBS ---
if (all(c("Reverse", "Potential.contaminant", "Only.identified.by.site") %in% colnames(pG_HV3TS))) {
  pG_HV3TS <- pG_HV3TS[
    Reverse != "+" &
    Potential.contaminant != "+" &
    Only.identified.by.site != "+"
  ]
} else if ("Reverse" %in% colnames(pG_HV3TS)) {
  pG_HV3TS <- pG_HV3TS[Reverse != "+"]
}

# --- Log2 Transformation of iBAQ Columns ---
ibaq_cols <- grep("iBAQ", colnames(pG_HV3TS), value = TRUE)
pG_HV3TS[, (ibaq_cols) := lapply(.SD, function(x) {
  x <- as.numeric(as.character(x))
  x[is.na(x)] <- 0
  log2(x + 1)
}), .SDcols = ibaq_cols]

# --- Median Centering of iBAQ Columns ---
pG_HV3TS[, (ibaq_cols) := lapply(.SD, function(x) x - median(x, na.rm = TRUE)), .SDcols = ibaq_cols]

# --- Output directories ---
plot_dir <- "../../../R_export/HV3TS/data_distribution/"
heatmap_dir <- "../../../R_export/HV3TS/heatmaps/"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(heatmap_dir, showWarnings = FALSE, recursive = TRUE)

# --- iBAQ Histogram Plots ---
ibaq_cols_hist <- grep("_summed_iBAQ$", colnames(pG_HV3TS), value = TRUE)
if (length(ibaq_cols_hist) > 0) {
  plots <- lapply(seq_along(ibaq_cols_hist), function(i) {
    ggplot(pG_HV3TS, aes_string(x = ibaq_cols_hist[i])) +
      geom_histogram(binwidth = 0.7, fill = "#105CB6", color = "black") +
      labs(title = paste("Channel", i), x = "Reporter Intensity (iBAQ)", y = "Frequency") +
      theme_classic() +
      xlim(-25, 15) +
      theme(
        plot.title = element_text(size = 12, hjust = 0.5),
        axis.title = element_text(size = 10),
        axis.text = element_text(size = 10)
      )
  })

  title <- grid::textGrob("pG_HV3TS", gp = grid::gpar(fontsize = 16, fontface = "bold"))
  plot_grid <- gridExtra::arrangeGrob(grobs = plots, ncol = 3, top = title)

  ggplot2::ggsave(
    filename = file.path(plot_dir, "pG_HV3TS_iBAQ_distribution.png"),
    plot = plot_grid,
    width = 12,
    height = 9,
    scale = 0.9
  )
  message("Saved iBAQ distribution plot for HV3TS.")
} else {
  message("No _summed_iBAQ columns found for HV3TS.")
}

# --- Heatmap Generation ---
ibaq_cols_heatmap <- grep("_iBAQ$", colnames(pG_HV3TS), value = TRUE)
if (length(ibaq_cols_heatmap) >= 2) {
  mat_raw <- as.matrix(pG_HV3TS[, ..ibaq_cols_heatmap])
  rownames(mat_raw) <- pG_HV3TS$Protein.IDs_pG
  mat_raw[is.na(mat_raw)] <- -23
  mat_scaled <- t(scale(t(mat_raw)))
  mat_scaled <- mat_scaled[apply(mat_scaled, 1, function(x) all(!is.na(x))), ]

  # Save static PNG
  heatmap_file <- file.path(heatmap_dir, "pG_HV3TS_heatmap_scaled.png")
  pheatmap::pheatmap(
    mat_scaled,
    filename = heatmap_file,
    cluster_rows = TRUE,
    cluster_cols = FALSE,
    show_rownames = FALSE,
    labels_col = paste0("channel", 1:length(ibaq_cols_heatmap))
  )
  message("Saved heatmap for HV3TS (", length(ibaq_cols_heatmap), " iBAQ columns).")
} else {
  message("Skipping heatmap for HV3TS - only ", length(ibaq_cols_heatmap), " _iBAQ columns found (need ≥2).")
}

```

##Normalisation 
- In this script we take our protein groups with ibaq values and we filter, transform and normaliye them so that thez are readz to be used for quantiative comparisons  


```{r}
## ============================================================
## Batch 1 — HV3TS  (CLEAN PIPELINE: reverse removal → filtering 
## → median normalization → log2 transform)
## ============================================================

library(data.table)
library(ggplot2)
library(pheatmap)

# ------------------------------------------------------------
# STEP 1 — Load data and remove reverse hits
# ------------------------------------------------------------
file_path_b1 <- "../../../R_import/HV3TS/pG_ibaq/pG_BAR_b1.csv"
b1_BAR_raw <- fread(file_path_b1)

if ("Reverse_pG" %in% colnames(b1_BAR_raw)) {
  n_before <- nrow(b1_BAR_raw)
  b1_BAR_raw <- b1_BAR_raw[Reverse_pG != "+", ]
  n_after <- nrow(b1_BAR_raw)
  cat(sprintf("Removed %d reverse hits. Remaining proteins: %d\n",
              n_before - n_after, n_after))
}

# Identify iBAQ intensity columns
intensity_cols <- grep("_summed_iBAQ$", colnames(b1_BAR_raw), value = TRUE)
stopifnot(length(intensity_cols) > 0)

# ------------------------------------------------------------
# STEP 2 — Remove proteins with >4 missing values (<0.1 counts as missing)
# ------------------------------------------------------------
n_before <- nrow(b1_BAR_raw)
keep_rows <- apply(b1_BAR_raw[, ..intensity_cols], 1,
                   function(x) sum(is.na(x) | x < 0.1) <= 4)
b1_BAR_raw <- b1_BAR_raw[keep_rows]
n_after <- nrow(b1_BAR_raw)
cat(sprintf("Removed %d proteins with >4 missing/<0.1 values. Remaining: %d\n",
            n_before - n_after, n_after))

# ------------------------------------------------------------
# Helper functions for QC plots
# ------------------------------------------------------------
make_qc_plot <- function(df, title, intensity_cols) {
  sums    <- colSums(df[, ..intensity_cols], na.rm = TRUE)
  medians <- sapply(intensity_cols, function(cn) median(df[[cn]], na.rm = TRUE))

  stats_df <- data.table(
    Channel = factor(paste0("Ch", seq_along(intensity_cols)),
                     levels = paste0("Ch", seq_along(intensity_cols))),
    Total_Intensity = sums,
    Median_Intensity = medians
  )

  # Rescale median curve to fit bar scale
  scale_factor <- max(stats_df$Total_Intensity) / max(stats_df$Median_Intensity)

  ggplot(stats_df, aes(x = Channel)) +
    geom_bar(aes(y = Total_Intensity), stat = "identity", fill = "steelblue", alpha = 0.6) +
    geom_point(aes(y = Median_Intensity * scale_factor), color = "red", size = 2.5) +
    geom_line(aes(y = Median_Intensity * scale_factor), color = "red", linewidth = 0.7) +
    scale_y_continuous(
      name = "Total Intensity (sum)",
      sec.axis = sec_axis(~ . / scale_factor, name = "Median Intensity")
    ) +
    labs(title = title) +
    theme_classic()
}

make_heatmap <- function(df, title, intensity_cols, outfile = NULL, max_rows = 1000) {
  m <- as.matrix(df[, ..intensity_cols])
  mode(m) <- "numeric"
  m[!is.finite(m)] <- NA

  m_log <- log10(m + 1)

  if (nrow(m_log) > max_rows) {
    vars <- apply(m_log, 1, var, na.rm = TRUE)
    idx <- order(vars, decreasing = TRUE)[1:max_rows]
    m_log <- m_log[idx, , drop = FALSE]
  }

  colnames(m_log) <- paste0("Ch", seq_along(intensity_cols))

  pheatmap(
    m_log,
    filename = outfile,
    main = title,
    scale = "row",
    cluster_rows = TRUE,
    cluster_cols = FALSE,
    show_rownames = FALSE,
    border_color = NA,
    silent = !is.null(outfile)
  )
}

# ------------------------------------------------------------
# STEP 3 — QC BEFORE normalization
# ------------------------------------------------------------
export_dir <- "../../../R_export/HV3TS/pg_iBAQ_intra/b1"
if (!dir.exists(export_dir)) dir.create(export_dir, recursive = TRUE)

# Barplot
p_pre <- make_qc_plot(b1_BAR_raw, "QC — pre-normalization", intensity_cols)
ggsave(file.path(export_dir, "qc_preNorm.png"), p_pre, width = 10, height = 5, dpi = 150)

# Heatmap
make_heatmap(b1_BAR_raw, "Heatmap — pre-normalization", intensity_cols,
             outfile = file.path(export_dir, "heatmap_preNorm.png"))

# ------------------------------------------------------------
# STEP 4 — MEDIAN NORMALIZATION
# ------------------------------------------------------------
sample_medians <- sapply(intensity_cols,
                         function(cn) median(b1_BAR_raw[[cn]], na.rm = TRUE))

grand_median <- median(sample_medians, na.rm = TRUE)
cor_factors <- grand_median / sample_medians

cat("Median normalization factors:\n")
print(cor_factors)

# Apply correction
for (col in intensity_cols) {
  b1_BAR_raw[[col]] <- b1_BAR_raw[[col]] * cor_factors[col]
}

cat("Median normalization applied.\n")

# ------------------------------------------------------------
# STEP 5 — Log2 transformation
# ------------------------------------------------------------
for (col in intensity_cols) {
  b1_BAR_raw[[col]] <- log2(b1_BAR_raw[[col]] + 1)
}
cat("Applied log2(+1) transform.\n")

# ------------------------------------------------------------
# STEP 6 — QC AFTER normalization + Save final matrix
# ------------------------------------------------------------
p_post <- make_qc_plot(b1_BAR_raw, "QC — post-normalization", intensity_cols)
ggsave(file.path(export_dir, "qc_postNorm.png"), p_post, width = 10, height = 5, dpi = 150)

make_heatmap(b1_BAR_raw, "Heatmap — post-normalization", intensity_cols,
             outfile = file.path(export_dir, "heatmap_postNorm.png"))

out_file <- "../../../R_Data/HV3TS/pg_iBAQ_intra/pg_iBAQ_BAR_b1_mediannorm.csv"
fwrite(b1_BAR_raw, out_file)

cat(sprintf("Saved normalized + log2-transformed dataset to:\n  %s\n", out_file))


```

#In this chunk we map intensity values to our protein groups 

```{r}

## ============================================
## HV3TS — Attach functional annotation to BAR data (single batch, 6 time points)
## ============================================


# ==========================================================
# STEP 0 — Load the normalized + log2-transformed BAR file
# ==========================================================
cat("Loading b1_BAR_raw (normalized + log2-transformed)...\n")

b1_BAR_path <- "../../../R_Data/HV3TS/pg_iBAQ_intra/pg_iBAQ_BAR_b1_mediannorm.csv"
if (!file.exists(b1_BAR_path)) {
  stop("❌ Could not find b1_BAR_raw file at: ", b1_BAR_path)
}

b1_BAR_raw <- fread(b1_BAR_path)
cat("✅ Loaded b1_BAR_raw with", nrow(b1_BAR_raw), "proteins and", ncol(b1_BAR_raw), "columns.\n")


# ==========================================================
# STEP 1 — Load DB matches (barley vs barley UniProt)
# ==========================================================
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_up"

barley_hvup <- read.delim(file.path(path, "barke_upbarley.tsv"),
                          sep = "\t", header = FALSE, stringsAsFactors = FALSE)
barley_hvup <- barley_hvup[, 1:2]
colnames(barley_hvup) <- c("barleyup_entry", "barley_entry")

# Clean and deduplicate
barley_hvup$barleyup_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_hvup$barleyup_entry)
barley_hvup <- barley_hvup %>% distinct(barley_entry, .keep_all = TRUE)


# ==========================================================
# STEP 2 — Attach UniProt functional annotation (barley)
# ==========================================================
path <- "../../../R_import/Hv6/anno/masters_up"

up_Hv <- read.delim(file.path(path, "unip_Hv_1505.tsv"),
                    sep = "\t", header = TRUE, stringsAsFactors = FALSE)
up_Hv <- up_Hv[, c(3,4,5,7,8,9)]
colnames(up_Hv) <- c("barleyup_entryname", "barleyup_pname", "barleyup_gname",
                     "barleyup_GO_bp", "barleyup_GO_mf", "barleyup_GO_cc")

barley_anno <- merge(
  barley_hvup,
  up_Hv,
  by.x = "barleyup_entry",
  by.y = "barleyup_entryname",
  all.x = TRUE
)


# ==========================================================
# STEP 3 — Add Arabidopsis ortholog mapping and annotation
# ==========================================================
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_upAt"

barley_ATup <- read.delim(file.path(path, "araT_barke.tsv"),
                          sep = "\t", header = FALSE, stringsAsFactors = FALSE)
barley_ATup <- barley_ATup[, 1:2]
colnames(barley_ATup) <- c("barley_entry", "araT_entry")
barley_ATup$araT_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_ATup$araT_entry)
barley_ATup <- barley_ATup %>% distinct(barley_entry, .keep_all = TRUE)

# Merge with barley annotation
barley_anno_expat <- merge(
  barley_anno,
  barley_ATup,
  by = "barley_entry",
  all.x = TRUE
)

# Arabidopsis UniProt annotation
At_UP_annotated <- read_tsv("../../../R_import/Hv6/anno/masters_up/unip_At_1505.tsv")
At_UP_annotated <- At_UP_annotated[, c(2,3,4,5,6,7,8)]
At_UP_annotated <- At_UP_annotated %>% mutate(TAIR = sub(";.*", "", TAIR))

# Merge full annotation
barley_full_anno <- merge(
  barley_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)

barley_full_anno <- barley_full_anno[, c(2,3,4,5,6,7,8,1,14,9,10,11,12,13)]
colnames(barley_full_anno) <- c("barley_entry","barleyup_entry","barleyup_pname","barleyup_gname",
                                "barleyup_GO_bp","barleyup_GO_mf","barleyup_GO_cc","araT_entry",
                                "TAIR","atup_pname","atup_gname","at_GO_bp","at_GO_mf","at_GO_cc")


# ==========================================================
# STEP 4 — Merge annotations with intensity matrix
# ==========================================================
setDT(barley_full_anno)

HV3TS_final <- merge(
  b1_BAR_raw,
  barley_full_anno,
  by.x = "Protein.IDs_pG",
  by.y = "barley_entry",
  all.x = TRUE,
  sort  = FALSE
)

# Move annotation columns next to ID
anno_cols <- setdiff(names(barley_full_anno), "barley_entry")
setcolorder(HV3TS_final,
            c("Protein.IDs_pG", anno_cols,
              setdiff(names(HV3TS_final), c("Protein.IDs_pG", anno_cols))))


# ==========================================================
# STEP 5 — Save final annotated + intensity matrix
# ==========================================================
out_dir <- "../../../R_Data/HV3TS/final_annotation"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

out_path <- file.path(out_dir, "HV3TS_BAR_final.csv")
fwrite(HV3TS_final, out_path)

message("✅ Saved final annotated HV3TS BAR matrix to: ", out_path)

```

#Generate heatmaps for specific protein families 

```{r}

## ============================================
## HV3TS — Expression of enzyme families (tissue-specific)
## ============================================

library(data.table)
library(dplyr)
library(pheatmap)
library(RColorBrewer)

# -------------------------
# STEP 1 — Load data
# -------------------------
out_path <- "../../../R_Data/HV3TS/final_annotation/HV3TS_BAR_final.csv"
final_HV3TS_mat <- fread(out_path)
cat("✅ Loaded HV3TS dataset with", nrow(final_HV3TS_mat), "proteins and", ncol(final_HV3TS_mat), "columns.\n")

# -------------------------
# STEP 2 — Select intensity columns
# -------------------------
int_cols <- grep("corrected\\.\\d_ev_ratio_TMT_corrected_summed_iBAQ$", names(final_HV3TS_mat), value = TRUE)
if (length(int_cols) == 0) stop("❌ No valid intensity columns found.")
cat("Selected", length(int_cols), "intensity columns.\n")

# Rename for clarity (tissue order)
time_points <- c("Aleurone", "Coat", "Embryo", "Endosperm", "Scutellum", "Seed")
setnames(final_HV3TS_mat, old = int_cols, new = time_points)
int_cols <- time_points

# -------------------------
# STEP 3 — Define enzyme families
# -------------------------
all_families_ID <- list(
  # --- STARCH DEGRADATION ---
  `Alpha-amylase` = c(
    "HORVU.BARKE.PROJ.3HG00259920.1","HORVU.BARKE.PROJ.5HG00441970.1","HORVU.BARKE.PROJ.6HG00559240.1","HORVU.BARKE.PROJ.6HG00561370.1",
    "HORVU.BARKE.PROJ.7HG00653310.1","HORVU.BARKE.PROJ.7HG00653330.1"
  ),
  `Beta-amylase` = c(
    "HORVU.BARKE.PROJ.4HG00303690.1","HORVU.BARKE.PROJ.1HG00036740.1","HORVU.BARKE.PROJ.2HG00126310.1",
    "HORVU.BARKE.PROJ.4HG00376360.1","HORVU.BARKE.PROJ.4HG00380960.1"
  ),
  `Alpha-glucosidase` = c(
    "HORVU.BARKE.PROJ.1HG00052860.1","HORVU.BARKE.PROJ.7HG00595360.1","HORVU.BARKE.PROJ.2HG00192840.1",
    "HORVU.BARKE.PROJ.2HG00192880.1","HORVU.BARKE.PROJ.3HG00202550.1",
    "HORVU.BARKE.PROJ.4HG00360800.1","HORVU.BARKE.PROJ.7HG00666400.1"
  ),
  `Isoamylase` = c(
    "HORVU.BARKE.PROJ.1HG00057560.1","HORVU.BARKE.PROJ.2HG00149340.1","HORVU.BARKE.PROJ.2HG00154890.1",
    "HORVU.BARKE.PROJ.5HG00443610.1","HORVU.BARKE.PROJ.7HG00618070.1",
    "HORVU.BARKE.PROJ.7HG00645630.1","HORVU.BARKE.PROJ.7HG00681480.1","HORVU.BARKE.PROJ.7HG00823000.1"
  ),

  # --- CELL WALL DEGRADATION ---
  `Beta-glucanase` = c(
    "HORVU.BARKE.PROJ.7HG00638840.1","HORVU.BARKE.PROJ.5HG00427900.1","HORVU.BARKE.PROJ.5HG00453760.1",
    "HORVU.BARKE.PROJ.7HG00596870.1","HORVU.BARKE.PROJ.7HG00606250.1",
    "HORVU.BARKE.PROJ.7HG00805710.1","HORVU.BARKE.PROJ.5HG00434990.1","HORVU.BARKE.PROJ.6HG00508740.1",
    "HORVU.BARKE.PROJ.7HG00596880.1","HORVU.BARKE.PROJ.1HG00014980.1","HORVU.BARKE.PROJ.6HG00561940.1",
    "HORVU.BARKE.PROJ.1HG00073020.1","HORVU.BARKE.PROJ.7HG00653770.1","HORVU.BARKE.PROJ.1HG00018950.1",
    "HORVU.BARKE.PROJ.5HG00452800.1","HORVU.BARKE.PROJ.4HG00311880.1","HORVU.BARKE.PROJ.4HG00345900.1",
    "HORVU.BARKE.PROJ.2HG00162350.1","HORVU.BARKE.PROJ.2HG00192140.1","HORVU.BARKE.PROJ.3HG00226090.1",
    "HORVU.BARKE.PROJ.4HG00379800.1","HORVU.BARKE.PROJ.1HG00078380.1","HORVU.BARKE.PROJ.6HG00554880.1",
    "HORVU.BARKE.PROJ.7HG00607220.1","HORVU.BARKE.PROJ.2HG00154880.1","HORVU.BARKE.PROJ.4HG00315590.1",
    "HORVU.BARKE.PROJ.3HG00268790.1","HORVU.BARKE.PROJ.7HG00639580.1",
    "HORVU.BARKE.PROJ.1HG00046400.1","HORVU.BARKE.PROJ.1HG00038030.1"
  ),
  `Beta-glucosidase` = c(
    "HORVU.BARKE.PROJ.2HG00165000.1","HORVU.BARKE.PROJ.2HG00165020.1","HORVU.BARKE.PROJ.3HG00285690.1",
    "HORVU.BARKE.PROJ.3HG00270250.1","HORVU.BARKE.PROJ.5HG00449910.1","HORVU.BARKE.PROJ.1HG00054020.1",
    "HORVU.BARKE.PROJ.5HG00449920.1","HORVU.BARKE.PROJ.5HG00449830.1","HORVU.BARKE.PROJ.5HG00449840.1",
    "HORVU.BARKE.PROJ.2HG00187930.1","HORVU.BARKE.PROJ.2HG00159250.1","HORVU.BARKE.PROJ.2HG00159220.1",
    "HORVU.BARKE.PROJ.7HG00616950.1","HORVU.BARKE.PROJ.7HG00612580.1","HORVU.BARKE.PROJ.7HG00649790.1",
    "HORVU.BARKE.PROJ.5HG00446020.1","HORVU.BARKE.PROJ.4HG00361530.1","HORVU.BARKE.PROJ.7HG00663760.1",
    "HORVU.BARKE.PROJ.3HG00279420.1","HORVU.BARKE.PROJ.3HG00199770.1","HORVU.BARKE.PROJ.2HG00107620.1",
    "HORVU.BARKE.PROJ.4HG00308030.1","HORVU.BARKE.PROJ.4HG00308050.1","HORVU.BARKE.PROJ.7HG00653840.1",
    "HORVU.BARKE.PROJ.2HG00164980.1","HORVU.BARKE.PROJ.5HG00463840.1","HORVU.BARKE.PROJ.5HG00463920.1",
    "HORVU.BARKE.PROJ.7HG00812770.1","HORVU.BARKE.PROJ.1HG00027930.1","HORVU.BARKE.PROJ.1HG00694950.1",
    "HORVU.BARKE.PROJ.3HG00266590.1","HORVU.BARKE.PROJ.5HG00463890.1","HORVU.BARKE.PROJ.ptg000375lG00684260.1",
    "HORVU.BARKE.PROJ.5HG00463790.1","HORVU.BARKE.PROJ.3HG00231710.1","HORVU.BARKE.PROJ.5HG00463960.1",
    "HORVU.BARKE.PROJ.1HG00067000.1","HORVU.BARKE.PROJ.2HG00181540.1","HORVU.BARKE.PROJ.2HG00181530.1"
  ),
  `Arabinofuranosidase` = c(
    "HORVU.BARKE.PROJ.5HG00420480.1","HORVU.BARKE.PROJ.4HG00325990.1","HORVU.BARKE.PROJ.4HG00346830.1",
    "HORVU.BARKE.PROJ.2HG00101670.1","HORVU.BARKE.PROJ.4HG00326450.1"
  ),
  `Mannanase` = c(
    "HORVU.BARKE.PROJ.3HG00263090.1","HORVU.BARKE.PROJ.7HG00612160.1","HORVU.BARKE.PROJ.3HG00203070.1",
    "HORVU.BARKE.PROJ.5HG00771660.1","HORVU.BARKE.PROJ.6HG00561230.1","HORVU.BARKE.PROJ.5HG00479880.1",
    "HORVU.BARKE.PROJ.3HG00253120.1","HORVU.BARKE.PROJ.6HG00561220.1"
  ),
  `Alpha-galactosidase` = c(
    "HORVU.BARKE.PROJ.5HG00772750.1","HORVU.BARKE.PROJ.1HG00041240.1","HORVU.BARKE.PROJ.5HG00772740.1",
    "HORVU.BARKE.PROJ.1HG00041210.1","HORVU.BARKE.PROJ.2HG00102930.1","HORVU.BARKE.PROJ.6HG00537080.1",
    "HORVU.BARKE.PROJ.6HG00556460.1"
  ),
  `Polygalacturonase` = c(
    "HORVU.BARKE.PROJ.6HG00502310.1","HORVU.BARKE.PROJ.3HG00294100.1","HORVU.BARKE.PROJ.1HG00694870.1","HORVU.BARKE.PROJ.5HG00475050.1",
    "HORVU.BARKE.PROJ.7HG00599930.1","HORVU.BARKE.PROJ.5HG00475020.1","HORVU.BARKE.PROJ.6HG00562290.1","HORVU.BARKE.PROJ.5HG00445800.1",
    "HORVU.BARKE.PROJ.3HG00278800.1","HORVU.BARKE.PROJ.2HG00170720.1","HORVU.BARKE.PROJ.3HG00221410.1","HORVU.BARKE.PROJ.2HG00170710.1",
    "HORVU.BARKE.PROJ.7HG00622760.2","HORVU.BARKE.PROJ.5HG00476730.1","HORVU.BARKE.PROJ.7HG00656610.1","HORVU.BARKE.PROJ.2HG00180220.1",
    "HORVU.BARKE.PROJ.7HG00622760.1","HORVU.BARKE.PROJ.2HG00177430.1","HORVU.BARKE.PROJ.3HG00731530.1","HORVU.BARKE.PROJ.3HG00250260.1",
    "HORVU.BARKE.PROJ.7HG00656560.1","HORVU.BARKE.PROJ.7HG00641670.1","HORVU.BARKE.PROJ.1HG00072280.1","HORVU.BARKE.PROJ.1HG00072330.1",
    "HORVU.BARKE.PROJ.1HG00072290.1","HORVU.BARKE.PROJ.4HG00314970.1","HORVU.BARKE.PROJ.3HG00745990.1","HORVU.BARKE.PROJ.3HG00279930.1",
    "HORVU.BARKE.PROJ.7HG00640290.1","HORVU.BARKE.PROJ.5HG00404030.1","HORVU.BARKE.PROJ.3HG00238660.1","HORVU.BARKE.PROJ.3HG00204800.1",
    "HORVU.BARKE.PROJ.5HG00481490.1","HORVU.BARKE.PROJ.1HG00062490.1","HORVU.BARKE.PROJ.5HG00436830.1","HORVU.BARKE.PROJ.7HG00576880.1",
    "HORVU.BARKE.PROJ.6HG00519420.1","HORVU.BARKE.PROJ.7HG00656560.2","HORVU.BARKE.PROJ.7HG00656680.1","HORVU.BARKE.PROJ.7HG00656650.1",
    "HORVU.BARKE.PROJ.4HG00315000.1","HORVU.BARKE.PROJ.4HG00376400.1","HORVU.BARKE.PROJ.3HG00248590.1","HORVU.BARKE.PROJ.1HG00013170.1",
    "HORVU.BARKE.PROJ.4HG00360780.1","HORVU.BARKE.PROJ.1HG00013190.1","HORVU.BARKE.PROJ.2HG00173550.1"
  ),
  `Pectin esterase` = c(
    "HORVU.BARKE.PROJ.7HG00656950.1","HORVU.BARKE.PROJ.7HG00598930.1","HORVU.BARKE.PROJ.5HG00473670.1","HORVU.BARKE.PROJ.7HG00639410.1",
    "HORVU.BARKE.PROJ.3HG00268440.1","HORVU.BARKE.PROJ.2HG00188100.1","HORVU.BARKE.PROJ.4HG00349290.1",
    "HORVU.BARKE.PROJ.2HG00100460.1","HORVU.BARKE.PROJ.3HG00232130.1","HORVU.BARKE.PROJ.2HG00157350.1",
    "HORVU.BARKE.PROJ.7HG00593290.1","HORVU.BARKE.PROJ.5HG00391190.1","HORVU.BARKE.PROJ.2HG00103570.1",
    "HORVU.BARKE.PROJ.1HG00045130.1","HORVU.BARKE.PROJ.3HG00230000.1","HORVU.BARKE.PROJ.6HG00564170.1",
    "HORVU.BARKE.PROJ.1HG00085260.1","HORVU.BARKE.PROJ.1HG00058460.1","HORVU.BARKE.PROJ.1HG00054490.1",
    "HORVU.BARKE.PROJ.5HG00437860.1","HORVU.BARKE.PROJ.2HG00154320.1","HORVU.BARKE.PROJ.3HG00280980.1",
    "HORVU.BARKE.PROJ.1HG00069330.1","HORVU.BARKE.PROJ.2HG00102050.1","HORVU.BARKE.PROJ.2HG00706320.1",
    "HORVU.BARKE.PROJ.2HG00115940.1","HORVU.BARKE.PROJ.5HG00403060.1","HORVU.BARKE.PROJ.2HG00107880.1",
    "HORVU.BARKE.PROJ.2HG00105260.1","HORVU.BARKE.PROJ.2HG00105250.1","HORVU.BARKE.PROJ.2HG00107890.1",
    "HORVU.BARKE.PROJ.3HG00300590.1","HORVU.BARKE.PROJ.2HG00171960.1","HORVU.BARKE.PROJ.4HG00366450.1",
    "HORVU.BARKE.PROJ.4HG00328640.1","HORVU.BARKE.PROJ.4HG00328630.1"
  ),
  `Xylan endohydrolase` = c(
    "HORVU.BARKE.PROJ.3HG00205970.1","HORVU.BARKE.PROJ.4HG00305690.1","HORVU.BARKE.PROJ.5HG00471950.1",
    "HORVU.BARKE.PROJ.3HG00205980.1","HORVU.BARKE.PROJ.5HG00779790.1","HORVU.BARKE.PROJ.5HG00471990.1",
    "HORVU.BARKE.PROJ.2HG00105770.1"
  ),
  `Beta-xylosidase` = c("HORVU.BARKE.PROJ.2HG00188490.1"),
  `Chitinase` = c("HORVU.BARKE.PROJ.2HG00171400.1"),

  # --- LIPID DEGRADATION ---
  `Phospholipase A1` = c(
    "HORVU.BARKE.PROJ.2HG00164160.1","HORVU.BARKE.PROJ.1HG00083010.1","HORVU.BARKE.PROJ.1HG00060140.1","HORVU.BARKE.PROJ.3HG00251690.1",
    "HORVU.BARKE.PROJ.3HG00251820.1","HORVU.BARKE.PROJ.1HG00081650.1","HORVU.BARKE.PROJ.1HG00083090.1","HORVU.BARKE.PROJ.3HG00251700.1",
    "HORVU.BARKE.PROJ.1HG00082950.1","HORVU.BARKE.PROJ.1HG00053440.1","HORVU.BARKE.PROJ.3HG00287780.1","HORVU.BARKE.PROJ.2HG00108030.1",
    "HORVU.BARKE.PROJ.3HG00279020.1","HORVU.BARKE.PROJ.4HG00313040.1","HORVU.BARKE.PROJ.4HG00313000.1","HORVU.BARKE.PROJ.6HG00543590.1",
    "HORVU.BARKE.PROJ.1HG00051590.1","HORVU.BARKE.PROJ.7HG00635870.1","HORVU.BARKE.PROJ.3HG00295580.1","HORVU.BARKE.PROJ.1HG00058050.1",
    "HORVU.BARKE.PROJ.3HG00295590.1","HORVU.BARKE.PROJ.3HG00259290.1","HORVU.BARKE.PROJ.3HG00248560.1","HORVU.BARKE.PROJ.3HG00231880.1",
    "HORVU.BARKE.PROJ.5HG00389370.1"
  ),
  `Phospholipase A2` = c(  # renamed from Phospholipase A2_Hv
    "HORVU.BARKE.PROJ.7HG00665330.1","HORVU.BARKE.PROJ.4HG00355590.1","HORVU.BARKE.PROJ.4HG00308310.1",
    "HORVU.BARKE.PROJ.2HG00180400.1","HORVU.BARKE.PROJ.1HG00072800.1",
    "HORVU.BARKE.PROJ.1HG00006150.1","HORVU.BARKE.PROJ.6HG00784580.1","HORVU.BARKE.PROJ.2HG00107210.1"
  ),
  `Phospholipase D` = c(
    "HORVU.BARKE.PROJ.6HG00499590.1","HORVU.BARKE.PROJ.7HG00621120.1","HORVU.BARKE.PROJ.3HG00270400.1",
    "HORVU.BARKE.PROJ.2HG00147990.1","HORVU.BARKE.PROJ.3HG00270400.2","HORVU.BARKE.PROJ.4HG00315710.1","HORVU.BARKE.PROJ.4HG00315700.1",
    "HORVU.BARKE.PROJ.6HG00556090.1","HORVU.BARKE.PROJ.2HG00706940.1","HORVU.BARKE.PROJ.1HG00024430.1","HORVU.BARKE.PROJ.2HG00116340.1",
    "HORVU.BARKE.PROJ.2HG00093030.1","HORVU.BARKE.PROJ.3HG00265370.1","HORVU.BARKE.PROJ.6HG00788600.1","HORVU.BARKE.PROJ.1HG00024450.1",
    "HORVU.BARKE.PROJ.7HG00616200.1","HORVU.BARKE.PROJ.2HG00720350.1","HORVU.BARKE.PROJ.6HG00533900.1","HORVU.BARKE.PROJ.6HG00521550.1",
    "HORVU.BARKE.PROJ.1HG00074020.1","HORVU.BARKE.PROJ.7HG00651940.1","HORVU.BARKE.PROJ.3HG00221500.1","HORVU.BARKE.PROJ.1HG00075940.1",
    "HORVU.BARKE.PROJ.1HG00025160.1","HORVU.BARKE.PROJ.4HG00381620.1","HORVU.BARKE.PROJ.1HG00075980.1","HORVU.BARKE.PROJ.4HG00381810.1",
    "HORVU.BARKE.PROJ.5HG00439260.1","HORVU.BARKE.PROJ.1HG00045880.1","HORVU.BARKE.PROJ.4HG00377390.1","HORVU.BARKE.PROJ.1HG00072310.1",
    "HORVU.BARKE.PROJ.5HG00455460.1","HORVU.BARKE.PROJ.5HG00434890.1","HORVU.BARKE.PROJ.6HG00499580.1","HORVU.BARKE.PROJ.1HG00054620.1"
  ),
  `Phospholipase C` = c(
    "HORVU.BARKE.PROJ.1HG00014000.1","HORVU.BARKE.PROJ.2HG00099060.1","HORVU.BARKE.PROJ.1HG00014040.1",
    "HORVU.BARKE.PROJ.5HG00409430.1","HORVU.BARKE.PROJ.5HG00423610.1","HORVU.BARKE.PROJ.4HG00350900.1"
  ),
  `Lipase` = c(
    "HORVU.BARKE.PROJ.2HG00116590.1","HORVU.BARKE.PROJ.5HG00452660.1","HORVU.BARKE.PROJ.7HG00622910.1",
    "HORVU.BARKE.PROJ.5HG00450050.1","HORVU.BARKE.PROJ.5HG00424780.1","HORVU.BARKE.PROJ.7HG00654330.1",
    "HORVU.BARKE.PROJ.2HG00143600.1","HORVU.BARKE.PROJ.7HG00620220.1","HORVU.BARKE.PROJ.7HG00654320.1",
    "HORVU.BARKE.PROJ.3HG00279020.1","HORVU.BARKE.PROJ.4HG00313040.1","HORVU.BARKE.PROJ.3HG00248560.1",
    "HORVU.BARKE.PROJ.3HG00231880.1","HORVU.BARKE.PROJ.3HG00265210.1","HORVU.BARKE.PROJ.5HG00476290.1"
  ),
  `Lipoxygenase-1` = c(
    "HORVU.BARKE.PROJ.5HG00462380.1","HORVU.BARKE.PROJ.4HG00307600.1","HORVU.BARKE.PROJ.2HG00188750.1","HORVU.BARKE.PROJ.4HG00307640.1"
  ),
  `Lipoxygenase-2` = c(
    "HORVU.BARKE.PROJ.3HG00735580.1","HORVU.BARKE.PROJ.4HG00760420.1","HORVU.BARKE.PROJ.6HG00519710.1","HORVU.BARKE.PROJ.6HG00490920.1",
    "HORVU.BARKE.PROJ.5HG00383870.1","HORVU.BARKE.PROJ.7HG00579310.1","HORVU.BARKE.PROJ.7HG00616920.1"
  ),
  `Hydroperoxide lyase` = c("HORVU.BARKE.PROJ.6HG00493590.1"),
  `Allene oxide synthase` = c(  # renamed from Allene oxide synthase_Hv
    "HORVU.BARKE.PROJ.3HG00297970.1","HORVU.BARKE.PROJ.5HG00481590.1","HORVU.BARKE.PROJ.1HG00072850.1"
  ),

  # --- PROTEIN DEGRADATION ---
  `Subtilisin-like protease` = c(
    "HORVU.BARKE.PROJ.7HG00594380.1","HORVU.BARKE.PROJ.5HG00387520.1","HORVU.BARKE.PROJ.7HG00652880.1","HORVU.BARKE.PROJ.3HG00258310.1",
    "HORVU.BARKE.PROJ.6HG00518840.1","HORVU.BARKE.PROJ.2HG00092410.1","HORVU.BARKE.PROJ.2HG00177680.1","HORVU.BARKE.PROJ.3HG00236140.1",
    "HORVU.BARKE.PROJ.1HG00045900.1","HORVU.BARKE.PROJ.1HG00053580.1","HORVU.BARKE.PROJ.1HG00052650.1","HORVU.BARKE.PROJ.2HG00171370.1",
    "HORVU.BARKE.PROJ.2HG00091080.1","HORVU.BARKE.PROJ.4HG00371060.1","HORVU.BARKE.PROJ.6HG00545200.1","HORVU.BARKE.PROJ.2HG00185200.1",
    "HORVU.BARKE.PROJ.4HG00361750.1","HORVU.BARKE.PROJ.2HG00169600.1","HORVU.BARKE.PROJ.6HG00562150.1","HORVU.BARKE.PROJ.6HG00562050.1",
    "HORVU.BARKE.PROJ.6HG00562140.1","HORVU.BARKE.PROJ.5HG00768700.1","HORVU.BARKE.PROJ.6HG00504300.1","HORVU.BARKE.PROJ.6HG00561990.1",
    "HORVU.BARKE.PROJ.6HG00562070.1","HORVU.BARKE.PROJ.4HG00377350.1","HORVU.BARKE.PROJ.1HG00057060.1","HORVU.BARKE.PROJ.5HG00763810.1",
    "HORVU.BARKE.PROJ.1HG00075820.1","HORVU.BARKE.PROJ.3HG00282630.1","HORVU.BARKE.PROJ.2HG00171120.1","HORVU.BARKE.PROJ.1HG00065070.1",
    "HORVU.BARKE.PROJ.4HG00357630.1","HORVU.BARKE.PROJ.7HG00646960.1","HORVU.BARKE.PROJ.1HG00053700.1","HORVU.BARKE.PROJ.2HG00101890.1",
    "HORVU.BARKE.PROJ.6HG00493740.1","HORVU.BARKE.PROJ.7HG00669560.1","HORVU.BARKE.PROJ.5HG00444980.1","HORVU.BARKE.PROJ.6HG00489550.1",
    "HORVU.BARKE.PROJ.5HG00465760.1","HORVU.BARKE.PROJ.3HG00273400.1","HORVU.BARKE.PROJ.2HG00171140.1","HORVU.BARKE.PROJ.2HG00171130.1",
    "HORVU.BARKE.PROJ.2HG00153370.1","HORVU.BARKE.PROJ.2HG00185220.1","HORVU.BARKE.PROJ.2HG00185190.1","HORVU.BARKE.PROJ.4HG00313730.1",
    "HORVU.BARKE.PROJ.5HG00436650.1","HORVU.BARKE.PROJ.2HG00101990.1","HORVU.BARKE.PROJ.7HG00608940.1","HORVU.BARKE.PROJ.3HG00274210.1",
    "HORVU.BARKE.PROJ.1HG00068460.1","HORVU.BARKE.PROJ.3HG00266160.1","HORVU.BARKE.PROJ.2HG00168300.1","HORVU.BARKE.PROJ.7HG00659580.1",
    "HORVU.BARKE.PROJ.3HG00268750.1","HORVU.BARKE.PROJ.1HG00082000.1","HORVU.BARKE.PROJ.5HG00452960.1","HORVU.BARKE.PROJ.6HG00521310.1",
    "HORVU.BARKE.PROJ.3HG00199430.1","HORVU.BARKE.PROJ.6HG00571850.1","HORVU.BARKE.PROJ.7HG00592920.1","HORVU.BARKE.PROJ.7HG00592920.2",
    "HORVU.BARKE.PROJ.3HG00223220.1","HORVU.BARKE.PROJ.7HG00592910.1","HORVU.BARKE.PROJ.3HG00223260.1","HORVU.BARKE.PROJ.4HG00314900.1",
    "HORVU.BARKE.PROJ.6HG00506350.1","HORVU.BARKE.PROJ.7HG00592890.1","HORVU.BARKE.PROJ.7HG00592900.1","HORVU.BARKE.PROJ.7HG00592880.1",
    "HORVU.BARKE.PROJ.4HG00365940.1","HORVU.BARKE.PROJ.2HG00092420.1","HORVU.BARKE.PROJ.5HG00444660.1","HORVU.BARKE.PROJ.3HG00261180.1"
  ),
  `Papain-like cysteine proteinase` = c(
    "HORVU.BARKE.PROJ.6HG00784430.1","HORVU.BARKE.PROJ.6HG00495310.1","HORVU.BARKE.PROJ.6HG00495260.1","HORVU.BARKE.PROJ.6HG00498700.1",
    "HORVU.BARKE.PROJ.6HG00494350.1","HORVU.BARKE.PROJ.6HG00495680.1","HORVU.BARKE.PROJ.6HG00495640.1","HORVU.BARKE.PROJ.6HG00495690.1",
    "HORVU.BARKE.PROJ.6HG00495700.1","HORVU.BARKE.PROJ.6HG00495660.1","HORVU.BARKE.PROJ.6HG00495710.1","HORVU.BARKE.PROJ.6HG00495650.1",
    "HORVU.BARKE.PROJ.1HG00077400.1","HORVU.BARKE.PROJ.6HG00495670.1","HORVU.BARKE.PROJ.7HG00679490.1","HORVU.BARKE.PROJ.7HG00679480.1",
    "HORVU.BARKE.PROJ.2HG00194010.1","HORVU.BARKE.PROJ.4HG00382300.1","HORVU.BARKE.PROJ.5HG00459870.1","HORVU.BARKE.PROJ.2HG00097640.1",
    "HORVU.BARKE.PROJ.3HG00208070.1"
  ),
  `Aminopeptidase` = c(
    "HORVU.BARKE.PROJ.4HG00305500.1","HORVU.BARKE.PROJ.5HG00766870.1","HORVU.BARKE.PROJ.5HG00432120.1",
    "HORVU.BARKE.PROJ.6HG00524520.1","HORVU.BARKE.PROJ.7HG00589820.1"
  ),
  `Methionyl-aminopeptidase` = c(
    "HORVU.BARKE.PROJ.1HG00055420.1","HORVU.BARKE.PROJ.1HG00044760.1","HORVU.BARKE.PROJ.6HG00536050.1",
    "HORVU.BARKE.PROJ.2HG00129520.1","HORVU.BARKE.PROJ.2HG00174090.1",
    "HORVU.BARKE.PROJ.6HG00559610.1","HORVU.BARKE.PROJ.7HG00622240.1"
  ),
  `Carboxypeptidase` = c(
    "HORVU.BARKE.PROJ.6HG00499600.1","HORVU.BARKE.PROJ.4HG00364910.2","HORVU.BARKE.PROJ.4HG00304700.1",
    "HORVU.BARKE.PROJ.4HG00316830.1","HORVU.BARKE.PROJ.2HG00148330.1","HORVU.BARKE.PROJ.1HG00041660.1",
    "HORVU.BARKE.PROJ.4HG00760680.1","HORVU.BARKE.PROJ.7HG00668820.1","HORVU.BARKE.PROJ.1HG00070610.1"
  ),
  `Prolyl endopeptidase` = c(
    "HORVU.BARKE.PROJ.7HG00679290.1","HORVU.BARKE.PROJ.6HG00538760.1","HORVU.BARKE.PROJ.5HG00444180.1",
    "HORVU.BARKE.PROJ.6HG00504720.1","HORVU.BARKE.PROJ.2HG00170920.1","HORVU.BARKE.PROJ.5HG00441560.1"
  ),
  `Dipeptidyl peptidase` = c("HORVU.BARKE.PROJ.5HG00406500.1"),
  `Aspartyl aminopeptidase` = c("HORVU.BARKE.PROJ.5HG00436310.1","HORVU.BARKE.PROJ.3HG00298490.1"),
  `Oligopeptidase` = c(
    "HORVU.BARKE.PROJ.4HG00757340.1","HORVU.BARKE.PROJ.6HG00573710.1","HORVU.BARKE.PROJ.2HG00713210.1",
    "HORVU.BARKE.PROJ.2HG00711450.1","HORVU.BARKE.PROJ.2HG00714030.1"
  ),
  `Glutamyl endopeptidase` = c("HORVU.BARKE.PROJ.4HG00348310.1","HORVU.BARKE.PROJ.2HG00100820.1"),

  # --- OTHER ---
  `Catalase` = c("HORVU.BARKE.PROJ.6HG00495590.1","HORVU.BARKE.PROJ.7HG00681400.1","HORVU.BARKE.PROJ.4HG00374170.1"),
  `Phytase`  = c("HORVU.BARKE.PROJ.5HG00419620.1","HORVU.BARKE.PROJ.3HG00231330.1","HORVU.BARKE.PROJ.5HG00454460.1","HORVU.BARKE.PROJ.5HG00430080.1")
)



# -------------------------
# STEP 4 — Define family-to-category mapping
# -------------------------
family_category <- c(
  "Alpha-amylase"     = "Starch degradation",
  "Beta-amylase"      = "Starch degradation",
  "Alpha-glucosidase" = "Starch degradation",
  "Isoamylase"        = "Starch degradation",
  
  "Beta-glucanase"       = "Cell wall degradation",
  "Beta-glucosidase"     = "Cell wall degradation",
  "Arabinofuranosidase"  = "Cell wall degradation",
  "Mannanase"            = "Cell wall degradation",
  "Alpha-galactosidase"  = "Cell wall degradation",
  "Polygalacturonase"    = "Cell wall degradation",
  "Pectin esterase"      = "Cell wall degradation",
  "Xylan endohydrolase"  = "Cell wall degradation",
  "Beta-xylosidase"      = "Cell wall degradation",
  "Chitinase"            = "Cell wall degradation",
  
  "Phospholipase A1"         = "Lipid degradation",
  "Phospholipase A2_Hv"      = "Lipid degradation",
  "Phospholipase D"          = "Lipid degradation",
  "Phospholipase C"          = "Lipid degradation",
  "Lipase"                   = "Lipid degradation",
  "Lipoxygenase-1"           = "Lipid degradation",
  "Lipoxygenase-2"           = "Lipid degradation",
  "Hydroperoxide lyase"      = "Lipid degradation",
  "Allene oxide synthase_Hv" = "Lipid degradation",
  
  "Subtilisin-like protease"        = "Protein degradation",
  "Papain-like cysteine proteinase" = "Protein degradation",
  "Aminopeptidase"                  = "Protein degradation",
  "Methionyl-aminopeptidase"        = "Protein degradation",
  "Carboxypeptidase"                = "Protein degradation",
  "Prolyl endopeptidase"            = "Protein degradation",
  "Dipeptidyl peptidase"            = "Protein degradation",
  "Aspartyl aminopeptidase"         = "Protein degradation",
  "Oligopeptidase"                  = "Protein degradation",
  "Glutamyl endopeptidase"          = "Protein degradation",
  
  "Catalase" = "Other",
  "Phytase"  = "Other"
)

# -------------------------
# STEP 5 — Collect proteins per family
# -------------------------
malting_rows <- lapply(names(all_families_ID), function(fam) {
  ids <- all_families_ID[[fam]]
  sub <- final_HV3TS_mat %>% filter(Protein.IDs_pG %in% ids)
  if (nrow(sub) > 0) {
    sub$enzyme_family <- fam
    sub$Category <- family_category[[fam]]
    return(sub)
  } else return(NULL)
})

malting_all <- bind_rows(malting_rows)
if (nrow(malting_all) == 0) stop("❌ No proteins found for the provided IDs.")

# -------------------------
# STEP 6 — Impose the family order (handle duplicates safely)
# -------------------------
# Create ordered list of protein IDs according to `all_families_ID`
ordered_ids <- unlist(all_families_ID, use.names = FALSE)

# Remove duplicated entries while keeping the *first occurrence* to preserve order
ordered_ids_unique <- ordered_ids[!duplicated(ordered_ids)]

# Filter and reorder malting_all accordingly
malting_all <- malting_all %>%
  filter(Protein.IDs_pG %in% ordered_ids_unique) %>%
  mutate(Protein.IDs_pG = factor(Protein.IDs_pG, levels = ordered_ids_unique)) %>%
  arrange(factor(Protein.IDs_pG, levels = ordered_ids_unique))

cat(sprintf("✔ Imposed custom enzyme order (%d unique proteins).\n", length(ordered_ids_unique)))
# -------------------------
# STEP 7 — Build numeric matrix (keep order)
# -------------------------
malting_mat <- as.matrix(malting_all[, ..int_cols])
mode(malting_mat) <- "numeric"
malting_mat[!is.finite(malting_mat)] <- NA
rownames(malting_mat) <- make.unique(as.character(malting_all$Protein.IDs_pG))

# -------------------------
# STEP 8 — Annotation metadata
# -------------------------
annotation_row <- data.frame(
  Family   = malting_all$enzyme_family,
  Category = malting_all$Category
)
rownames(annotation_row) <- rownames(malting_mat)

family_names <- names(all_families_ID)
category_names <- unique(family_category)

annotation_row$Family   <- factor(annotation_row$Family,   levels = family_names)
annotation_row$Category <- factor(annotation_row$Category, levels = category_names)

# -------------------------
# STEP 9 — Colors
# -------------------------
palette <- colorRampPalette(brewer.pal(12, "Set3"))(length(family_names))
family_colors <- setNames(palette, family_names)

category_colors <- c(
  "Starch degradation"    = "#FFD92F",
  "Cell wall degradation" = "#A6D854",
  "Lipid degradation"     = "#E78AC3",
  "Protein degradation"   = "#FC8D62",
  "Other"                 = "#8DA0CB"
)

ann_colors <- list(Family = family_colors, Category = category_colors)

# -------------------------
# STEP 10 — Export heatmaps (imposed order, no clustering)
# -------------------------
out_dir <- "../../../R_export/HV3TS/malting_enzymes"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
heat_colors <- colorRampPalette(c("#4579F1", "white", "#D73020"))(130)

# JPEG
jpeg(file.path(out_dir, "HV3TS_allFamilies_ordered.jpeg"), width = 6, height = 10, units = "in", res = 400)
pheatmap(
  malting_mat,
  scale = "row",
  color= heat_colors, 
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_row = annotation_row,
  annotation_colors = ann_colors,
  show_rownames = TRUE,
  show_colnames = TRUE,
  legend = TRUE,
  annotation_legend = TRUE,
  fontsize_row = 6,
  fontsize_col = 8,
  # ↓↓↓ reduce legend bar height ↓↓↓
  legend_labels_angle = 0,
  legend_key_height = unit(0.01, "cm")   # smaller colorbar height
)
dev.off()
# PDF
pdf(file.path(out_dir, "HV3TS_allFamilies_ordered.pdf"), width = 6, height = 8)
pheatmap(
  malting_mat,
  scale = "row",
  color = heat_colors,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_row = annotation_row,
  annotation_colors = ann_colors,
  show_rownames = TRUE,
  show_colnames = TRUE,
  legend = TRUE,
  annotation_legend = TRUE,
  fontsize_row = 6,
  fontsize_col = 8,
  main = "HV3TS: Enzyme family expression (fixed order, row-scaled log2 intensities)",
  legend_labels_angle = 0,
  legend_key_height = unit(0.2, "cm")
)
dev.off()

# -------------------------
# STEP 11 — Heatmap for “Starch degradation” family only
# -------------------------

# Filter for starch degradation family
starch_idx <- annotation_row$Category == "Starch degradation"
malting_mat_starch <- malting_mat[starch_idx, ]
annotation_row_starch <- annotation_row[starch_idx, , drop = FALSE]

# Output directory
out_dir_starch <- file.path(out_dir, "starch_only")
if (!dir.exists(out_dir_starch)) dir.create(out_dir_starch, recursive = TRUE)

# Heatmap color palette (same as before)
heat_colors <- colorRampPalette(c("#4579F1", "white", "#D73020"))(130)

# --- JPEG export ---
jpeg(file.path(out_dir_starch, "HV3TS_starchFamily_ordered.jpeg"),
     width = 8, height = 6, units = "in", res = 400)
pheatmap(
  malting_mat_starch,
  scale = "row",
  color = heat_colors,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_row = annotation_row_starch,
  annotation_colors = ann_colors,
  show_rownames = TRUE,
  show_colnames = TRUE,
  legend = TRUE,
  annotation_legend = TRUE,
  fontsize_row = 7,
  fontsize_col = 8,
  legend_breaks = c(-2, -1, 0, 1, 2),
  legend_labels = c("-2", "-1", "0", "1", "2"),
  legend_labels_angle = 0,
  legend_key_height = unit(0.15, "cm")
)
dev.off()

# --- PDF export ---
pdf(file.path(out_dir_starch, "HV3TS_starchFamily_ordered.pdf"),
    width = 8, height = 6)
pheatmap(
  malting_mat_starch,
  scale = "row",
  color = heat_colors,
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  annotation_row = annotation_row_starch,
  annotation_colors = ann_colors,
  show_rownames = TRUE,
  show_colnames = TRUE,
  legend = TRUE,
  annotation_legend = TRUE,
  fontsize_row = 7,
  fontsize_col = 8,
  main = "HV3TS: Starch degradation enzymes (row-scaled log2 intensities)",
  legend_breaks = c(-2, -1, 0, 1, 2),
  legend_labels = c("-2", "-1", "0", "1", "2"),
  legend_labels_angle = 0,
  legend_key_height = unit(0.25, "cm")
)
dev.off()


```



