---
title: "DApipe_HvGerm_Exp5"
author: "Guido Giordano"
date: "2025-07-31"
output: html_document
---
#This is the pipeline for data analysis for barley germination experiment 6 

#Load packages
```{r}
BiocManager::install("cleave")
BiocManager::install("Mfuzz")
BiocManager::install("Biobase")
BiocManager::install("org.At.tair.db")
BiocManager::install("clusterProfiler")

library(AnnotationDbi)f
library(org.At.tair.db)
library(clusterProfiler)
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(ggbreak)
library(cowplot)
library(stringr)
library(kableExtra)
library(clusterProfiler)
library(mice)
library(reshape2)
library(data.table)
library(cleaver)
library(Mfuzz)
library(Biobase)
library(textshape)
library(tidyr)
library(gridExtra)
library(grid)
library(pcaMethods)
library(impute)
library(pheatmap)
library(ggh4x)
```
#Calculate the digestion efficiency. This is directly connected with the goodness of protein identification. 

```{r}
#Load evidence (TD) file. 

evidence_b1 <- read.table("../../../R_import/Hv5/evidence/evidence2.txt", header = TRUE, sep = "\t")

#Generate a bar plot from the column missed cleavages
# Batch 1
Try_Dig_b1 <- ggplot(evidence_b1, aes(x = factor(Missed.cleavages))) +
  geom_bar(fill = "#105CB6", color = "black") +
  labs(title = "Tryptic digestion efficiency (Hv5)",
       x = "Missed Cleavages",
       y = "Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, family = "", face = "plain"),  
        axis.title = element_text(size = 12, family = "", face = "plain"),
        axis.text = element_text(size = 12, family = "", face = "plain")) +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)),
        axis.title.x = element_text(margin = margin(t = 8)),
        axis.title.y = element_text(margin = margin(r = 15)))
ggsave("../../../R_export/Hv5/trypdig/Hv5_b1_trypdig.png")


# Save the combined plot as a single image
ggsave("../../../R_export/Hv5/trypdig/combo_tryp_dig_Hv5.png", 
       plot = Try_Dig_b1, scale=0.8)

```


#Calculate the total number of MS3 spectra acquired per batch

```{r}

# List of file paths and batch labels
summary_b1 <- read.table("../../../R_import/Hv5/summary/summary.txt", header = TRUE, sep = "\t")


# ---- Experimente bereinigen ----
experiments <- summary_b1 %>%
  mutate(Experiment = str_trim(as.character(Experiment))) %>%
  filter(!is.na(Experiment) & Experiment != "") %>%
  distinct(Experiment) %>%
  pull()

# ---- Plot-Funktion ----
plot_ms3_per_fraction <- function(df, exp_label, max_fraction = 48) {
  df_exp <- df %>%
    filter(str_trim(Experiment) == exp_label) %>%
    filter(!is.na(Fraction)) %>%
    mutate(Fraction = suppressWarnings(as.integer(as.character(Fraction)))) %>%
    group_by(Fraction) %>%
    summarise(MS3 = sum(MS3, na.rm = TRUE), .groups = "drop") %>%
    tidyr::complete(Fraction = 1:max_fraction, fill = list(MS3 = 0)) %>%
    mutate(Fraction = factor(Fraction, levels = 1:max_fraction))

  ggplot(df_exp, aes(x = Fraction, y = MS3)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_minimal() +
    labs(
      title = paste0("MS3 scans (", exp_label, ")"),
      x = NULL,
      y = "MS3 Identifications"
    ) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, size = 6),
      plot.title  = element_text(size = 12, hjust = 0.5)
    )
}

# ---- Plots für alle Experimente erzeugen ----
plots <- lapply(experiments, function(exp) plot_ms3_per_fraction(summary_b1, exp))

# ---- Kombinierter Plot ----
combined_plot <- wrap_plots(plots, ncol = 2)

# Anzeigen
print(combined_plot)

# ---- Optional speichern ----
ggsave("../../../R_export/Hv5/ms3_scans/ms3_Hv5.png", combined_plot, width = 14, height = 10, dpi = 300, scale = 0.5)

```

#Calculate the sequence coverage 

```{r}
# --- Define input paths and output ---
pgroup_files <- list(
  "Batch 1" = "../../../R_import/Hv5/pg_ibaq/pG_BAR_standard.csv",
  "Batch 2" = "../../../R_import/Hv5/pg_ibaq/pG_BAR_heat.csv",
  "Batch 3" = "../../../R_import/Hv5/pg_ibaq/pG_BAR_drought.csv",
  "Batch 4" = "../../../R_import/Hv5/pg_ibaq/pG_BAR_combined.csv"
)

output_dir <- "../../../R_export/Hv5/sequence_cov/"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

# --- Read files into list ---
pGroups_list <- lapply(pgroup_files, function(path) {
  read.csv(path, header = TRUE, stringsAsFactors = FALSE)
})
names(pGroups_list) <- names(pgroup_files)

# --- Generate one plot per batch ---
seqcov_plots <- list()

for (batch_name in names(pGroups_list)) {
  df <- pGroups_list[[batch_name]]
  coverage <- df[[33]]  # column 21: sequence coverage

  p <- ggplot(data.frame(Coverage = coverage), aes(x = "", y = Coverage)) +
    geom_violin(trim = FALSE, fill = "lightblue", color = "black") +
    geom_boxplot(width = 0.1, fill = "white", color = "black", outlier.size = 0.5) +
    labs(
      title = paste0("seqcov_", tolower(gsub(" ", "", batch_name))),
      x = NULL,
      y = "Sequence Coverage"
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(size = 12, face = "plain", hjust = 0.5),
      axis.title = element_text(size = 10),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  
  seqcov_plots[[length(seqcov_plots) + 1]] <- p
}

# --- Combine plots with patchwork ---
combined_seqcov <- (seqcov_plots[[1]] | seqcov_plots[[2]]) /
                   (seqcov_plots[[3]] | seqcov_plots[[4]])

# --- Save combined figure ---
ggsave(
  filename = file.path(output_dir, "combined_seqcov_Hv5.png"),
  plot = combined_seqcov,
  width = 12,
  height = 10,
  dpi = 300, 
  scale=0.6
)

message("Saved combined sequence coverage plot.")


```

#Check data distribution and structure 
Here using ibaq protein groups after conversion with ibaq pipeline 
```{r}
# --- Load files ---
data_dir <- "../../../R_import/Hv5/pG_ibaq/"
file_list <- list.files(path = data_dir, pattern = "^pG_.*\\.csv$", full.names = TRUE)

for (file in file_list) {
  var_name <- gsub("^.*/pG_(.*)\\.csv$", "pG_\\1", file)
  assign(var_name, fread(file), envir = .GlobalEnv)
}

# --- Remove reverse + rows with all last 11 < 0.1 ---
for (var_name in ls(pattern = "^pG_")) {
  df <- get(var_name)
  n_before <- nrow(df)
  
  # remove reverse hits
  if ("Reverse_pG" %in% colnames(df)) {
    df <- df[df$Reverse_pG != "+", ]
  }
  
  # find last 11 columns
  last_11_cols <- tail(colnames(df), 11)
  mat <- as.matrix(df[, ..last_11_cols])
  
  # keep rows that are not all <0.1
  keep <- rowSums(mat < 0.1, na.rm = TRUE) < 11
  df <- df[keep, ]
  
  n_after <- nrow(df)
  cat(var_name, ": removed", n_before - n_after, "rows\n")
  
  assign(var_name, df, envir = .GlobalEnv)
}

# --- Log2 Transformation of iBAQ Columns ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)
  ibaq_cols <- grep("summed_iBAQ", colnames(df), value = TRUE)
  df[, (ibaq_cols) := lapply(.SD, function(x) {
    x <- as.numeric(as.character(x))
    x[is.na(x)] <- 0
    log2(x + 1)
  }), .SDcols = ibaq_cols]
  assign(var_name, df, envir = .GlobalEnv)
}


# --- Median Centering of iBAQ Columns ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)
  ibaq_cols <- grep("summed_iBAQ", colnames(df), value = TRUE)
  df[, (ibaq_cols) := lapply(.SD, function(x) x - median(x, na.rm = TRUE)), .SDcols = ibaq_cols]
  assign(var_name, df, envir = .GlobalEnv)
}

# --- Output directories ---
plot_dir <- "../../../R_export/Hv5/data_distribution/"
heatmap_dir <- "../../../R_export/Hv5/heatmaps/"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(heatmap_dir, showWarnings = FALSE, recursive = TRUE)

# --- Plotting & Heatmap Generation ---
histogram_plots <- list()
heatmap_plots <- list()

for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)

  if (nrow(df) == 0) {
    message("Skipping ", var_name, " - dataset is empty.")
    next
  }

  ## iBAQ histogram plots
  ibaq_cols_hist <- grep("_summed_iBAQ$", colnames(df), value = TRUE)
  if (length(ibaq_cols_hist) > 0) {
    plots <- lapply(seq_along(ibaq_cols_hist), function(i) {
      ggplot(df, aes_string(x = ibaq_cols_hist[i])) +
        geom_histogram(binwidth = 0.7, fill = "#105CB6", color = "black") +
        labs(title = paste("Batch", i), x = "Reporter Intensity (iBAQ)", y = "Frequency") +
        theme_classic() +
        xlim(-15, 15) +
        theme(
          plot.title = element_text(size = 12, hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 10)
        )
    })

    if (length(plots) > 0) {
      title <- grid::textGrob(var_name, gp = grid::gpar(fontsize = 16, fontface = "bold"))
      plot_grid <- gridExtra::arrangeGrob(grobs = plots, ncol = 3, top = title)

      ggplot2::ggsave(
        filename = file.path(plot_dir, paste0(var_name, "_iBAQ_distribution.png")),
        plot = plot_grid,
        width = 12,
        height = 9, 
        scale=0.9
      )

      # For patchwork: use first plot + title
      first_plot <- plots[[1]] + ggtitle(paste0("histogram_", sub(".*b(\\d+).*", "b\\1", var_name)))
      histogram_plots[[length(histogram_plots) + 1]] <- first_plot
      message("Saved distribution plot for ", var_name)
    }
  }

  ## Heatmap generation
  ibaq_cols_heatmap <- grep("_iBAQ$", colnames(df), value = TRUE)
  if (length(ibaq_cols_heatmap) == 11) {
    mat_raw <- as.matrix(df[, ..ibaq_cols_heatmap])
    rownames(mat_raw) <- df$Protein.IDs_pG
    mat_raw[is.na(mat_raw)] <- -23
    mat_scaled <- t(scale(t(mat_raw)))
    mat_scaled <- mat_scaled[apply(mat_scaled, 1, function(x) all(!is.na(x))), ]

    # Save static PNG
    heatmap_file <- file.path(heatmap_dir, paste0(var_name, "_heatmap_scaled.png"))
    pheatmap::pheatmap(mat_scaled,
      filename = heatmap_file,
      cluster_rows = TRUE,
      cluster_cols = FALSE,
      show_rownames = FALSE,
      labels_col = paste0("channel", 1:11)
    )
    message("Saved heatmap for ", var_name)

    # For patchwork (render in R, not saving image)
    heatmap_plot <- pheatmap::pheatmap(
      mat_scaled,
      silent = TRUE,
      cluster_rows = TRUE,
      cluster_cols = FALSE,
      show_rownames = FALSE,
      labels_col = paste0("channel", 1:11),
      main = paste0("heatmap_", sub(".*b(\\d+).*", "b\\1", var_name))
    )
    heatmap_plots[[length(heatmap_plots) + 1]] <- grid::grid.grabExpr(grid::grid.draw(heatmap_plot$gtable))
  } else {
    message("Skipping heatmap for ", var_name, " - found ", length(ibaq_cols_heatmap), " _iBAQ columns")
  }
}

# --- Combine and save 2x2 grid of histograms ---
if (length(histogram_plots) >= 4) {
  combined_hist <- (histogram_plots[[1]] | histogram_plots[[2]]) /
                   (histogram_plots[[3]] | histogram_plots[[4]])

  ggplot2::ggsave(
    filename = file.path(plot_dir, "combined_histograms.png"),
    plot = combined_hist,
    width = 16,
    height = 14, 
    scale=0.9
  )
  message("Saved combined histogram figure.")
}

# --- Combine and save 2x2 grid of heatmaps ---
if (length(heatmap_plots) >= 4) {
  combined_heat <- (wrap_elements(heatmap_plots[[1]]) | wrap_elements(heatmap_plots[[2]])) /
                   (wrap_elements(heatmap_plots[[3]]) | wrap_elements(heatmap_plots[[4]]))

  ggplot2::ggsave(
    filename = file.path(heatmap_dir, "combined_heatmaps.png"),
    plot = combined_heat,
    width = 16,
    height = 12, 
    scale=0.6
  )
  message("Saved combined heatmap figure.")
}

```

##Normalisation 
- In this script we take our protein groups with ibaq values and we filter, transform and normalise them so that they are ready to be used for quantiative comparisons  

remember to modify this in the script to exclude bridge from inputation 
intensity_cols <- setdiff(names(df)[2:11], "iBAQ_std_0hs")



```{r}
## ============================================================
## Hv5 — Adaptive Imputation + TSN (protecting imputed cells) + Bridge-based Inter-Batch Scaling
## ============================================================

suppressPackageStartupMessages({
  library(data.table)
  library(ggplot2)
  library(pheatmap)
  library(magick)
})

## =========================
## Moving-average smoothing (k = 3)
## =========================
smooth_transition <- function(x, k = 3) {
  if (all(is.na(x))) return(x)
  x_ext <- c(rep(x[1], floor(k/2)), x, rep(x[length(x)], floor(k/2)))
  y <- stats::filter(x_ext, rep(1/k, k), sides = 2)
  y <- y[(floor(k/2)+1):(length(y)-floor(k/2))]
  as.numeric(y)
}

smooth_matrix <- function(mat, k = 3) {
  t(apply(mat, 1, smooth_transition, k = k))
}

## --- Parameters ---
threshold <- 4
min_val   <- 0.1
heat_colors <- colorRampPalette(c("#4579F1", "white", "#D73020"))(200)
mod_masks <- new.env(parent = emptyenv())  # mask storage

## --- Condition-specific thresholds ---
cond_thresholds <- list(
  "standard" = 3,
  "heat"     = 4,
  "drought"  = 4,
  "combined" = 4
)

## --- Condition-specific channel correction rules ---
cond_specific_rules <- list(
  "standard" = list(
    list(target = 8, neighbors = c(7, 9)),
    list(target = 6, neighbors = c(5, 7))
  )
)

## =========================
## 0) Load files
## =========================
data_dir <- "../../../R_import/Hv5/pG_ibaq/"
file_list <- list.files(path = data_dir, pattern = "^pG_.*\\.csv$", full.names = TRUE)
if (!length(file_list)) stop("❌ No pG_*.csv files found in: ", data_dir)

for (file in file_list) {
  var_name <- gsub("^.*/pG_(.*)\\.csv$", "pG_\\1", file)
  assign(var_name, fread(file), envir = .GlobalEnv)
}

## =========================
## 1) Clean: remove Reverse; drop rows with all last 11 < 0.1
## =========================
for (var_name in ls(pattern = "^pG_")) {
  df <- get(var_name)
  n_before <- nrow(df)

  if ("Reverse_pG" %in% names(df)) df <- df[df$Reverse_pG != "+", ]
  else if ("Reverse" %in% names(df)) df <- df[df$Reverse != "+", ]

  last_11_cols <- tail(names(df), 11)
  mat <- suppressWarnings(as.matrix(df[, ..last_11_cols]))
  keep <- rowSums(mat < 0.1, na.rm = TRUE) < 11
  df <- df[keep, , drop = FALSE]

  n_after <- nrow(df)
  cat(sprintf("%s: removed %d rows (low intensity)\n", var_name, n_before - n_after))
  assign(var_name, as.data.table(df), .GlobalEnv)
}

## =========================
## 2) Rename, reorder, and merge 0 h replicates (keep bridge)
## =========================
for (var_name in ls(pattern = "^pG_")) {
  df <- get(var_name)
  if (ncol(df) < 12) next

  last_11_idx <- tail(seq_len(ncol(df)), 11)
  df <- df[, c(1, last_11_idx), with = FALSE]

  cond <- tolower(gsub("^pG_(BAR_)?", "", var_name))
  cond <- dplyr::case_when(
    grepl("standard|std", cond) ~ "standard",
    grepl("heat", cond)         ~ "heat",
    grepl("drought", cond)      ~ "drought",
    grepl("comb|combined|combo", cond) ~ "combined",
    TRUE ~ cond
  )

  prot_col <- names(df)[1]
  chan_cols <- names(df)[2:12]
  new_order <- c(chan_cols[11], chan_cols[1:10])
  df <- df[, c(prot_col, new_order), with = FALSE]

  new_names <- c(
    "iBAQ_std_0hs",
    paste0("iBAQ_", cond, "_0hs_r1"),
    paste0("iBAQ_", cond, "_0hs_r2"),
    paste0("iBAQ_", cond, "_", seq(12, 96, by = 12), "hs")
  )
  setnames(df, old = names(df)[2:12], new = new_names)

  df[, iBAQ_mean_0hs := rowMeans(.SD, na.rm = TRUE),
     .SDcols = paste0("iBAQ_", cond, c("_0hs_r1", "_0hs_r2"))]
  df <- df[, !paste0("iBAQ_", cond, c("_0hs_r1", "_0hs_r2")), with = FALSE]

  setcolorder(df, c("Protein.IDs_pG", "iBAQ_std_0hs", "iBAQ_mean_0hs",
                    paste0("iBAQ_", cond, "_", seq(12, 96, by = 12), "hs")))
  assign(var_name, df, .GlobalEnv)
}

## =========================
## 3) Adaptive Imputation + Outlier Correction + Mask build
## =========================
cat("\n--- Adaptive imputation & outlier correction ---\n")

for (var_name in ls(pattern = "^pG_")) {
  df <- get(var_name)
  if (is.null(df) || ncol(df) < 11) next

  cond <- tolower(gsub("^pG_(BAR_)?", "", var_name))
  cond <- dplyr::case_when(
    grepl("standard|std", cond) ~ "standard",
    grepl("heat", cond)         ~ "heat",
    grepl("drought", cond)      ~ "drought",
    grepl("comb|combined|combo", cond) ~ "combined",
    TRUE ~ cond
  )

  thr <- ifelse(cond %in% names(cond_thresholds), cond_thresholds[[cond]], threshold)
  rules <- cond_specific_rules[[cond]]

  cat(sprintf("→ %s (condition = %s, threshold = %.1f)\n", var_name, cond, thr))

  intensity_cols <- names(df)[2:11] #here the correct thing to do is to exclude the bridge channel from adaptative imputation but currently we are not doing it 
  #intensity_cols <- setdiff(names(df)[2:11], "iBAQ_std_0hs")

  df[, (intensity_cols) := lapply(.SD, as.numeric), .SDcols = intensity_cols]
  mask <- matrix(FALSE, nrow = nrow(df), ncol = length(intensity_cols))

  n_imp <- 0; n_corr <- 0

  for (i in seq_len(nrow(df))) {
    row_vals <- unlist(df[i, ..intensity_cols])

    # --- Step 1: custom rules
    if (!is.null(rules)) {
      for (rule in rules) {
        target <- rule$target
        neigh_idx <- rule$neighbors
        if (all(neigh_idx %in% seq_along(row_vals))) {
          val <- row_vals[target]
          neigh_vals <- row_vals[neigh_idx]
          neigh_vals <- neigh_vals[is.finite(neigh_vals) & neigh_vals > 0]
          if (length(neigh_vals)) {
            neigh_mean <- mean(neigh_vals)
            if (is.na(val) || val < min_val) {
              row_vals[target] <- neigh_mean
              mask[i, target] <- TRUE; n_imp <- n_imp + 1
            } else {
              log2fc <- log2((val + 1) / (neigh_mean + 1))
              if (abs(log2fc) > thr) {
                row_vals[target] <- neigh_mean
                mask[i, target] <- TRUE; n_corr <- n_corr + 1
              }
            }
          }
        }
      }
    }

    # --- Step 2: general adaptive correction
    rule_targets <- if (!is.null(rules)) vapply(rules, `[[`, numeric(1), "target") else integer(0)
    for (j in seq_along(row_vals)) {
      if (j %in% rule_targets) next
      val <- row_vals[j]
      neigh_idx <- c(j - 1L, j + 1L)
      neigh_idx <- neigh_idx[neigh_idx >= 1L & neigh_idx <= length(row_vals)]
      neigh_vals <- row_vals[neigh_idx]
      neigh_vals <- neigh_vals[is.finite(neigh_vals) & neigh_vals > 0]
      if (!length(neigh_vals)) next
      neigh_mean <- mean(neigh_vals)
      if (is.na(val) || val < min_val) {
        row_vals[j] <- neigh_mean; mask[i, j] <- TRUE; n_imp <- n_imp + 1
      } else {
        log2fc <- log2((val + 1) / (neigh_mean + 1))
        if (abs(log2fc) > thr) {
          row_vals[j] <- neigh_mean; mask[i, j] <- TRUE; n_corr <- n_corr + 1
        }
      }
    }
    df[i, (intensity_cols) := as.list(row_vals)]
  }

  cat(sprintf("✅ %s — corrected: %d (imputed=%d, outliers=%d)\n",
              var_name, n_imp + n_corr, n_imp, n_corr))
  assign(var_name, df, .GlobalEnv)
  mod_masks[[var_name]] <- mask
}

## =========================
## 4) QC function
## =========================
make_qc_plot <- function(df, title) {
  intensity_cols <- names(df)[-1]
  sums <- colSums(df[, ..intensity_cols], na.rm = TRUE)
  meds <- vapply(intensity_cols, function(cn) median(df[[cn]], na.rm = TRUE), numeric(1))
  stats <- data.table(Channel = factor(intensity_cols, levels = intensity_cols),
                      Total_Intensity = sums, Median_Intensity = meds)
  max_total <- max(stats$Total_Intensity, na.rm = TRUE)
  max_med   <- max(stats$Median_Intensity, na.rm = TRUE)
  scale_factor <- if (is.finite(max_total / max_med) && max_med > 0) max_total / max_med else 1
  ggplot(stats, aes(x = Channel)) +
    geom_bar(aes(y = Total_Intensity), stat = "identity", fill = "steelblue", alpha = 0.6) +
    geom_point(aes(y = Median_Intensity * scale_factor), color = "red", size = 2.5) +
    geom_line(aes(y = Median_Intensity * scale_factor, group = 1), color = "red") +
    scale_y_continuous(name = "Total Intensity",
                       sec.axis = sec_axis(~ . / scale_factor, name = "Median Intensity")) +
    labs(title = title, x = NULL) +
    theme_classic() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

## =========================
## 5) Output directories
## =========================
out_dir_pre   <- "../../../R_export/Hv5/TS_preNorm"
out_dir_intra <- "../../../R_export/Hv5/TS_intraNorm"
out_dir_inter <- "../../../R_export/Hv5/TS_postInterNorm"

for (d in c(out_dir_pre, out_dir_intra, out_dir_inter)) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}

## =========================
## 6) PRE-NORM QC
## =========================
cat("\n--- Pre-normalization QC ---\n")
for (nm in ls(pattern = "^pG_")) {
  df <- get(nm)
  ggsave(file.path(out_dir_pre, paste0(nm, "_preNorm_sumMedian.png")),
         make_qc_plot(df, paste("Total & Median Intensity —", nm, "(pre-norm)")),
         width = 10, height = 5, dpi = 150)
}

## =========================
## 7) INTRA-BATCH TOTAL-SUM NORMALIZATION + MOVING AVERAGE SMOOTHING
## =========================
cat("\n--- Intra-batch total-sum normalization + smoothing (protecting modified cells) ---\n")
batch_names <- ls(pattern = "^pG_")

for (nm in batch_names) {

  df <- get(nm)
  cols_norm <- setdiff(names(df), c("Protein.IDs_pG", "iBAQ_std_0hs"))

  mask <- mod_masks[[nm]]
  if (is.null(mask)) mask <- matrix(FALSE, nrow = nrow(df), ncol = length(cols_norm))

  ## -----------------------------
  ## Total Sum Normalization (TSN)
  ## -----------------------------
  col_sums <- sapply(seq_along(cols_norm), function(j) {
    v <- df[[cols_norm[j]]]
    idx <- !mask[, j] & is.finite(v)
    sum(v[idx], na.rm = TRUE)
  })

  target <- median(col_sums[is.finite(col_sums)], na.rm = TRUE)
  factors <- target / col_sums

  for (j in seq_along(cols_norm)) {
    v <- df[[cols_norm[j]]]
    idx <- !mask[, j] & is.finite(v)
    v[idx] <- v[idx] * factors[j]
    df[[cols_norm[j]]] <- v
  }

  ## ----------------------------------------
  ## NEW: Edge-replicated moving average (k=3)
  ## ----------------------------------------
  mat_postTSN <- as.matrix(df[, ..cols_norm])
  mat_smooth  <- smooth_matrix(mat_postTSN, k = 3)

  df[, (cols_norm) := as.data.table(mat_smooth)]

  cat(sprintf("✨ %s — Moving-average smoothing applied (k=3)\n", nm))

  assign(nm, df, .GlobalEnv)
}

## =========================
## 8) INTER-BATCH SCALING (bridge-based, using MEDIAN intensity)
## =========================
cat("\n--- Inter-batch scaling using bridge channel MEDIAN (protecting modified cells) ---\n")

# OLD:
# bridge_totals <- vapply(batch_names, function(nm) {
#   df <- get(nm)
#   sum(df[["iBAQ_std_0hs"]], na.rm = TRUE)
# }, numeric(1))

# NEW: use MEDIAN OF INTENSITIES, not SUM
bridge_medians <- vapply(batch_names, function(nm) {
  df <- get(nm)
  median(df[["iBAQ_std_0hs"]], na.rm = TRUE)
}, numeric(1))

# Target level = median of bridge medians
grand_level <- median(bridge_medians, na.rm = TRUE)

# Scaling factors = target median / batch median
batch_factors <- grand_level / bridge_medians

for (nm in batch_names) {
  df <- get(nm)
  fac <- batch_factors[[nm]]
  mask <- mod_masks[[nm]]
  if (is.null(mask)) mask <- matrix(FALSE, nrow = nrow(df), ncol = ncol(df) - 1)
  cols <- setdiff(names(df), "Protein.IDs_pG")

  for (j in seq_along(cols)) {
    v <- df[[cols[j]]]
    idx <- !mask[, j] & is.finite(v)
    v[idx] <- v[idx] * fac
    df[[cols[j]]] <- v
  }
  assign(nm, df, .GlobalEnv)
}

## =========================
## 9) REMOVE BRIDGE before QC + export
## =========================
for (nm in batch_names) {
  df <- get(nm)
  if ("iBAQ_std_0hs" %in% names(df)) df <- df[, !c("iBAQ_std_0hs"), with = FALSE]
  assign(nm, df, .GlobalEnv)
}

## =========================
## 10) POST-INTER QC + COLLAGES
## =========================
cat("\n--- Post-inter QC ---\n")
for (nm in batch_names) {
  df <- get(nm)
  ggsave(file.path(out_dir_inter, paste0(nm, "_postInter_sumMedian.png")),
         make_qc_plot(df, paste("Total & Median Intensity —", nm, "(post-inter)")),
         width = 10, height = 5, dpi = 150)
}

## =========================
## 11) Combine QC plots
## =========================
out_dir_combo <- "../../../R_export/Hv5/TS_postInterNorm_combined"
if (!dir.exists(out_dir_combo)) dir.create(out_dir_combo, recursive = TRUE)

bar_files <- list.files(out_dir_inter, pattern = "sumMedian\\.png$", full.names = TRUE)
if (length(bar_files)) {
  imgs <- lapply(bar_files, function(f) image_resize(image_read(f), "1600x1600"))
  collage <- image_montage(image_join(imgs), tile = "2x2", geometry = "+10+10")
  image_write(collage, file.path(out_dir_combo, "AllBatches_postInter_BarplotCollage.png"), format = "png")
  cat("✅ Combined barplot collage saved.\n")
}

## =========================
## 12) EXPORT FINAL log2(+1)
## =========================
final_out_dir <- "../../../R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly"
if (!dir.exists(final_out_dir)) dir.create(final_out_dir, recursive = TRUE)

for (nm in batch_names) {
  df <- get(nm)
  cols <- names(df)[-1]
  df[, (cols) := lapply(.SD, function(x) log2(as.numeric(x) + 1)), .SDcols = cols]
  fwrite(df, file.path(final_out_dir, paste0(nm, "_norm_log2.csv")))
}

cat("\n✅ Hv5 pipeline completed successfully.\n")



```

##In this script we add intensity values and annotations to our protein groups
In this case is easier because we don't need to perform orthology analysis, we just add uniprot IDs to our pangenome IDs and then add intensity values 

```{r}
#Load the DB matches between the species specific fasta and the uniprot species specific fasta, so that each fasta entry is associated with a uniprot entry 
# set the folder path
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_up"

# Load each file separately
barley_hvup <- read.delim(file.path(path, "barke_upbarley.tsv"), sep = "\t", header = FALSE, stringsAsFactors = FALSE)
# Keep only first 2 columns and rename them consistently
barley_hvup  <- barley_hvup[, 1:2]
colnames(barley_hvup)  <- c("barleyup_entry", "barley_entry")

# Clean annotation files to allow matching 
barley_hvup$barleyup_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_hvup$barleyup_entry)

#deduplicate entries 
barley_hvup <- barley_hvup %>%
  distinct(barley_entry, .keep_all = TRUE)

##Attach functional annotation from uniprot to the uniprot matched entries 
# set the folder path
path <- "../../../R_import/Hv6/anno/masters_up"

# Load each file separately
up_Hv <- read.delim(file.path(path, "unip_Hv_1505.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Keep only first 2 columns and rename them consistently
up_Hv  <- up_Hv[, c(3,4,5,7,8,9)]
colnames(up_Hv)  <- c("barleyup_entryname", "barleyup_pname", "barleyup_gname", "barleyup_GO_bp", "barleyup_GO_mf", "barleyup_GO_cc")

#Attach the functional annotation to the DB matches and then to the orthology file. I set all.x=TRUE, because i dont want to have annotations for proteins that are not matching 

barley_anno <- merge(
  barley_hvup,
  up_Hv,
  by.x = "barleyup_entry",
  by.y = "barleyup_entryname",
  all.x = TRUE
)

#Now associate each barley protein to an arabidopsis protein and attach at uniprot annotations 
# set the folder path
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_upAt"

# Load each file separately
barley_ATup <- read.delim(file.path(path, "araT_barke.tsv"), sep = "\t", header = FALSE, stringsAsFactors = FALSE)

# Keep only first 2 columns and rename them consistently
barley_ATup  <- barley_ATup[, 1:2]
colnames(barley_ATup)  <- c("barley_entry",  "araT_entry")

# Clean annotation files to allow matching 
barley_ATup$araT_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_ATup$araT_entry)

#deduplicate entries 
barley_ATup <- barley_ATup %>%
  distinct(barley_entry, .keep_all = TRUE)

#Merge each annotation file with arabidopsis thaliana entries 
barley_anno_expat <- merge(
  barley_anno,
  barley_ATup,
  by = "barley_entry",
  all.x = TRUE
)

#Attach functional annotation from Atuniprot
At_UP_annotated <- read_tsv("../../../R_import/Hv6/anno/masters_up/unip_At_1505.tsv")
At_UP_annotated  <- At_UP_annotated[, c(2,3,4,5,6,7,8)]
At_UP_annotated <- At_UP_annotated %>%
  mutate(TAIR = sub(";.*", "", TAIR))

#Finally get the final annotation file 
barley_full_anno <- merge(
  barley_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)
barley_full_anno <- barley_full_anno[, c(2, 3, 4, 5, 6, 7, 8, 1, 14, 9, 10, 11, 12, 13)]
colnames(barley_full_anno) <- c("barley_entry","barleyup_entry","barleyup_pname","barleyup_gname",                               "barleyup_GO_bp","barleyup_GO_mf","barleyup_GO_cc","araT_entry",
                                "TAIR","atup_pname","atup_gname","at_GO_bp",
                                "at_GO_mf","at_GO_cc")


# Save the final annotation file

write.csv(
  barley_full_anno,
  "../../../R_Data/Hv5/func_anno/barley_full_anno.csv",
  row.names = FALSE
)

#Now add intensity values to the table

##Load protein group files normalised ----
# Where the script saved the final files
final_dir <- "../../../R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly"

# Find all final normalized + log2 files
final_files <- list.files(final_dir, pattern = "_norm_log2\\.csv$", full.names = TRUE)
if (!length(final_files)) stop("No final normalized files found in: ", final_dir)

# Helper: get a clean object name from file name (e.g., "pG_BAR_standard_norm_log2" -> "pG_BAR_standard")
name_from_file <- function(x) sub("_norm_log2\\.csv$", "", basename(x))

# Read all into a named list and also assign each to the Global Env with suffix "_final"
final_list <- vector("list", length(final_files))
names(final_list) <- sapply(final_files, name_from_file)

for (i in seq_along(final_files)) {
  dt <- fread(final_files[i])
  # keep as data.table
  dt <- as.data.table(dt)
  final_list[[i]] <- dt
  # also create a variable per file for convenience
  assign(paste0(names(final_list)[i], "_final"), dt, envir = .GlobalEnv)
}

##Merge the protein group files normalised in one matrix ----
## -------- settings --------
# suffixes per condition
suffix_map <- c(
  standard = "_std",
  heat     = "_h",
  drought  = "_d",
  combined = "_c"
)

# detect condition from object name
detect_condition <- function(objname) {
  nm <- tolower(objname)
  if (grepl("standard|std", nm))  return("standard")
  if (grepl("heat", nm))          return("heat")
  if (grepl("drought", nm))       return("drought")
  if (grepl("comb|combined|combo", nm)) return("combined")
  return("unknown")
}

## -------- collect candidate tables --------
cand_objs <- ls(pattern = "^pG_.*_final$")
if (!length(cand_objs)) stop("No *_final objects found in the environment.")

# build a list of data.tables, each with suffixed non-key columns
dt_list <- list()
for (nm in cand_objs) {
  dt <- as.data.table(get(nm))

  # standardize key column name to 'Protein.IDs_pG'
  key_col <- names(dt)[1]
  if (key_col != "Protein.IDs_pG") setnames(dt, key_col, "Protein.IDs_pG")

  cond <- detect_condition(nm)
  suf  <- suffix_map[cond]
  if (is.na(suf)) suf <- paste0("_", cond)  # fallback if 'unknown'

  # suffix all non-key columns to avoid any name collisions (e.g., iBAQ_std_0hs)
  nonkey <- setdiff(names(dt), "Protein.IDs_pG")
  setnames(dt, nonkey, paste0(nonkey, suf))

  dt_list[[nm]] <- dt
}

# sanity: show what we’re about to merge
message("Merging the following datasets (with suffixes):")
print(data.table(
  object = names(dt_list),
  condition = sapply(names(dt_list), detect_condition),
  nrows = sapply(dt_list, nrow),
  ncols = sapply(dt_list, ncol)
))

## -------- merge by key (full outer join) --------
merged_dt <- Reduce(function(x, y) merge(x, y, by = "Protein.IDs_pG", all = TRUE),
                    dt_list)

# optional: order columns by condition order (std, heat, drought, combined)
order_by_suffix <- function(cols) {
  # keep key first
  key <- "Protein.IDs_pG"
  others <- setdiff(cols, key)

  # group by suffix appearance
  grp <- list(
    std = grep("_std$", others, value = TRUE),
    h   = grep("_h$",   others, value = TRUE),
    d   = grep("_d$",   others, value = TRUE),
    c   = grep("_c$",   others, value = TRUE),
    rest = setdiff(others, c(
      grep("_std$", others, value = TRUE),
      grep("_h$",   others, value = TRUE),
      grep("_d$",   others, value = TRUE),
      grep("_c$",   others, value = TRUE)
    ))
  )
  c(key, unlist(grp, use.names = FALSE))
}
setcolorder(merged_dt, order_by_suffix(names(merged_dt)))

# inspect
message("Merged matrix dims: ", nrow(merged_dt), " proteins x ", ncol(merged_dt)-1, " channels")

## -------- (optional) save to disk --------
out_dir <- "../../../R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
out_file <- file.path(out_dir, "_MERGED_allBatches_wide_log2.csv")
fwrite(merged_dt, out_file)
message("Saved merged matrix to: ", out_file)

##Let's now merge annotations and intensities ----
setDT(merged_dt)
setDT(barley_full_anno)

# Left join by ID (keep all proteins from merged_dt)
final_Hv5_mat <- merge(
  merged_dt,
  barley_full_anno,
  by.x  = "Protein.IDs_pG",
  by.y  = "barley_entry",
  all.x = TRUE,
  sort  = FALSE
)

# (Optional) move annotation columns right after the ID for readability
anno_cols <- setdiff(names(barley_full_anno), "barley_entry")
setcolorder(
  final_Hv5_mat,
  c("Protein.IDs_pG", anno_cols, setdiff(names(final_Hv5_mat), c("Protein.IDs_pG", anno_cols)))
)

# (Optional) save
out_path <- "../../../R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly/Hv5_final.csv"
fwrite(final_Hv5_mat, out_path)

```

#Map carlsberg gene IDs to Barke based on my Diamond BLAST search (MorexV3 against Morex pangenome) and Manuel Spannagl orthofinder search (040126)


```{r}
bamy <- c(
  "HORVU.MOREX.PROJ.2HG00122200.1","HORVU.MOREX.PROJ.2HG00122210.1",
  "HORVU.MOREX.PROJ.1HG00035230.1","HORVU.MOREX.PROJ.1HG00035280.1",
  "HORVU.MOREX.PROJ.1HG00049680.1","HORVU.MOREX.PROJ.3HG00221430.1",
  "HORVU.MOREX.PROJ.4HG00294880.1","HORVU.MOREX.PROJ.4HG00365950.1",
  "HORVU.MOREX.PROJ.4HG00370340.1","HORVU.MOREX.PROJ.6HG00488540.1"
)

aamy <- c(
  "HORVU.MOREX.PROJ.2HG00144590.1","HORVU.MOREX.PROJ.3HG00252270.1",
  "HORVU.MOREX.PROJ.5HG00429890.1","HORVU.MOREX.PROJ.6HG00545380.1",
  "HORVU.MOREX.PROJ.6HG00545400.1","HORVU.MOREX.PROJ.6HG00545420.1",
  "HORVU.MOREX.PROJ.6HG00545440.1","HORVU.MOREX.PROJ.6HG00545450.1",
  "HORVU.MOREX.PROJ.6HG00545470.1","HORVU.MOREX.PROJ.6HG00547690.1",
  "HORVU.MOREX.PROJ.7HG00636260.1","HORVU.MOREX.PROJ.7HG00636410.1",
  "HORVU.MOREX.PROJ.7HG00636430.1"
)

aglu <- c(
  "HORVU.MOREX.PROJ.1HG00051530.1","HORVU.MOREX.PROJ.2HG00187140.1",
  "HORVU.MOREX.PROJ.2HG00187150.1","HORVU.MOREX.PROJ.2HG00187200.1",
  "HORVU.MOREX.PROJ.2HG00187220.1","HORVU.MOREX.PROJ.3HG00196410.1",
  "HORVU.MOREX.PROJ.3HG00227410.1","HORVU.MOREX.PROJ.4HG00350740.1",
  "HORVU.MOREX.PROJ.7HG00649600.1","HORVU.MOREX.PROJ.7HG00580490.1"
)

iso <- c(
  "HORVU.MOREX.PROJ.1HG00055940.1","HORVU.MOREX.PROJ.5HG00431450.1",
  "HORVU.MOREX.PROJ.7HG00602160.1","HORVU.MOREX.PROJ.2HG00145410.1",
  "HORVU.MOREX.PROJ.2HG00149670.1","HORVU.MOREX.PROJ.7HG00628700.1",
  "HORVU.MOREX.PROJ.7HG00663740.1","HORVU.MOREX.PROJ.7HG00830550.1",
  "HORVU.MOREX.PROJ.7HG00830560.1"
)


enzyme_sets <- list(
  bamy = bamy,
  aamy = aamy,
  aglu = aglu,
  iso  = iso
)


HO_N1 <- read_tsv("../../../R_import/Hv4/Ortho/N1.tsv")

map_fun <- function(ids) {
  HO_N1 %>%
    filter(Morex %in% ids, !is.na(Barke)) %>%
    group_by(Morex) %>%
    slice(1) %>%          # first valid Barke hit
    ungroup() %>%
    pull(Barke) %>%
    unique()
}



barke_vectors <- lapply(enzyme_sets, map_fun)

barke_csv <- lapply(barke_vectors, function(x)
  paste(x, collapse = ",")
)

barke_csv




##We also add a script with the quick mapping visual 
library(dplyr)
library(readr)

# 1:1 mapper: for each Morex, keep the first non-NA Barke hit
map_table_fun <- function(ids, set_name) {
  HO_N1 %>%
    filter(Morex %in% ids, !is.na(Barke)) %>%
    group_by(Morex) %>%
    slice(1) %>%                 # first valid Barke hit per Morex
    ungroup() %>%
    transmute(
      enzyme_set = set_name,
      Morex = Morex,
      Barke = Barke
    )
}

# mapping tables per enzyme set
mapping_tables <- Map(map_table_fun, enzyme_sets, names(enzyme_sets))

# combined mapping table (all enzyme sets)
mapping_1to1 <- bind_rows(mapping_tables) %>%
  distinct(enzyme_set, Morex, .keep_all = TRUE)

mapping_1to1
```

#Generate heatmaps from specific gene families 
In this chunk we generate heatmaps from specific gene families 

```{r}

## ============================================================
## Hv5 — Malting enzyme heatmaps (Hv4-style z-scores, scale="none")
## ============================================================

suppressPackageStartupMessages({
  library(data.table)
  library(dplyr)
  library(tidyr)
  library(pheatmap)
  library(RColorBrewer)
})

# --- 1) Load data ---
out_path <- "Z:/PhD Munchen/Guido personal PhD Project_1/Barley germination proteome/Resources for Data analysis/R/R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly/Hv5_final.csv"
final_Hv5_mat <- fread(out_path)

# --- 2) Select intensity columns (remove unwanted ones) ---
int_cols <- grep("^iBAQ_", names(final_Hv5_mat), value = TRUE)

# Explicitly drop undesired columns
drop_cols <- c(
  "iBAQ_std_0hs_std", "iBAQ_std_0hs_r2_std", "iBAQ_std_0hs_h",
  "iBAQ_heat_0hs_r2_h", "iBAQ_std_0hs_d", "iBAQ_drought_0hs_r2_d",
  "iBAQ_std_0hs_c", "iBAQ_combo_0hs_r2_c"
)
int_cols <- setdiff(int_cols, drop_cols)

if (length(int_cols) == 0) stop("❌ No valid intensity columns found after filtering unwanted ones.")

cat("✅ Retained", length(int_cols), "intensity columns for heatmap\n")



barke_gene_map <- c(

# =========================
# Alpha-amylase (Carlsberg group)
# =========================
"HORVU.BARKE.PROJ.6HG00559210.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00559200.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00559230.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00559240.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00559250.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00559270.1" = "Amy1-1",
"HORVU.BARKE.PROJ.6HG00561370.1" = "Amy1-2",
"HORVU.BARKE.PROJ.7HG00653200.1" = "Amy2-1",
"HORVU.BARKE.PROJ.7HG00653330.1" = "Amy2-3",
"HORVU.BARKE.PROJ.7HG00653310.1" = "Amy2-2",
"HORVU.BARKE.PROJ.5HG00441970.1" = "Amy3",
"HORVU.BARKE.PROJ.2HG00148940.1" = "Amy4-1",
"HORVU.BARKE.PROJ.3HG00259920.1" = "Amy4-2",

# =========================
# Beta-amylase (Carlsberg group)
# =========================
# Real gene symbols
"HORVU.BARKE.PROJ.2HG00126310.1" = "Bmy2",
"HORVU.BARKE.PROJ.4HG00380960.1" = "Bmy1",

# No real gene symbol
"HORVU.BARKE.PROJ.1HG00036700.1" = "chr1H3670",
"HORVU.BARKE.PROJ.1HG00036740.1" = "chr1H3674",
"HORVU.BARKE.PROJ.1HG00051030.1" = "chr1H5103",
"HORVU.BARKE.PROJ.2HG00105510.1" = "chr2H10551",
"HORVU.BARKE.PROJ.2HG00126320.1" = "chr2H12632",
"HORVU.BARKE.PROJ.3HG00228930.1" = "chr3H22893",
"HORVU.BARKE.PROJ.4HG00303690.1" = "chr4H30369",
"HORVU.BARKE.PROJ.4HG00376360.1" = "chr4H37636",
"HORVU.BARKE.PROJ.6HG00502080.1" = "chr6H50208",

# =========================
# Limit dextrinase (Carlsberg group)
# =========================
"HORVU.BARKE.PROJ.7HG00595270.1" = "Lda",

# =========================
# Isoamylase (Carlsberg group)
# =========================
# Real gene symbols
"HORVU.BARKE.PROJ.7HG00618070.1" = "Iso1",
"HORVU.BARKE.PROJ.1HG00057560.1" = "Iso2",
"HORVU.BARKE.PROJ.5HG00443610.1" = "Iso3",

# No real gene symbol
"HORVU.BARKE.PROJ.2HG00154890.1" = "chr2H15489",
"HORVU.BARKE.PROJ.2HG00149340.1" = "chr2H14934",
"HORVU.BARKE.PROJ.7HG00681480.1" = "chr7H68148",
"HORVU.BARKE.PROJ.7HG00822990.1" = "chr7H82299",
"HORVU.BARKE.PROJ.7HG00823000.1" = "chr7H82300",
"HORVU.BARKE.PROJ.7HG00645630.1" = "chr7H64563",

# =========================
# Alpha-glucosidase (Carlsberg group)
# =========================
# Real gene symbols
"HORVU.BARKE.PROJ.7HG00666400.1" = "Agl1",
"HORVU.BARKE.PROJ.3HG00202550.1" = "Agl2",
"HORVU.BARKE.PROJ.1HG00052860.1" = "Agl3",
"HORVU.BARKE.PROJ.4HG00360800.1" = "Agl4",

# No real gene symbol
"HORVU.BARKE.PROJ.2HG00192840.1" = "chr2H19284",
"HORVU.BARKE.PROJ.2HG00192870.1" = "chr2H19287",
"HORVU.BARKE.PROJ.2HG00192880.1" = "chr2H19288",
"HORVU.BARKE.PROJ.3HG00234540.1" = "chr3H23454",
"HORVU.BARKE.PROJ.6HG00561650.1" = "chr6H56165",
"HORVU.BARKE.PROJ.7HG00595360.1" = "chr7H59536",

# =========================
# Alpha glucan phosphorylase (Carlsberg group)
# =========================
# Real gene symbol
"HORVU.BARKE.PROJ.5HG00465050.1" = "Phs1",

# No real gene symbol
"HORVU.BARKE.PROJ.3HG00275600.1" = "chr3H27560",
"HORVU.BARKE.PROJ.2HG00708120.1" = "chr2H70812",
"HORVU.BARKE.PROJ.3HG00275620.1" = "chr3H27562",

# =========================
# Lichenases (uniprot)
# =========================
"HORVU.BARKE.PROJ.1HG00053050.1" = "Glb1",
"HORVU.BARKE.PROJ.7HG00679910.1" = "Glb2",
# =========================
# Endo-beta-1,3-glucanases (barkepgv2)
# =========================

# ---- chr1 ----
"HORVU.BARKE.PROJ.1HG00010560.1" = "chr1H1056",
"HORVU.BARKE.PROJ.1HG00010610.1" = "chr1H1061",
"HORVU.BARKE.PROJ.1HG00062010.1" = "chr1H6201",
"HORVU.BARKE.PROJ.1HG00067490.1" = "chr1H6749",
"HORVU.BARKE.PROJ.1HG00084470.1" = "chr1H8447",
"HORVU.BARKE.PROJ.1HG00084660.1" = "chr1H8466",
"HORVU.BARKE.PROJ.1HG00690880.1" = "chr1H69088",

# ---- chr2 ----
"HORVU.BARKE.PROJ.2HG00150170.1" = "chr2H15017",

# ---- chr3 ----
"HORVU.BARKE.PROJ.3HG00259580.1" = "chr3H25958",
"HORVU.BARKE.PROJ.3HG00269220.1" = "chr3H26922",
"HORVU.BARKE.PROJ.3HG00274530.1" = "chr3H27453",
"HORVU.BARKE.PROJ.3HG00291000.1" = "chr3H29100",
"HORVU.BARKE.PROJ.3HG00291190.1" = "chr3H29119",
"HORVU.BARKE.PROJ.3HG00291270.1" = "chr3H29127",
"HORVU.BARKE.PROJ.3HG00291300.1" = "chr3H29130",
"HORVU.BARKE.PROJ.3HG00291320.1" = "chr3H29132",
"HORVU.BARKE.PROJ.3HG00291730.1" = "chr3H29173",

# ---- chr4 ----
"HORVU.BARKE.PROJ.4HG00304900.1" = "chr4H30490",
"HORVU.BARKE.PROJ.4HG00310050.1" = "chr4H31005",
"HORVU.BARKE.PROJ.4HG00358890.1" = "chr4H35889",
"HORVU.BARKE.PROJ.4HG00360390.1" = "chr4H36039",

# ---- chr5 ----
"HORVU.BARKE.PROJ.5HG00442540.1" = "chr5H44254",
"HORVU.BARKE.PROJ.5HG00448060.1" = "chr5H44806",
"HORVU.BARKE.PROJ.5HG00453330.1" = "chr5H45333",
"HORVU.BARKE.PROJ.5HG00464130.1" = "chr5H46413",
"HORVU.BARKE.PROJ.5HG00470160.1" = "chr5H47016",

# ---- chr6 ----
"HORVU.BARKE.PROJ.6HG00496480.1" = "chr6H49648",

# ---- chr7 ----
"HORVU.BARKE.PROJ.7HG00642370.1" = "chr7H64237",
"HORVU.BARKE.PROJ.7HG00653120.1" = "chr7H65312",

# =========================
# Cellulase (barkepgv2)
# =========================

# ---- chr1 ----
"HORVU.BARKE.PROJ.1HG00014980.1" = "chr1H1498",
"HORVU.BARKE.PROJ.1HG00018950.1" = "chr1H1895",
"HORVU.BARKE.PROJ.1HG00073020.1" = "chr1H7302",
"HORVU.BARKE.PROJ.1HG00078380.1" = "chr1H7838",

# ---- chr2 ----
"HORVU.BARKE.PROJ.2HG00154880.1" = "chr2H15488",
"HORVU.BARKE.PROJ.2HG00160110.1" = "chr2H16011",
"HORVU.BARKE.PROJ.2HG00162350.1" = "chr2H16235",
"HORVU.BARKE.PROJ.2HG00192140.1" = "chr2H19214",

# ---- chr3 ----
"HORVU.BARKE.PROJ.3HG00226090.1" = "chr3H22609",
"HORVU.BARKE.PROJ.3HG00727900.1" = "chr3H72790",

# ---- chr4 ----
"HORVU.BARKE.PROJ.4HG00311880.1" = "chr4H31188",
"HORVU.BARKE.PROJ.4HG00345900.1" = "chr4H34590",
"HORVU.BARKE.PROJ.4HG00379800.1" = "chr4H37980",

# ---- chr5 ----
"HORVU.BARKE.PROJ.5HG00427900.1" = "chr5H42790",
"HORVU.BARKE.PROJ.5HG00434990.1" = "chr5H43499",
"HORVU.BARKE.PROJ.5HG00452800.1" = "chr5H45280",
"HORVU.BARKE.PROJ.5HG00453760.1" = "chr5H45376",

# ---- chr6 ----
"HORVU.BARKE.PROJ.6HG00502950.1" = "chr6H50295",
"HORVU.BARKE.PROJ.6HG00508740.1" = "chr6H50874",
"HORVU.BARKE.PROJ.6HG00554880.1" = "chr6H55488",
"HORVU.BARKE.PROJ.6HG00558010.1" = "chr6H55801",
"HORVU.BARKE.PROJ.6HG00561940.1" = "chr6H56194",

# ---- chr7 ----
"HORVU.BARKE.PROJ.7HG00596870.1" = "chr7H59687",
"HORVU.BARKE.PROJ.7HG00596880.1" = "chr7H59688",
"HORVU.BARKE.PROJ.7HG00606250.1" = "chr7H60625",
"HORVU.BARKE.PROJ.7HG00607220.1" = "chr7H60722",
"HORVU.BARKE.PROJ.7HG00638840.1" = "chr7H63884",
"HORVU.BARKE.PROJ.7HG00653770.1" = "chr7H65377",
"HORVU.BARKE.PROJ.7HG00805710.1" = "chr7H80571",

# =========================
# Exo-β-hydrolases (uniprot)
# =========================

# ---- chr5 ----
"HORVU.BARKE.PROJ.5HG00463840.1" = "Exo2",
"HORVU.BARKE.PROJ.5HG00463920.1" = "Exo1",

# =========================
# Beta-glucosidase (barkepgv2)
# =========================

# ---- chr1 ----
"HORVU.BARKE.PROJ.1HG00054020.1" = "Bgq60",

# ---- chr2 ----
"HORVU.BARKE.PROJ.2HG00107620.1" = "chr2H10762",
"HORVU.BARKE.PROJ.2HG00159250.1" = "chr2H15925",
"HORVU.BARKE.PROJ.2HG00164980.1" = "chr2H16498",
"HORVU.BARKE.PROJ.2HG00165000.1" = "chr2H16500",
"HORVU.BARKE.PROJ.2HG00165020.1" = "chr2H16502",
"HORVU.BARKE.PROJ.2HG00187930.1" = "chr2H18793",

# ---- chr3 ----
"HORVU.BARKE.PROJ.3HG00199770.1" = "chr3H19977",
"HORVU.BARKE.PROJ.3HG00232460.1" = "chr3H23246",
"HORVU.BARKE.PROJ.3HG00270230.1" = "chr3H27023",
"HORVU.BARKE.PROJ.3HG00270250.1" = "chr3H27025",
"HORVU.BARKE.PROJ.3HG00270270.1" = "chr3H27027",
"HORVU.BARKE.PROJ.3HG00279420.1" = "chr3H27942",
"HORVU.BARKE.PROJ.3HG00285690.1" = "chr3H28569",
"HORVU.BARKE.PROJ.3HG00285720.1" = "chr3H28572",
"HORVU.BARKE.PROJ.3HG00285810.1" = "chr3H28581",

# ---- chr4 ----
"HORVU.BARKE.PROJ.4HG00308030.1" = "chr4H30803",
"HORVU.BARKE.PROJ.4HG00308050.1" = "chr4H30805",
"HORVU.BARKE.PROJ.4HG00328690.1" = "chr4H32869",
"HORVU.BARKE.PROJ.4HG00361530.1" = "chr4H36153",

# ---- chr5 ----
"HORVU.BARKE.PROJ.5HG00446020.1" = "chr5H44602",
"HORVU.BARKE.PROJ.5HG00449830.1" = "chr5H44983",
"HORVU.BARKE.PROJ.5HG00449840.1" = "chr5H44984",
"HORVU.BARKE.PROJ.5HG00449910.1" = "chr5H44991",
"HORVU.BARKE.PROJ.5HG00449950.1" = "chr5H44995",

# ---- chr7 ----
"HORVU.BARKE.PROJ.7HG00612580.1" = "chr7H61258",
"HORVU.BARKE.PROJ.7HG00616950.1" = "chr7H61695",
"HORVU.BARKE.PROJ.7HG00649790.1" = "chr7H64979",
"HORVU.BARKE.PROJ.7HG00653840.1" = "chr7H65384",
"HORVU.BARKE.PROJ.7HG00663760.1" = "chr7H66376",


# =========================
# Endo-xylanase (Betts et al., 2020)
# =========================
"HORVU.BARKE.PROJ.5HG00471990.1" = "Xyn1",
"HORVU.BARKE.PROJ.5HG00471950.1" = "Xyn2",
"HORVU.BARKE.PROJ.5HG00779790.1" = "Xyn3",
"HORVU.BARKE.PROJ.3HG00205980.1" = "Xyn4",
"HORVU.BARKE.PROJ.3HG00205970.1" = "Xyn5",
"HORVU.BARKE.PROJ.2HG00105770.1" = "Xyn6",
"HORVU.BARKE.PROJ.4HG00305690.1" = "Xyn7",
"HORVU.BARKE.PROJ.1HG00022700.1" = "Xyn8",
"HORVU.BARKE.PROJ.2HG00186320.1" = "Xyn9",
"HORVU.BARKE.PROJ.2HG00194590.1" = "Xyn10",
"HORVU.BARKE.PROJ.5HG00431450.1" = "Xyn11",
"HORVU.BARKE.PROJ.4HG00363430.1" = "Xyn12",
"HORVU.BARKE.PROJ.4HG00363480.1" = "Xyn13",
"HORVU.BARKE.PROJ.2HG00097040.1" = "Xyn14",
"HORVU.BARKE.PROJ.4HG00357530.1" = "Xyn15",

# =========================
# Beta-xylosidase (Betts et al., 2020)
# =========================
"HORVU.BARKE.PROJ.6HG00556300.1" = "Xyl1",
"HORVU.BARKE.PROJ.4HG00306120.1" = "Xyl2",
"HORVU.BARKE.PROJ.4HG00362200.1" = "Xyl3",
"HORVU.BARKE.PROJ.4HG00312970.1" = "Xyl4",
"HORVU.BARKE.PROJ.4HG00312980.1" = "Xyl4",
"HORVU.BARKE.PROJ.2HG00167010.1" = "Xyl5",
"HORVU.BARKE.PROJ.4HG00328350.1" = "Xyl6",
"HORVU.BARKE.PROJ.6HG00543870.1" = "Xyl7",
"HORVU.BARKE.PROJ.2HG00188490.1" = "Ara1",

# =========================
# Arabinofuranosidases (uniprot)
# =========================
"HORVU.BARKE.PROJ.5HG00420480.1" = "Axah1",
"HORVU.BARKE.PROJ.4HG00326450.1" = "Axah2",
"HORVU.BARKE.PROJ.4HG00325990.1" = "Axah3",
"HORVU.BARKE.PROJ.4HG00346830.1" = "Axah4",
"HORVU.BARKE.PROJ.2HG00101670.1" = "Axah5",

# =========================
# Mannanases (Iglesias-Fernández et al., 2020)
# =========================
"HORVU.BARKE.PROJ.3HG00253120.1" = "Man1",
"HORVU.BARKE.PROJ.3HG00263090.1" = "Man2",
"HORVU.BARKE.PROJ.5HG00479880.1" = "Man3",
"HORVU.BARKE.PROJ.6HG00561220.1" = "Man4",
"HORVU.BARKE.PROJ.5HG00771660.1" = "Man5",
"HORVU.BARKE.PROJ.7HG00612160.1" = "Man6",

# =========================
#  Alpha-galactosidase (uniprot)
# =========================

"HORVU.BARKE.PROJ.1HG00041210.1" = "chr1H4121",
"HORVU.BARKE.PROJ.1HG00041240.1" = "chr1H4124",
"HORVU.BARKE.PROJ.2HG00102930.1" = "chr2H10293",
"HORVU.BARKE.PROJ.5HG00772740.1" = "chr5H77274",
"HORVU.BARKE.PROJ.5HG00772750.1" = "chr5H77275",
"HORVU.BARKE.PROJ.6HG00537080.1" = "chr6H53708",
"HORVU.BARKE.PROJ.6HG00556460.1" = "chr6H55646",

# =========================
#Polygalacturonase (uniprot)
# =========================

"HORVU.BARKE.PROJ.4HG00360780.1" = "chr4H36078",
"HORVU.BARKE.PROJ.5HG00475020.1" = "chr5H47502",
"HORVU.BARKE.PROJ.5HG00475050.1" = "chr5H47505",
"HORVU.BARKE.PROJ.6HG00502310.1" = "chr6H50231",
"HORVU.BARKE.PROJ.7HG00599930.1" = "chr7H59993",

#  Subtilases (serine endopeptidases) (barkepgv2) 

"HORVU.BARKE.PROJ.1HG00053700.1" = "chr1H5370",
"HORVU.BARKE.PROJ.2HG00101890.1" = "chr2H10189",
"HORVU.BARKE.PROJ.2HG00168300.1" = "chr2H16830",
"HORVU.BARKE.PROJ.4HG00357630.1" = "chr4H35763",
"HORVU.BARKE.PROJ.6HG00493740.1" = "chr6H49374",
"HORVU.BARKE.PROJ.7HG00646960.1" = "chr7H64696",


# ---- Papain-like cysteine protease (uniprot)
"HORVU.BARKE.PROJ.4HG00313340.1" = "Epa",
"HORVU.BARKE.PROJ.3HG00281420.1" = "Epb1",
"HORVU.BARKE.PROJ.3HG00281440.1" = "Epb2",
"HORVU.BARKE.PROJ.3HG00281450.1" = "Epb2",
"HORVU.BARKE.PROJ.3HG00281460.1" = "Epb2",
"HORVU.BARKE.PROJ.4HG00311260.1" = "CatB1",
"HORVU.BARKE.PROJ.6HG00495660.1" = "CatB2",
"HORVU.BARKE.PROJ.6HG00495670.1" = "chr6H49567",
"HORVU.BARKE.PROJ.5HG00436460.1" = "Ale",

# =========================
# Leucine Aminopeptidase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.1HG00021950.1" = "chr1H2195",
"HORVU.BARKE.PROJ.6HG00565260.1" = "chr6H56526",
"HORVU.BARKE.PROJ.7HG00805360.1" = "chr7H80536",

# =========================
# Dipeptidyl peptidase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.2HG00100820.1" = "chr2H10082",
"HORVU.BARKE.PROJ.4HG00348310.1" = "chr4H34831",
"HORVU.BARKE.PROJ.5HG00406500.1" = "chr5H40650",

# =========================
# Tripeptidyl peptidase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.6HG00545020.1" = "chr6H54502",
"HORVU.BARKE.PROJ.6HG00545030.1" = "chr6H54503",

# =========================
# Serine carboxypeptidase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.3HG00285590.1" = "Cbp1",
"HORVU.BARKE.PROJ.1HG00018290.1" = "Cbp2",
"HORVU.BARKE.PROJ.6HG00494930.1" = "Cbp3",
"HORVU.BARKE.PROJ.3HG00222540.1" = "Cpb2-1",
"HORVU.BARKE.PROJ.3HG00222540.2" = "Cpb2-1",
"HORVU.BARKE.PROJ.7HG00597910.1" = "Cpb2-2",
"HORVU.BARKE.PROJ.2HG00107490.1" = "Cpb2-3",

"HORVU.BARKE.PROJ.1HG00044340.1" = "chr1H4434",
"HORVU.BARKE.PROJ.1HG00044350.1" = "chr1H4435",
"HORVU.BARKE.PROJ.2HG00093780.1" = "chr2H9378",
"HORVU.BARKE.PROJ.2HG00171840.1" = "chr2H17184",
"HORVU.BARKE.PROJ.2HG00171910.1" = "chr2H17191",
"HORVU.BARKE.PROJ.2HG00171920.1" = "chr2H17192",
"HORVU.BARKE.PROJ.2HG00722350.1" = "chr2H72235",
"HORVU.BARKE.PROJ.3HG00215140.1" = "chr3H21514",
"HORVU.BARKE.PROJ.3HG00260410.1" = "chr3H26041",
"HORVU.BARKE.PROJ.3HG00265950.1" = "chr3H26595",
"HORVU.BARKE.PROJ.3HG00269040.1" = "chr3H26904",
"HORVU.BARKE.PROJ.4HG00311090.1" = "chr4H31109",
"HORVU.BARKE.PROJ.4HG00311110.1" = "chr4H31111",
"HORVU.BARKE.PROJ.4HG00315620.1" = "chr4H31562",
"HORVU.BARKE.PROJ.4HG00321770.1" = "chr4H32177",
"HORVU.BARKE.PROJ.4HG00375970.1" = "chr4H37597",
"HORVU.BARKE.PROJ.5HG00445740.1" = "chr5H44574",
"HORVU.BARKE.PROJ.5HG00445760.1" = "chr5H44576",
"HORVU.BARKE.PROJ.5HG00451630.1" = "chr5H45163",
"HORVU.BARKE.PROJ.5HG00451680.1" = "chr5H45168",
"HORVU.BARKE.PROJ.7HG00593540.1" = "chr7H59354",

# =========================
# Serpins (uniprot)
# =========================
"HORVU.BARKE.PROJ.4HG00314040.1" = "Paz1",
"HORVU.BARKE.PROJ.4HG00314080.1" = "PazX",
"HORVU.BARKE.PROJ.5HG00476910.1" = "Bsz7",

# =========================
# Lipoxygenase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.4HG00307600.1" = "Lox1",
"HORVU.BARKE.PROJ.5HG00383870.1" = "Lox2-1",
"HORVU.BARKE.PROJ.7HG00579310.1" = "Lox2-1",
"HORVU.BARKE.PROJ.3HG00735580.1" = "Lox2-2",
"HORVU.BARKE.PROJ.4HG00760420.1" = "Lox2-2",
"HORVU.BARKE.PROJ.6HG00490920.1" = "Lox2-2",
"HORVU.BARKE.PROJ.6HG00519710.1" = "Lox2-3",

"HORVU.BARKE.PROJ.1HG00050580.1" = "chr1H5058",
"HORVU.BARKE.PROJ.2HG00156010.1" = "chr2H15601",
"HORVU.BARKE.PROJ.2HG00188750.1" = "chr2H18875",
"HORVU.BARKE.PROJ.2HG00188780.1" = "chr2H18878",
"HORVU.BARKE.PROJ.4HG00307640.1" = "chr4H30764",
"HORVU.BARKE.PROJ.4HG00369200.1" = "chr4H36920",
"HORVU.BARKE.PROJ.4HG00369210.1" = "chr4H36921",
"HORVU.BARKE.PROJ.5HG00462390.1" = "chr5H46239",
"HORVU.BARKE.PROJ.5HG00462410.1" = "chr5H46241",
"HORVU.BARKE.PROJ.5HG00462420.1" = "chr5H46242",
"HORVU.BARKE.PROJ.5HG00462430.1" = "chr5H46243",
"HORVU.BARKE.PROJ.5HG00462440.1" = "chr5H46244",
"HORVU.BARKE.PROJ.5HG00481340.1" = "chr5H48134",
"HORVU.BARKE.PROJ.7HG00616920.1" = "chr7H61692",

# =========================
# Allene oxide synthase (barkepgv2)
# =========================
"HORVU.BARKE.PROJ.4HG00359650.1" = "Aos1",
"HORVU.BARKE.PROJ.4HG00359600.1" = "Aos2",
"HORVU.BARKE.PROJ.5HG00466330.1" = "chr5H46633",
"HORVU.BARKE.PROJ.6HG00493590.1" = "chr6H49359",
"HORVU.BARKE.PROJ.6HG00524700.1" = "chr6H52470"
)


# --- 1) Define the family-to-category mapping (in desired order) ---
# ================================================================
# Family → Category mapping (ALIGNED WITH all_families_ID)
# ================================================================

family_category <- c(

  # ======================
  # Amylolysis
  # ======================
  "Alpha-amylase"              = "Amylolysis",
  "Beta-amylase"               = "Amylolysis",
  "Limit dextrinase"           = "Amylolysis",
  "Isoamylase"                 = "Amylolysis",
  "Alpha-glucosidase"          = "Amylolysis",
  "Alpha glucan phosphorylase" = "Amylolysis",

  # ======================
  # Cytolysis (cell wall)
  # ======================
  "Cellulase"              = "Cytolysis",
  "Endo-beta-1,3-glucanase"= "Cytolysis",
  "Beta-glucan exohydrolase" = "Cytolysis",
  "Lichenase"              = "Cytolysis",
  "Xylanase"               = "Cytolysis",
  "Beta-xylosidase"        = "Cytolysis",
  "Arabinofuranosidase"    = "Cytolysis",
  "Mannanase"              = "Cytolysis",
  "Alpha-galactosidase"    = "Cytolysis",
  "Polygalacturonase"      = "Cytolysis",

  # ======================
  # Proteolysis
  # ======================
  "Subtilase"                         = "Proteolysis",
  "Papain-like cysteine protease" = "Proteolysis",
  "Leucine aminopeptidase"          = "Proteolysis",
  "Dipeptidyl-peptidase"            = "Proteolysis",
  "Tripeptidyl-peptidase"           = "Proteolysis",
  "Serine carboxypeptidase"         = "Proteolysis",
  "Serpin"                          = "Proteolysis",

  # ======================
  # Lipid metabolism
  # ======================
  "Lipoxygenase"          = "Lipolysis",
  "Allene oxide synthase" = "Lipolysis"
)
all_families_ID <- readRDS("Z:/PhD Munchen/Guido personal PhD Project_1/Barley germination proteome/Resources for Data analysis/R/R_Data/shared_malting_maps/all_families_ID.rds")

# remove families not present in the category map
all_families_ID <- all_families_ID[names(all_families_ID) %in% names(family_category)]

# --- 4) Collect proteins for each family ---
malting_rows <- lapply(names(all_families_ID), function(fam) {
  ids <- all_families_ID[[fam]]
  sub <- final_Hv5_mat %>% filter(Protein.IDs_pG %in% ids)
  if (nrow(sub) > 0) {
    sub$enzyme_family <- fam
    sub$Category <- family_category[[fam]]
    return(sub)
  }
  return(NULL)
})

malting_all <- bind_rows(malting_rows)

if (nrow(malting_all) == 0) stop("❌ No proteins matched the ID list!")

# ------------------------------------------------
# Map BARKE IDs → gene names (DISPLAY ONLY)
# ------------------------------------------------
malting_all$gene_label <- barke_gene_map[malting_all$Protein.IDs_pG]

# Fallback: if not mapped, keep BARKE ID
malting_all$gene_label[is.na(malting_all$gene_label)] <-
  malting_all$Protein.IDs_pG[is.na(malting_all$gene_label)]

gene_order_ref <- names(barke_gene_map)

malting_all$gene_order <- match(
  malting_all$Protein.IDs_pG,
  gene_order_ref
)

malting_all$gene_order[is.na(malting_all$gene_order)] <- Inf

# -----------------------------
# DEFINE ORDERING BEFORE MATRIX
# -----------------------------
family_names <- names(family_category)

category_names <- c(
  "Amylolysis",
  "Cytolysis",
  "Lipolysis",
  "Proteolysis"
)

# ----------------------------------------
# APPLY ORDERING BEFORE BUILDING MATRIX
# ----------------------------------------
malting_all$Category <- factor(
  malting_all$Category,
  levels = category_names
)

malting_all$enzyme_family <- factor(
  malting_all$enzyme_family,
  levels = family_names
)

malting_all <- arrange(
  malting_all,
  Category,
  enzyme_family,
  gene_order
)

# --- 5) Build numeric matrix ---
malting_mat <- as.matrix(malting_all[, ..int_cols])
mode(malting_mat) <- "numeric"
malting_mat[!is.finite(malting_mat)] <- NA

# --- 6) Row labels and annotation ---
rownames(malting_mat) <- make.unique(malting_all$gene_label)

annotation_row <- data.frame(
  Family   = malting_all$enzyme_family,
  Category = malting_all$Category,
  row.names = rownames(malting_mat)
)

# --- 9) Colors (unchanged) ---
heat_colors <- colorRampPalette(c("#FFFFFF","#FEE8C8","#D73000"))(150)

palette <- colorRampPalette(brewer.pal(12, "Set3"))(length(family_names))
family_colors <- setNames(palette, family_names)

category_colors <- c(
  "Amylolysis"  = "#E6C229",
  "Cytolysis"   = "#7FB685",
  "Proteolysis" = "#E07A5F",
  "Lipolysis"   = "#8DA0CB"
)

ann_colors <- list(Family = family_colors, Category = category_colors)

# ------------------------------------------------
# Adjust column names to time points (repeated)
# ------------------------------------------------
time_labels <- c("0 h","12 h","24 h","36 h","48 h","60 h","72 h","84 h","96 h")
colnames(malting_mat) <- rep(time_labels, length.out = ncol(malting_mat))

# ------------------------------------------------
# ComplexHeatmap setup
# ------------------------------------------------
library(ComplexHeatmap)
library(circlize)
library(grid)
 
# ================================================================
# Heatmaps for correlation with enzymatic data 
# ================================================================

library(grid)
library(gridExtra)

panel_families <- c(
    "Alpha-amylase",
    "Beta-amylase",
    "Lichenase",
    "Limit dextrinase"
)

# -----------------------------
# Subset data
# -----------------------------
subset_multi <- malting_all %>%
    dplyr::filter(enzyme_family %in% panel_families)

if (nrow(subset_multi) == 0) {
    stop("❌ No proteins found for selected families.")
}

# -----------------------------
# Build raw matrix
# -----------------------------
subset_mat_raw <- as.matrix(subset_multi[, ..int_cols])
mode(subset_mat_raw) <- "numeric"
subset_mat_raw[!is.finite(subset_mat_raw)] <- NA
subset_multi$gene_label <- barke_gene_map[subset_multi$Protein.IDs_pG]
subset_multi$gene_label[is.na(subset_multi$gene_label)] <-
  subset_multi$Protein.IDs_pG[is.na(subset_multi$gene_label)]

rownames(subset_mat_raw) <- make.unique(subset_multi$gene_label)

# -----------------------------
# Z-score matrix (row-wise)
# -----------------------------
subset_mat_z <- t(scale(t(subset_mat_raw)))

# ================================================================
# FIXED CELL SIZE (global)
# ================================================================
cell_height <- 12
cell_width  <- 12

# ================================================================
# Helper: build stacked heatmap figure
# ================================================================
build_panel <- function(mat, title) {
    
    grobs <- list()
    
    for (i in seq_along(panel_families)) {
        
        fam <- panel_families[i]
        fam_rows <- which(subset_multi$enzyme_family == fam)
        if (length(fam_rows) == 0) next
        
        fam_mat <- mat[fam_rows, , drop = FALSE]
        
        ph <- pheatmap::pheatmap(
            fam_mat,
            scale = "none",
            cluster_rows = FALSE,
            cluster_cols = FALSE,
            treeheight_row = 0,
            treeheight_col = 0,
            color = heat_colors,
            show_rownames = TRUE,
            show_colnames = (i == 1),   # <-- ONLY TOP PANEL
            legend = TRUE,
            fontsize_row = 6,
            fontsize_col = 6,
            cellheight = cell_height,
            cellwidth  = cell_width,
            main = fam,
            silent = TRUE
        )
        
        # Remove any dendrogram remnants
        ph$gtable$grobs <- ph$gtable$grobs[
            !grepl("dendrogram", ph$gtable$layout$name)
        ]
        
        grobs[[fam]] <- ph$gtable
    }
    
    # Make panels closer together
    arrangeGrob(
        grobs = grobs,
        ncol = 1,
        heights = rep(1, length(grobs)),  # <-- compact vertical spacing
        top = textGrob(title, gp = gpar(fontsize = 14, fontface = "bold"))
    )
}

# ================================================================
# BUILD FIGURES
# ================================================================
panel_z   <- build_panel(subset_mat_z,   "Hv5 enzyme panels (z-scored log2 iBAQ)")
panel_raw <- build_panel(subset_mat_raw, "Hv5 enzyme panels (raw log2 iBAQ)")
out_dir <- "Z:/PhD Munchen/Guido personal PhD Project_1/Barley germination proteome/Resources for Data analysis/R/R_export/Hv5/malting_enzymes"

# ================================================================
# EXPORT (NO SVG)
# ================================================================
for (ext in c("pdf", "jpeg")) {
    
    # Z-SCORED
    out_z <- file.path(out_dir, paste0("Hv5_EnzymePanels_Zscore_fixedCell.", ext))
    if (ext == "pdf") {
        pdf(out_z, width = 10, height = 8)
    } else {
        jpeg(out_z, width = 10, height = 8, units = "in", res = 350)
    }
    grid.draw(panel_z)
    dev.off()
    
    # RAW
    out_r <- file.path(out_dir, paste0("Hv5_EnzymePanels_Raw_fixedCell.", ext))
    if (ext == "pdf") {
        pdf(out_r, width = 10, height = 8)
    } else {
        jpeg(out_r, width = 10, height = 8, units = "in", res = 350)
    }
    grid.draw(panel_raw)
    dev.off()
}
# ================================================================
# Bmy1 ONLY — 3 HEATMAPS (LOG2 10–30, RAW, Z-SCORE)
# Timepoints stacked (rows) × Conditions (columns)
# ================================================================

target_id <- "HORVU.BARKE.PROJ.4HG00380960.1"  # Bmy1

bmy1_df <- malting_all %>%
  dplyr::filter(Protein.IDs_pG == target_id)

if (nrow(bmy1_df) == 0) {
  stop("❌ Bmy1 not found.")
}

if (nrow(bmy1_df) > 1) {
  message("⚠️ Multiple entries found for Bmy1 — using first")
  bmy1_df <- bmy1_df[1, ]
}

# -----------------------------
# Extract numeric values
# -----------------------------
bmy1_vec <- as.numeric(bmy1_df[, ..int_cols])
names(bmy1_vec) <- colnames(bmy1_df[, ..int_cols])

# -----------------------------
# Parse condition + time
# -----------------------------
bmy1_long <- data.frame(
  value = bmy1_vec,
  colname = names(bmy1_vec)
) %>%
  mutate(
    condition = sub("_.*", "", colname),
    time      = sub(".*_", "", colname)
  ) %>%
  filter(!is.na(condition), !is.na(time)) %>%
  group_by(condition, time) %>%
  summarise(value = mean(value), .groups = "drop")

# -----------------------------
# Build matrix (time × condition)
# -----------------------------
time_levels <- c("0hs","12hs","24hs","36hs","48hs","60hs","72hs","84hs","96hs")

bmy1_mat <- bmy1_long %>%
  pivot_wider(names_from = condition, values_from = value)

# enforce time order
bmy1_mat <- bmy1_mat %>%
  mutate(time = factor(time, levels = time_levels)) %>%
  arrange(time)

rownames_vec <- bmy1_mat$time

bmy1_mat <- bmy1_mat %>%
  select(-time) %>%
  as.matrix()

rownames(bmy1_mat) <- rownames_vec
mode(bmy1_mat) <- "numeric"

bmy1_mat[!is.finite(bmy1_mat)] <- NA

# -----------------------------
# LOG2
# -----------------------------
bmy1_log2 <- log2(bmy1_mat)

# -----------------------------
# Z-SCORE (row-wise = per timepoint)
# -----------------------------
bmy1_z <- t(scale(t(bmy1_log2)))

# -----------------------------
# RAW (already linear iBAQ)
# -----------------------------
bmy1_raw <- bmy1_mat

# -----------------------------
# LOG2 CLIPPED (10–30)
# -----------------------------
bmy1_log2_clip <- bmy1_log2
bmy1_log2_clip[bmy1_log2_clip < 10] <- 10
bmy1_log2_clip[bmy1_log2_clip > 30] <- 30

log2_breaks <- seq(10, 30, length.out = 200)

# -----------------------------
# COLORS
# -----------------------------
heat_cols <- colorRampPalette(c("#FFFFFF","#FEE8C8","#D73000"))(200)

# -----------------------------
# OUTPUT DIR
# -----------------------------
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# ================================================================
# LOG2 HEATMAP (10–30)
# ================================================================
pdf(file.path(out_dir, "Bmy1_log2_10to30_timeRows.pdf"), width = 4, height = 6)

pheatmap::pheatmap(
  bmy1_log2_clip,
  color = heat_cols,
  breaks = log2_breaks,
  scale = "none",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  treeheight_row = 0,
  treeheight_col = 0,
  show_rownames = TRUE,
  show_colnames = TRUE,
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Bmy1 (log2 iBAQ, 10–30)"
)

dev.off()

# ================================================================
# RAW HEATMAP
# ================================================================
pdf(file.path(out_dir, "Bmy1_raw_timeRows.pdf"), width = 4, height = 6)

pheatmap::pheatmap(
  bmy1_raw,
  color = heat_cols,
  scale = "none",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  treeheight_row = 0,
  treeheight_col = 0,
  show_rownames = TRUE,
  show_colnames = TRUE,
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Bmy1 (raw iBAQ)"
)

dev.off()

# ================================================================
# Z-SCORE HEATMAP
# ================================================================
pdf(file.path(out_dir, "Bmy1_zscore_timeRows.pdf"), width = 4, height = 6)

pheatmap::pheatmap(
  bmy1_z,
  color = heat_cols,
  scale = "none",
  cluster_rows = FALSE,
  cluster_cols = FALSE,
  treeheight_row = 0,
  treeheight_col = 0,
  show_rownames = TRUE,
  show_colnames = TRUE,
  fontsize_row = 8,
  fontsize_col = 8,
  main = "Bmy1 (z-score per timepoint)"
)

dev.off()

message("✅ Bmy1 heatmaps (log2 clipped, raw, z-score) generated.")
```


##Unbiased cluster analysis to see if there are group of proteins that are systematically more abundnat in more modern varieties  

```{r}
# ============================================================
# Identify proteins uniquely downregulated in Heat, Drought, or Combo vs Std
# (≥7 of 9 timepoints) and perform GO enrichment (Hv5 dataset)
# ============================================================

library(data.table)
library(dplyr)
library(tidyr)
library(clusterProfiler)
library(org.At.tair.db)
library(stringr)
library(readr)
library(VennDiagram)
library(grid)


# ------------------------------------------------------------
# Step 1: Load dataset and clean unwanted columns
# ------------------------------------------------------------
out_path <- "../../../R_Data/Hv5/pG_iBAQ_postnorm_bridgeOnly/Hv5_final.csv"
final_Hv5_mat <- fread(out_path)

# Identify intensity columns
int_cols <- grep("^iBAQ_", names(final_Hv5_mat), value = TRUE)

# Columns to drop completely
drop_cols <- c(
  "iBAQ_std_0hs_std", "iBAQ_std_0hs_r2_std", "iBAQ_std_0hs_h",
  "iBAQ_heat_0hs_r2_h", "iBAQ_std_0hs_d", "iBAQ_drought_0hs_r2_d",
  "iBAQ_std_0hs_c", "iBAQ_combo_0hs_r2_c"
)

# Remove them from the dataset directly
final_Hv5_mat <- final_Hv5_mat[, !colnames(final_Hv5_mat) %in% drop_cols, with = FALSE]

# ------------------------------------------------------------
# Step 2: Convert to long format
# ------------------------------------------------------------
ibaq_long <- final_Hv5_mat %>%
  dplyr::select(Protein.IDs_pG, TAIR, all_of(grep("^iBAQ_", names(final_Hv5_mat), value = TRUE))) %>%
  pivot_longer(cols = starts_with("iBAQ_"),
               names_to = "condition",
               values_to = "iBAQ") %>%
  mutate(
    group = case_when(
      grepl("_std_", condition)     ~ "Std",
      grepl("_heat_", condition)    ~ "Heat",
      grepl("_drought_", condition) ~ "Drought",
      grepl("_combo_", condition)   ~ "Combo",
      TRUE ~ "Other"
    ),
    timepoint_h = str_extract(condition, "[0-9]+(?=hs)") %>% as.numeric()
  ) %>%
  filter(!is.na(timepoint_h))

# ------------------------------------------------------------
# Step 3: Average iBAQ per protein, group, timepoint
# ------------------------------------------------------------
group_avg <- ibaq_long %>%
  group_by(Protein.IDs_pG, TAIR, group, timepoint_h) %>%
  summarise(iBAQ = mean(iBAQ, na.rm = TRUE), .groups = "drop")

# ------------------------------------------------------------
# Step 4: Compare each stress vs control at each time point
# ------------------------------------------------------------
comparison_df <- group_avg %>%
  pivot_wider(names_from = group, values_from = iBAQ) %>%
  mutate(
    log2FC_heat    = Heat    - Std,
    log2FC_drought = Drought - Std,
    log2FC_combo   = Combo   - Std
  )

# ------------------------------------------------------------
# Step 5: Identify proteins downregulated in ≥7 of 9 timepoints
# ------------------------------------------------------------
downregulated <- comparison_df %>%
  mutate(
    down_heat    = log2FC_heat    <= -1,
    down_drought = log2FC_drought <= -1,
    down_combo   = log2FC_combo   <= -1
  ) %>%
  group_by(Protein.IDs_pG, TAIR) %>%
  summarise(
    down_heat_count    = sum(down_heat, na.rm = TRUE),
    down_drought_count = sum(down_drought, na.rm = TRUE),
    down_combo_count   = sum(down_combo, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  mutate(
    is_down_heat    = down_heat_count    >= 7,
    is_down_drought = down_drought_count >= 7,
    is_down_combo   = down_combo_count   >= 7
  )

# ------------------------------------------------------------
# Step 6: Define unique and shared sets
# ------------------------------------------------------------
heat_only <- downregulated %>%
  filter(is_down_heat & !is_down_drought & !is_down_combo)

drought_only <- downregulated %>%
  filter(is_down_drought & !is_down_heat & !is_down_combo)

combo_only <- downregulated %>%
  filter(is_down_combo & !is_down_heat & !is_down_drought)

all_three <- downregulated %>%
  filter(is_down_heat & is_down_drought & is_down_combo)

cat("✔ Heat-only:", nrow(heat_only), "\n")
cat("✔ Drought-only:", nrow(drought_only), "\n")
cat("✔ Combo-only:", nrow(combo_only), "\n")
cat("✔ Shared in all three:", nrow(all_three), "\n")

# ------------------------------------------------------------
# Step 7: Venn diagram (scaled)
# ------------------------------------------------------------
venn_list <- list(
  Heat = downregulated$Protein.IDs_pG[downregulated$is_down_heat],
  Drought = downregulated$Protein.IDs_pG[downregulated$is_down_drought],
  Combo = downregulated$Protein.IDs_pG[downregulated$is_down_combo]
)

venn_out <- "../../../R_export/Hv5/unbiased/downregulated_venn_scaled.pdf"
pdf(venn_out, width = 8, height = 8)

assign("overrideTriple", 1, envir = .GlobalEnv)
venn.plot <- venn.diagram(
  x = venn_list,
  filename = NULL,
  fill = c("#E69F00", "#56B4E9", "#009E73"),
  alpha = 0.5,
  cex = 1.6,
  cat.cex = 1.4,
  cat.pos = c(-20, 20, 0),
  cat.dist = 0.06,
  main = "Unique and Shared Downregulated Proteins (scaled)",
  main.cex = 1.6,
  scaled = TRUE,
  euler.d = TRUE,
  lty = "blank",
  fontface = "bold",
  cat.fontface = "bold"
)
grid::grid.draw(venn.plot)
dev.off()
rm(overrideTriple, envir = .GlobalEnv)
cat("✔ Scaled (approximate) Venn diagram saved to:", venn_out, "\n")

# ------------------------------------------------------------
# Step 8: GO enrichment (unique + shared)
# ------------------------------------------------------------
perform_enrichment <- function(df, universe, name) {
  genes <- df %>%
    filter(!is.na(TAIR) & TAIR != "") %>%
    distinct(TAIR) %>%
    mutate(TAIR = sub(";.*$", "", TAIR))
  
  if (nrow(genes) == 0) {
    cat("⚠ No valid TAIR IDs for", name, "\n")
    return(NULL)
  }
  
  enrichGO(
    gene = genes$TAIR,
    OrgDb = org.At.tair.db,
    keyType = "TAIR",
    ont = "BP",
    universe = universe$TAIR,pAdjustMethod="none", minGSSize = 1, maxGSSize = 50000,   qvalueCutoff = 1)
}

# Prepare universe
universe <- final_Hv5_mat %>%
  filter(!is.na(TAIR) & TAIR != "") %>%
  distinct(TAIR) %>%
  mutate(TAIR = sub(";.*$", "", TAIR))

enrich_heat_only    <- perform_enrichment(heat_only, universe, "Heat-only")
enrich_drought_only <- perform_enrichment(drought_only, universe, "Drought-only")
enrich_combo_only   <- perform_enrichment(combo_only, universe, "Combo-only")
enrich_all_three    <- perform_enrichment(all_three, universe, "Shared-all-three")


# ===================================================
# Step 10 — Simplify GO enrichment results for all sets
# ===================================================
library(GOSemSim)
library(igraph)

simplify_enrichment <- function(enrich_obj, name, ontology = "BP", cutoff = 0.95) {
  if (is.null(enrich_obj) || nrow(enrich_obj@result) == 0) {
    cat("⚠ No enrichment results for", name, "\n")
    return(NULL)
  }

  cat("\n🔹 Simplifying GO enrichment for", name, "...\n")

  # 1️⃣ Prepare GO data
  sem_data <- godata("org.At.tair.db", ont = ontology, computeIC = TRUE)

  # 2️⃣ Extract GO term IDs
  go_terms <- enrich_obj@result$ID

  # 3️⃣ Compute pairwise semantic similarity (Lin method)
  sim_matrix <- mgoSim(go_terms, go_terms, semData = sem_data, measure = "Lin", combine = NULL)
  sim_matrix[is.na(sim_matrix)] <- 0
  diag(sim_matrix) <- 1

  # 4️⃣ Build adjacency matrix for high similarity
  adj <- sim_matrix >= cutoff
  diag(adj) <- FALSE

  # 5️⃣ Create graph and find clusters
  g <- graph_from_adjacency_matrix(adj, mode = "undirected")
  clusters <- components(g)$membership

  # 6️⃣ Select representative GO term (lowest adjusted p-value)
  selected <- sapply(unique(clusters), function(cl) {
    members <- which(clusters == cl)
    pvals <- enrich_obj@result$p.adjust[members]
    members[which.min(pvals)]
  })

  # 7️⃣ Filter enrichment object to simplified terms
  simplified <- enrich_obj@result[selected, ]
  enrich_obj@result <- simplified

  cat("✅", nrow(simplified), "non-redundant GO terms retained for", name, "\n")
  return(enrich_obj)
}

# Apply simplification to all four enrichment results
enrich_heat_only_simplified    <- simplify_enrichment(enrich_heat_only, "Heat-only")
enrich_drought_only_simplified <- simplify_enrichment(enrich_drought_only, "Drought-only")
enrich_combo_only_simplified   <- simplify_enrichment(enrich_combo_only, "Combo-only")
enrich_all_three_simplified    <- simplify_enrichment(enrich_all_three, "Shared-all-three")


##create an image from this analysis
# 1️⃣ Collect simplified enrichments
simplified_list <- list(
  "Heat-only"        = enrich_heat_only,
  "Drought-only"     = enrich_drought_only,
  "Combo-only"       = enrich_combo_only,
  "Shared-all-three" = enrich_all_three
)

# 2️⃣ Filter each enrichGO object to top 5 terms (keeping structure)
simplified_list_top5 <- lapply(simplified_list, function(x) {
  if (!is.null(x) && nrow(x@result) > 0) {
    x@result <- x@result %>% arrange(p.adjust) %>% head(5)
    return(x)
  } else {
    return(NULL)
  }
})

# 3️⃣ Merge filtered enrichments into a compareClusterResult object
merged_top5 <- merge_result(simplified_list_top5)

# 4️⃣ Generate dotplot (with -log10(p.adjust) color scale)
pdf("../../../R_export/Hv5/unbiased/GO_dotplot_top.pdf", width = 8, height = 8)
dotplot(
  merged_top5,
  title = "Top 5 Enriched GO Terms per Condition (Simplified)",
  font.size = 8,
  showCategory = 20
) + 
  scale_color_continuous(
    name = expression(-log[10](adjusted~p)),
    trans = "reverse",
    guide = guide_colorbar(order = 1)
  )
dev.off()


##Add annotations to the downregulated proteins 

ance_res <- enrich_ance@result
ance_res2 <- ance_res %>%
  separate_rows(geneID, sep = "/")

At_UP_annotated <- read_tsv("../../../R_import/Cam_At_project/Ortho_At_AtUP/UpAt_annotations.tsv")
UpAt_annotations_dist <- At_UP_annotated%>%
  distinct(TAIR, .keep_all = TRUE)
UpAt_annotations_dist <- UpAt_annotations_dist %>%
  mutate(TAIR = str_remove(TAIR, ";$"))

ance_res_anno <- merge(
  ance_res2,
  UpAt_annotations_dist, 
  by.x  = "geneID",  
  by.y  = "TAIR",
  all.x = TRUE
)

```

#Enzymatic tests on stressed samples 

```{r}

# ===============================
# GLOBAL FONT → ARIAL
# ===============================
theme_set(theme_get() + theme(text = element_text(family = "Arial")))

# ===============================
# A) GRAIN TRAITS DATA
# ===============================
grain_data <- tribble(
  ~Condition, ~TGW,   ~Protein, ~Starch,
  "Control",         56.72,   15.0,    57.2,
  "Heat",            55.79,   14.6,    58.4,
  "Drought",         45.24,   21.0,    42.0,
  "Combined",        40.81,   21.8,    41.4
)

grain_long <- grain_data %>%
  pivot_longer(cols = c(TGW, Protein, Starch),
               names_to = "Trait",
               values_to = "Value") %>%
  mutate(
    Trait = factor(Trait, levels = c("TGW", "Protein", "Starch")),
    Condition = factor(Condition, levels = c("Control","Heat","Drought","Combined"))
  )


# ===============================
# B) ENZYME DATA
# ===============================
enz_raw <- tribble(
  ~Condition, ~alpha_amylase, ~beta_amylase, ~limit_dextrinase, ~beta_glucanase,
  "Drought",   230, 78, 9, 179,
  "Drought",   214, 73, 10, 169,
  "Heat",      167, 45, 13, 133,
  "Heat",      167, 46, 12, 125,
  "Combined",  162, 80, 8, 152,
  "Combined",  172, 88, 10, 152,
  "Control",   187, 46, 22, 138,
  "Control",   187, 44, 18, 139,
  "Control",   188, 45, 22, 135
)

enz_long <- enz_raw %>%
  pivot_longer(cols = c(alpha_amylase, beta_amylase, limit_dextrinase, beta_glucanase),
               names_to = "Enzyme", values_to = "Activity") %>%
  mutate(
    Enzyme = recode(Enzyme,
                    "alpha_amylase"    = "Alpha amylase",
                    "beta_amylase"     = "Beta amylase",
                    "limit_dextrinase" = "Limit dextrinase",
                    "beta_glucanase"   = "Beta glucanase"),
    Condition = factor(Condition, levels = c("Control","Heat","Drought","Combined"))
  )

# ===============================
# C) PROTEIN-NORMALISED ENZYME DATA
# ===============================

enz_long_norm <- enz_long %>%
  left_join(
    grain_data %>% select(Condition, Protein),
    by = "Condition"
  ) %>%
  mutate(
    Activity_norm = Activity / (Protein / 100),
    Condition = factor(Condition, levels = c("Control","Heat","Drought","Combined"))
  )


#=====================================================================
# GLOBAL FONT → ARIAL
# =====================================================================
theme_set(theme_get() + theme(text = element_text(family = "Arial")))


# =====================================================================
# CUSTOM NATURE-STYLE THEME (UPDATED & CLEAN)
# =====================================================================
theme_nature <- function() {
    theme_classic(base_size = 20) +
        theme(
            text = element_text(family = "Arial"),
            
            # Facet labels
            strip.background = element_blank(),
            strip.text = element_text(size = 14, 
                                      margin = margin(b = 12)),
            
            # Layout & spacing
            panel.grid = element_blank(),
            panel.spacing = unit(2, "lines"),
            
            # Axes
            axis.title.x = element_text(size = 12, margin = margin(t = 10)),
            axis.title.y = element_text(size = 12),
            axis.text.x  = element_text(size = 10, angle = 45, hjust = 1),  # ← tilted labels
            axis.text.y  = element_text(size = 10),
            
            axis.line  = element_line(size = 0.4, color = "black"),
            axis.ticks = element_line(size = 0.4),
            
            legend.position = "none"
        )
}


# =====================================================================
# COLOUR PALETTE
# =====================================================================
palette4 <- c(
    "Control"  = "#FFFFFF",
    "Heat"     = "#E36B5D",
    "Drought"  = "#F2C75C",
    "Combined" = "#7EB5E6"
)


# =====================================================================
# PLOT 1 — Grain Traits
# =====================================================================
p1 <- grain_long %>%
    ggplot(aes(x = Condition, y = Value, fill = Condition)) +
    geom_col(width = 0.65, color = "black") +
    geom_jitter(
        width  = 0.12,
        size   = 2.4,
        shape  = 21,
        fill   = "white",
        color  = "black",
        stroke = 0.5
    ) +
    scale_fill_manual(values = palette4) +
    scale_y_continuous(expand = c(0,0)) +
    coord_cartesian(clip = "off") +
    facet_wrap(~Trait, scales = "free_y", nrow = 1) +
    ylab("Value") +
    theme_nature()

ggsave(
    "../../../R_export/Hv5/carlsberg_Hv5/Hv5_GrainTraits_NatureStyle.png",
    p1,
    width  = 16,
    height = 8,
    dpi    = 300,
    scale  = 0.9
)



# =====================================================================
# PLOT 2 — Enzymatic activities
# =====================================================================
p2 <- enz_long %>%
    ggplot(aes(x = Condition, y = Activity, fill = Condition)) +
    
    stat_summary(
        fun   = mean,
        geom  = "col",
        width = 0.65,
        color = "black",
        alpha = 0.85
    ) +
    
    stat_summary(
        fun.data = mean_se,
        geom     = "errorbar",
        width    = 0.25,
        size     = 0.5
    ) +
    
    geom_jitter(
        width  = 0.12,
        size   = 2.4,
        shape  = 21,
        fill   = "white",
        color  = "black",
        stroke = 0.5
    ) +
    
    scale_fill_manual(values = palette4) +
    scale_y_continuous(expand = c(0,0)) +
    coord_cartesian(clip = "off") +
    facet_wrap(~Enzyme, scales = "free_y", nrow = 1) +
    ylab("Enzymatic activity") +
    theme_nature()

ggsave(
    "../../../R_export/Hv5/carlsberg_Hv5/Hv5_Enzymes_NatureStyle.png",
    p2,
    width  = 18,
    height = 8,
    dpi    = 300,
    scale  = 0.9
)



# ===============================
# PLOT 3 — Protein-normalised enzymatic activities
# ===============================
p3 <- enz_long_norm %>%
  ggplot(aes(x = Condition, y = Activity_norm, fill = Condition)) +
  stat_summary(fun = mean, geom = "col", width = 0.65,
               color = "black", alpha = 0.85) +
  stat_summary(fun.data = mean_se, geom = "errorbar",
               width = 0.25, size = 0.5) +
  geom_jitter(width = 0.12, size = 2.4, shape = 21,
              fill = "white", color = "black", stroke = 0.5) +
  scale_fill_manual(values = palette4) +
  scale_y_continuous(expand = c(0,0)) +
  coord_cartesian(clip = "off") +
  facet_wrap(~Enzyme, scales = "free_y", nrow = 1) +
  ylab("Enzymatic activity (per g protein)") +
  theme_nature()

ggsave(
  "../../../R_export/Hv5/carlsberg_Hv5/Hv5_Enzymes_ProteinNorm_NatureStyle.pdf",
  plot   = p3,
  width  = 14,
  height = 4,
  device = cairo_pdf
)


```

