---
title: "DApipe_HvGerm_Exp6"
author: "Guido Giordano"
date: "2025-07-31"
output: html_document
---
#This is the pipeline for data analysis for barley germination experiment 6 

#Load packages
```{r}
# List packages (duplicates removed)
pkgs <- c(
  "AnnotationDbi","org.At.tair.db","clusterProfiler","dplyr","readr","tidyr",
  "ggplot2","ggbreak","cowplot","stringr","kableExtra","mice","reshape2",
  "data.table","cleaver","Mfuzz","Biobase","textshape","gridExtra","grid",
  "pcaMethods","impute","pheatmap","ggh4x","patchwork"
)

# Install BiocManager if missing
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager", repos = "https://cloud.r-project.org")

# CRAN vs Bioconductor separation
cran_pkgs <- pkgs[!pkgs %in% rownames(BiocManager::available())]
bioc_pkgs <- pkgs[pkgs %in% rownames(BiocManager::available())]

# Install missing CRAN packages
install_if_missing <- function(p, from_bioc = FALSE) {
  if (!requireNamespace(p, quietly = TRUE)) {
    if (from_bioc) {
      BiocManager::install(p, ask = FALSE, update = FALSE)
    } else {
      install.packages(p, repos = "https://cloud.r-project.org")
    }
  }
}

# Install missing ones
lapply(cran_pkgs, install_if_missing, from_bioc = FALSE)
lapply(bioc_pkgs, install_if_missing, from_bioc = TRUE)

# Load all
lapply(pkgs, function(p) library(p, character.only = TRUE))
```
#Calculate the digestion efficiency. This is directly connected with the goodness of protein identification. 

```{r}
#Load evidence (TD) file. 

evidence_b1 <- read.table("../../../R_import/Hv6/evidences/evidence_b1.txt", header = TRUE, sep = "\t")
evidence_b2 <- read.table("../../../R_import/Hv6/evidences/evidence_b2.txt", header = TRUE, sep = "\t")
evidence_b3 <- read.table("../../../R_import/Hv6/evidences/evidence_b3.txt", header = TRUE, sep = "\t")
evidence_b4 <- read.table("../../../R_import/Hv6/evidences/evidence_b4.txt", header = TRUE, sep = "\t")


#Generate a bar plot from the column missed cleavages
# Batch 1
Try_Dig_b1 <- ggplot(evidence_b1, aes(x = factor(Missed.cleavages))) +
  geom_bar(fill = "#105CB6", color = "black") +
  labs(title = "Tryptic digestion efficiency (Hv6_batch1)",
       x = "Missed Cleavages",
       y = "Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, family = "", face = "plain"),  
        axis.title = element_text(size = 12, family = "", face = "plain"),
        axis.text = element_text(size = 12, family = "", face = "plain")) +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)),
        axis.title.x = element_text(margin = margin(t = 8)),
        axis.title.y = element_text(margin = margin(r = 15)))
ggsave("../../../R_export/Hv6/trypdig/Hv6_b1_trypdig.png")

# Batch 2
Try_Dig_b2 <- ggplot(evidence_b2, aes(x = factor(Missed.cleavages))) +
  geom_bar(fill = "#105CB6", color = "black") +
  labs(title = "Tryptic digestion efficiency (Hv6_batch2)",
       x = "Missed Cleavages",
       y = "Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, family = "", face = "plain"),  
        axis.title = element_text(size = 12, family = "", face = "plain"),
        axis.text = element_text(size = 12, family = "", face = "plain")) +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)),
        axis.title.x = element_text(margin = margin(t = 8)),
        axis.title.y = element_text(margin = margin(r = 15)))
ggsave("../../../R_export/Hv6/trypdig/Hv6_b2_trypdig.png")

# Batch 3
Try_Dig_b3 <- ggplot(evidence_b3, aes(x = factor(Missed.cleavages))) +
  geom_bar(fill = "#105CB6", color = "black") +
  labs(title = "Tryptic digestion efficiency (Hv6_batch3)",
       x = "Missed Cleavages",
       y = "Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, family = "", face = "plain"),  
        axis.title = element_text(size = 12, family = "", face = "plain"),
        axis.text = element_text(size = 12, family = "", face = "plain")) +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)),
        axis.title.x = element_text(margin = margin(t = 8)),
        axis.title.y = element_text(margin = margin(r = 15)))
ggsave("../../../R_export/Hv6/trypdig/Hv6_b3_trypdig.png")

# Batch 4
Try_Dig_b4 <- ggplot(evidence_b4, aes(x = factor(Missed.cleavages))) +
  geom_bar(fill = "#105CB6", color = "black") +
  labs(title = "Tryptic digestion efficiency (Hv6_batch4)",
       x = "Missed Cleavages",
       y = "Count") +
  theme_classic() +
  theme(plot.title = element_text(size = 14, family = "", face = "plain"),  
        axis.title = element_text(size = 12, family = "", face = "plain"),
        axis.text = element_text(size = 12, family = "", face = "plain")) +
  theme(plot.title = element_text(hjust = 0.5, margin = margin(b = 10)),
        axis.title.x = element_text(margin = margin(t = 8)),
        axis.title.y = element_text(margin = margin(r = 15)))
ggsave("../../../R_export/Hv6/trypdig/Hv6_b4_trypdig.png")


# Combine all 20 plots into a single object
combined_plot <- Try_Dig_b1 + Try_Dig_b2 + Try_Dig_b3 + Try_Dig_b4
# Save the combined plot as a single image
ggsave("../../../R_export/Hv6/trypdig/combo_tryp_dig_Hv6.png", 
       plot = combined_plot, width = 20, height = 16, scale=0.5)

```


#Calculate the total number of MS3 spectra acquired per batch

```{r}

# List of file paths and batch labels
summary_files <- list(
  "summary_b1" = "../../../R_import/Hv6/summary/summary_b1.txt",
  "summary_b2" = "../../../R_import/Hv6/summary/summary_b2.txt",
  "summary_b3" = "../../../R_import/Hv6/summary/summary_b3.txt",
  "summary_b4" = "../../../R_import/Hv6/summary/summary_b4.txt"
)

# Function to read, clean, and plot a summary file
plot_ms3_per_fraction <- function(filepath, batch_label) {
  data <- read.table(filepath, header = TRUE, sep = "\t")
  data_clean <- data %>%
    filter(!is.na(Fraction)) %>%
    mutate(Fraction = factor(Fraction, levels = 1:48))

  ggplot(data_clean, aes(x = Fraction, y = MS3)) +
    geom_bar(stat = "identity", fill = "steelblue") +
    theme_minimal() +
    labs(
      title = paste("MS3 scans (", batch_label, ")", sep = ""),
      x = NULL,  # Removed x-axis label
      y = "MS3 Identifications"
    ) +
    theme(
      axis.text.x = element_text(angle = 90, vjust = 0.5, size = 6),  # Smaller font
      plot.title = element_text(size = 12, hjust = 0.5)
    )
}

# Generate all plots
plots <- mapply(plot_ms3_per_fraction, summary_files, names(summary_files), SIMPLIFY = FALSE)

# Combine using patchwork in a 2x2 layout
combined_plot <- (plots[[1]] | plots[[2]]) / (plots[[3]] | plots[[4]])

# Show combined plot
print(combined_plot)

# Save combined plot
ggsave("../../../R_export/Hv6/ms3_scans/ms3_ID_Hv6.png",
       combined_plot, width = 14, height = 10, dpi = 300, scale = 0.5)


```

#Calculate the sequence coverage 

```{r}
# --- Define input paths and output ---
pgroup_files <- list(
  "Batch 1" = "../../../R_import/Hv6/pGroups/pGroups_Hv6_b1_FDR.txt",
  "Batch 2" = "../../../R_import/Hv6/pGroups/pGroups_Hv6_b2_FDR.txt",
  "Batch 3" = "../../../R_import/Hv6/pGroups/pGroups_Hv6_b3_FDR.txt",
  "Batch 4" = "../../../R_import/Hv6/pGroups/pGroups_Hv6_b4_FDR.txt"
)

output_dir <- "../../../R_export/Hv6/sequence_cov/"
dir.create(output_dir, showWarnings = FALSE, recursive = TRUE)

# --- Read files into list ---
pGroups_list <- lapply(pgroup_files, function(path) {
  read.table(path, header = TRUE, sep = "\t", stringsAsFactors = FALSE)
})
names(pGroups_list) <- names(pgroup_files)

# --- Generate one plot per batch ---
seqcov_plots <- list()

for (batch_name in names(pGroups_list)) {
  df <- pGroups_list[[batch_name]]
  coverage <- df[[21]]  # column 21: sequence coverage

  p <- ggplot(data.frame(Coverage = coverage), aes(x = "", y = Coverage)) +
    geom_violin(trim = FALSE, fill = "lightblue", color = "black") +
    geom_boxplot(width = 0.1, fill = "white", color = "black", outlier.size = 0.5) +
    labs(
      title = paste0("seqcov_", tolower(gsub(" ", "", batch_name))),
      x = NULL,
      y = "Sequence Coverage"
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(size = 12, face = "plain", hjust = 0.5),
      axis.title = element_text(size = 10),
      axis.text.x = element_blank(),
      axis.ticks.x = element_blank()
    )
  
  seqcov_plots[[length(seqcov_plots) + 1]] <- p
}

# --- Combine plots with patchwork ---
combined_seqcov <- (seqcov_plots[[1]] | seqcov_plots[[2]]) /
                   (seqcov_plots[[3]] | seqcov_plots[[4]])

# --- Save combined figure ---
ggsave(
  filename = file.path(output_dir, "combined_seqcov_Hv6.png"),
  plot = combined_seqcov,
  width = 12,
  height = 10,
  dpi = 300, 
  scale=0.6
)

message("Saved combined sequence coverage plot.")


```

#Check data distribution and structure 
Here using ibaq protein groups after conversion with ibaq pipeline 
```{r}

# --- Load files ---
data_dir <- "../../../R_import/Hv6/pG_ibaq/"
file_list <- list.files(path = data_dir, pattern = "^pG_.*\\.csv$", full.names = TRUE)

for (file in file_list) {
  var_name <- gsub("^.*/pG_(.*)\\.csv$", "pG_\\1", file)
  assign(var_name, data.table::fread(file), envir = .GlobalEnv)
}

# --- Remove Contaminants, Reverse, or OIBS ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)
  if (all(c("Reverse", "Potential.contaminant", "Only.identified.by.site") %in% colnames(df))) {
    df <- df[Reverse != "+" & Potential.contaminant != "+" & Only.identified.by.site != "+", ]
  } else if ("Reverse" %in% colnames(df)) {
    df <- df[Reverse != "+", ]
  }
  assign(var_name, df, envir = .GlobalEnv)
}

# --- Log2 Transformation of iBAQ Columns ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)
  ibaq_cols <- grep("iBAQ", colnames(df), value = TRUE)
  df[, (ibaq_cols) := lapply(.SD, function(x) {
    x <- as.numeric(as.character(x))
    x[is.na(x)] <- 0
    log2(x + 1)
  }), .SDcols = ibaq_cols]
  assign(var_name, df, envir = .GlobalEnv)
}

# --- Median Centering of iBAQ Columns ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)
  ibaq_cols <- grep("iBAQ", colnames(df), value = TRUE)
  df[, (ibaq_cols) := lapply(.SD, function(x) x - median(x, na.rm = TRUE)), .SDcols = ibaq_cols]
  assign(var_name, df, envir = .GlobalEnv)
}

# --- Output directories ---
plot_dir <- "../../../R_export/Hv6/data_distribution/"
heatmap_dir <- "../../../R_export/Hv6/heatmaps/"
dir.create(plot_dir, showWarnings = FALSE, recursive = TRUE)
dir.create(heatmap_dir, showWarnings = FALSE, recursive = TRUE)

# --- Plotting & Heatmap Generation ---
histogram_plots <- list()
heatmap_plots <- list()

for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, ignore.case = TRUE)) next
  df <- get(var_name)

  if (nrow(df) == 0) {
    message("Skipping ", var_name, " - dataset is empty.")
    next
  }

  ## iBAQ histogram plots
  ibaq_cols_hist <- grep("_summed_iBAQ$", colnames(df), value = TRUE)
  if (length(ibaq_cols_hist) > 0) {
    plots <- lapply(seq_along(ibaq_cols_hist), function(i) {
      ggplot(df, aes_string(x = ibaq_cols_hist[i])) +
        geom_histogram(binwidth = 0.7, fill = "#105CB6", color = "black") +
        labs(title = paste("Batch", i), x = "Reporter Intensity (iBAQ)", y = "Frequency") +
        theme_classic() +
        xlim(-25, 15) +
        theme(
          plot.title = element_text(size = 12, hjust = 0.5),
          axis.title = element_text(size = 10),
          axis.text = element_text(size = 10)
        )
    })

    if (length(plots) > 0) {
      title <- grid::textGrob(var_name, gp = grid::gpar(fontsize = 16, fontface = "bold"))
      plot_grid <- gridExtra::arrangeGrob(grobs = plots, ncol = 3, top = title)

      ggplot2::ggsave(
        filename = file.path(plot_dir, paste0(var_name, "_iBAQ_distribution.png")),
        plot = plot_grid,
        width = 12,
        height = 9, 
        scale=0.9
      )

      # For patchwork: use first plot + title
      first_plot <- plots[[1]] + ggtitle(paste0("histogram_", sub(".*b(\\d+).*", "b\\1", var_name)))
      histogram_plots[[length(histogram_plots) + 1]] <- first_plot
      message("Saved distribution plot for ", var_name)
    }
  }

  ## Heatmap generation
  ibaq_cols_heatmap <- grep("_iBAQ$", colnames(df), value = TRUE)
  if (length(ibaq_cols_heatmap) == 11) {
    mat_raw <- as.matrix(df[, ..ibaq_cols_heatmap])
    rownames(mat_raw) <- df$Protein.IDs_pG
    mat_raw[is.na(mat_raw)] <- -23
    mat_scaled <- t(scale(t(mat_raw)))
    mat_scaled <- mat_scaled[apply(mat_scaled, 1, function(x) all(!is.na(x))), ]

    # Save static PNG
    heatmap_file <- file.path(heatmap_dir, paste0(var_name, "_heatmap_scaled.png"))
    pheatmap::pheatmap(mat_scaled,
      filename = heatmap_file,
      cluster_rows = TRUE,
      cluster_cols = FALSE,
      show_rownames = FALSE,
      labels_col = paste0("channel", 1:11)
    )
    message("Saved heatmap for ", var_name)

    # For patchwork (render in R, not saving image)
    heatmap_plot <- pheatmap::pheatmap(
      mat_scaled,
      silent = TRUE,
      cluster_rows = TRUE,
      cluster_cols = FALSE,
      show_rownames = FALSE,
      labels_col = paste0("channel", 1:11),
      main = paste0("heatmap_", sub(".*b(\\d+).*", "b\\1", var_name))
    )
    heatmap_plots[[length(heatmap_plots) + 1]] <- grid::grid.grabExpr(grid::grid.draw(heatmap_plot$gtable))
  } else {
    message("Skipping heatmap for ", var_name, " - found ", length(ibaq_cols_heatmap), " _iBAQ columns")
  }
}

# --- Combine and save 2x2 grid of histograms ---
if (length(histogram_plots) >= 4) {
  combined_hist <- (histogram_plots[[1]] | histogram_plots[[2]]) /
                   (histogram_plots[[3]] | histogram_plots[[4]])

  ggplot2::ggsave(
    filename = file.path(plot_dir, "combined_histograms.png"),
    plot = combined_hist,
    width = 16,
    height = 14, 
    scale=0.9
  )
  message("Saved combined histogram figure.")
}

# --- Combine and save 2x2 grid of heatmaps ---
if (length(heatmap_plots) >= 4) {
  combined_heat <- (wrap_elements(heatmap_plots[[1]]) | wrap_elements(heatmap_plots[[2]])) /
                   (wrap_elements(heatmap_plots[[3]]) | wrap_elements(heatmap_plots[[4]]))

  ggplot2::ggsave(
    filename = file.path(heatmap_dir, "combined_heatmaps.png"),
    plot = combined_heat,
    width = 16,
    height = 12, 
    scale=0.6
  )
  message("Saved combined heatmap figure.")
}

```

##Normalisation 
- In this script we take our protein groups with ibaq values and we filter, transform and normaliye them so that thez are readz to be used for quantiative comparisons  


```{r}
## =============================================================
## Hv6 NORMALIZATION PIPELINE — Hv4 LOGIC
## Approach 2: Protect corrected/imputed cells from TSN + inter-batch scaling
## =============================================================

suppressPackageStartupMessages({
  library(data.table)
  library(ggplot2)
  library(pheatmap)
  library(ggplotify)
  library(magick)
})

## --- Parameters ---
min_val    <- 0.1
heat_colors <- colorRampPalette(c("#4579F1", "white", "#D73020"))(100)

## --- Hv6 batch-specific thresholds and correction rules ---
batch_thresholds <- list(
  "BAR" = 4.0,
  "WHE" = 4.0,
  "RYE" = 4.0,
  "SOR" = 4.0
)

## No special rules (keeps behavior deterministic)
batch_specific_rules <- list(
  "BAR" = NULL,
  "WHE" = NULL,
  "RYE" = NULL,
  "SOR" = NULL
)

## Store per-dataset modification masks (TRUE = imputed or corrected)
mod_masks <- new.env(parent = emptyenv())


## =========================
## Moving-average smoothing (k = 3)
## =========================
smooth_transition <- function(x, k = 3) {
  if (all(is.na(x))) return(x)
  x_ext <- c(rep(x[1], floor(k/2)), x, rep(x[length(x)], floor(k/2)))
  y <- stats::filter(x_ext, rep(1/k, k), sides = 2)
  y <- y[(floor(k/2)+1):(length(y)-floor(k/2))]
  as.numeric(y)
}

smooth_matrix <- function(mat, k = 3) {
  t(apply(mat, 1, smooth_transition, k = k))
}


## =============================================================
## 1️⃣ LOAD AND CLEAN
## =============================================================

## --- Load files ---
data_dir  <- "../../../R_import/Hv6/pG_ibaq/"
file_list <- list.files(path = data_dir, pattern = "^pG_.*\\.csv$", full.names = TRUE)

for (file in file_list) {
  var_name <- gsub("^.*/pG_(.*)\\.csv$", "pG_\\1", file)
  assign(var_name, fread(file), envir = .GlobalEnv)
}

## --- Remove Reverse_pG / Potential.contaminant_pG ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, TRUE)) next
  df <- get(var_name)

  for (col in c("Reverse_pG", "Potential.contaminant_pG")) {
    if (col %in% names(df)) df[[col]] <- as.character(df[[col]])
  }
  if ("Reverse_pG" %in% names(df))               df <- df[is.na(Reverse_pG) | Reverse_pG != "+", ]
  if ("Potential.contaminant_pG" %in% names(df)) df <- df[is.na(Potential.contaminant_pG) | Potential.contaminant_pG != "+", ]

  assign(var_name, df, .GlobalEnv)
}

## --- Keep columns 1 and 72:80 (1 ID + 9 intensity channels) ---
for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, TRUE)) next
  df <- get(var_name)
  if (ncol(df) >= 80) {
    ## Note: original selects 1 and 72:80 (inclusive)
    df <- df[, c(1, 72:80), with = FALSE]
  } else {
    warning("Skipping ", var_name, " - not enough columns")
  }
  assign(var_name, df, .GlobalEnv)
}

## =============================================================
## 2️⃣ ADAPTIVE IMPUTATION & OUTLIER CORRECTION (Hv4 logic, per batch)
##     + build modification mask (TRUE where value was imputed or corrected)
## =============================================================
cat("\n--- Adaptive imputation & outlier correction (Hv4 logic, batch-specific) ---\n")

for (var_name in ls(pattern = "^pG_")) {
  if (grepl("final", var_name, TRUE)) next
  df <- get(var_name)

  ## Expect 1 ID + 9 channels
  if (ncol(df) < 10) next
  intensity_cols <- names(df)[2:10]
  df[, (intensity_cols) := lapply(.SD, as.numeric), .SDcols = intensity_cols]

  ## init mask: nrow x 9 channels
  mask <- matrix(FALSE, nrow = nrow(df), ncol = length(intensity_cols))

  ## Identify batch (e.g., BAR, WHE, RYE, SOR)
  batch_name <- toupper(gsub("^pG_([A-Za-z]+).*", "\\1", var_name))
  thr <- if (batch_name %in% names(batch_thresholds)) batch_thresholds[[batch_name]] else 2
  rules <- batch_specific_rules[[batch_name]]

  n_imp <- 0L
  n_corr <- 0L

  for (i in seq_len(nrow(df))) {
    row_vals <- unlist(df[i, ..intensity_cols])

    ## --- Apply batch-specific correction rules first ---
    if (!is.null(rules)) {
      for (rule in rules) {
        target   <- rule$target
        neigh_idx <- rule$neighbors
        if (all(neigh_idx %in% seq_along(row_vals))) {
          val <- row_vals[target]
          neigh_vals <- row_vals[neigh_idx]
          neigh_vals <- neigh_vals[is.finite(neigh_vals) & neigh_vals > 0]
          if (length(neigh_vals)) {
            neigh_mean <- mean(neigh_vals, na.rm = TRUE)
            if (is.na(val) || val < min_val) {
              row_vals[target] <- neigh_mean
              n_imp <- n_imp + 1L
              mask[i, target] <- TRUE
            } else {
              log2fc <- log2((val + 1) / (neigh_mean + 1))
              if (abs(log2fc) > thr) {
                row_vals[target] <- neigh_mean
                n_corr <- n_corr + 1L
                mask[i, target] <- TRUE
              }
            }
          }
        }
      }
    }

    ## --- Default adaptive correction for all channels ---
    rule_targets <- if (!is.null(rules)) vapply(rules, function(x) x$target, numeric(1)) else integer(0)
    for (j in seq_along(row_vals)) {
      if (length(rule_targets) && j %in% rule_targets) next
      val <- row_vals[j]
      neigh_idx <- c(j - 1L, j + 1L)
      neigh_idx <- neigh_idx[neigh_idx >= 1L & neigh_idx <= length(row_vals)]
      neigh_vals <- row_vals[neigh_idx]
      neigh_vals <- neigh_vals[is.finite(neigh_vals) & neigh_vals > 0]
      if (!length(neigh_vals)) next
      neigh_mean <- mean(neigh_vals, na.rm = TRUE)
      if (is.na(val) || val < min_val) {
        row_vals[j] <- neigh_mean
        n_imp <- n_imp + 1L
        mask[i, j] <- TRUE
      } else {
        log2fc <- log2((val + 1) / (neigh_mean + 1))
        if (abs(log2fc) > thr) {
          row_vals[j] <- neigh_mean
          n_corr <- n_corr + 1L
          mask[i, j] <- TRUE
        }
      }
    }

    df[i, (intensity_cols) := as.list(row_vals)]
  }

  cat(sprintf("✅ %s — threshold=%.1f | corrected: %d (imputed=%d, outliers=%d)\n",
              var_name, thr, n_imp + n_corr, n_imp, n_corr))

  assign(var_name, df, .GlobalEnv)
  mod_masks[[var_name]] <- mask
}

## =============================================================
## 3️⃣ QC FUNCTIONS
## =============================================================
make_qc_plot <- function(df, title) {
  intensity_cols <- names(df)[2:10]
  sums    <- colSums(df[, ..intensity_cols], na.rm = TRUE)
  medians <- vapply(intensity_cols, function(cn) median(df[[cn]], na.rm = TRUE), numeric(1))

  stats_df <- data.table(Channel = factor(seq_along(intensity_cols)),
                         Total_Intensity = sums, Median_Intensity = medians)
  max_total <- max(stats_df$Total_Intensity, na.rm = TRUE)
  max_med   <- max(stats_df$Median_Intensity,  na.rm = TRUE)
  scale_factor <- if (is.finite(max_total / max_med) && max_med > 0) max_total / max_med else 1

  ggplot(stats_df, aes(x = Channel)) +
    geom_bar(aes(y = Total_Intensity), stat = "identity", fill = "steelblue", alpha = 0.6) +
    geom_point(aes(y = Median_Intensity * scale_factor), color = "red", size = 2.5) +
    geom_line(aes(y = Median_Intensity * scale_factor, group = 1), color = "red") +
    scale_y_continuous(name = "Total Intensity",
                       sec.axis = sec_axis(~ . / scale_factor, name = "Median Intensity")) +
    labs(title = title, x = "Channel") +
    theme_classic()
}

make_heatmap <- function(df, title, outfile = NA) {
  intensity_cols <- names(df)[2:10]
  mat <- as.matrix(df[, ..intensity_cols])
  mode(mat) <- "numeric"
  mat[!is.finite(mat)] <- NA
  mat_log <- log10(mat + 1)
  row_sd <- apply(mat_log, 1, sd, na.rm = TRUE)
  mat_log <- mat_log[is.finite(row_sd) & row_sd > 0, , drop = FALSE]
  colnames(mat_log) <- paste0("Ch", seq_len(ncol(mat_log)))
  pheatmap(mat_log, filename = outfile, main = title, scale = "row", border_color = NA,
           cluster_cols = FALSE, color = heat_colors, clustering_method = "average", silent = TRUE)
}

## =============================================================
## 4️⃣ NORMALIZATION, PLOTTING, EXPORT
## =============================================================

## --- Paths ---
out_dir_pre      <- "../../../R_export/Hv6/TS_mi"              ## Pre-normalization plots
out_dir_post_out <- "../../../R_export/Hv6/TS_mi_norm_inter"   ## Post inter-batch plots
out_dir_combo    <- "../../../R_export/Hv6/TS_mi_combined"     ## Collages of post-inter plots
final_out_dir    <- "../../../R_Data/Hv6/pG_iBAQ_postnorm"     ## Final normalized data

for (d in c(out_dir_pre, out_dir_post_out, out_dir_combo, final_out_dir)) {
  if (!dir.exists(d)) dir.create(d, recursive = TRUE)
}

## --- Pre-norm QC plots ---
cat("\n--- Pre-normalization summaries (bars + heatmaps) ---\n")
for (nm in ls(pattern = "^pG_")) {
  if (grepl("final", nm, TRUE)) next
  df <- get(nm)
  if (ncol(df) < 10) next
  ggsave(file.path(out_dir_pre, paste0(nm, "_preNorm_sumMedian.png")),
         make_qc_plot(df, paste("Total & Median Intensity —", nm, "(pre-norm)")),
         width = 10, height = 5, dpi = 150)
  make_heatmap(df, paste("Heatmap —", nm, "(pre-norm, log10, row-scaled)"),
               file.path(out_dir_pre, paste0(nm, "_preNorm_heatmap.png")))
}

## --- Intra-batch total-sum normalization (TSN) ---
## Compute factors from UNMODIFIED cells and apply scaling ONLY to UNMODIFIED cells
cat("\n--- Intra-batch total-sum normalization (protecting modified cells) ---\n")
for (nm in ls(pattern = "^pG_")) {
  if (grepl("final", nm, TRUE)) next
  df <- get(nm)
  if (ncol(df) < 10) next
  intensity_cols <- names(df)[2:10]
  mask <- mod_masks[[nm]]
  if (is.null(mask)) mask <- matrix(FALSE, nrow = nrow(df), ncol = length(intensity_cols))

  ## Column sums from UNMODIFIED cells only
  col_sums <- sapply(seq_along(intensity_cols), function(j) {
    v <- df[[intensity_cols[j]]]
    idx <- !mask[, j] & is.finite(v)
    sum(v[idx], na.rm = TRUE)
  })
  target  <- median(col_sums[is.finite(col_sums)], na.rm = TRUE)
  factors <- target / col_sums

  ## Apply scaling ONLY to UNMODIFIED cells
  for (j in seq_along(intensity_cols)) {
    col <- intensity_cols[j]
    v <- df[[col]]
    idx <- !mask[, j] & is.finite(v)
    v[idx] <- v[idx] * factors[j]
    df[[col]] <- v
  }
  
  ## ----------------------------------------
  ## Moving-average smoothing (k = 3)
  ## ----------------------------------------
  mat_postTSN <- as.matrix(df[, ..intensity_cols])
  mat_smooth  <- smooth_matrix(mat_postTSN, k = 3)

  df[, (intensity_cols) := as.data.table(mat_smooth)]

  cat(sprintf("✨ %s — Moving-average smoothing applied (k=3)\n", nm))

  
  assign(nm, df, .GlobalEnv)
}

## --- Inter-batch median centering (protect modified cells) ---
cat("\n--- Inter-batch median centering (protecting modified cells) ---\n")
batch_names <- Filter(function(x) !grepl("final", x, TRUE), ls(pattern = "^pG_"))

## Per-batch medians from UNMODIFIED cells only
batch_meds <- sapply(batch_names, function(nm) {
  df <- get(nm)
  if (ncol(df) < 10) return(NA_real_)
  intensity_cols <- names(df)[2:10]
  mask <- mod_masks[[nm]]
  if (is.null(mask)) mask <- matrix(FALSE, nrow = nrow(df), ncol = length(intensity_cols))
  vals <- c()
  for (j in seq_along(intensity_cols)) {
    v <- df[[intensity_cols[j]]]
    idx <- !mask[, j] & is.finite(v)
    if (any(idx)) vals <- c(vals, v[idx])
  }
  median(vals, na.rm = TRUE)
})

grand_median <- median(batch_meds, na.rm = TRUE)

for (nm in batch_names) {
  df <- get(nm)
  if (ncol(df) < 10) next
  intensity_cols <- names(df)[2:10]
  mask <- mod_masks[[nm]]
  if (is.null(mask)) mask <- matrix(FALSE, nrow = nrow(df), ncol = length(intensity_cols))

  factor <- grand_median / batch_meds[nm]

  ## Apply inter-batch scaling ONLY to UNMODIFIED cells
  for (j in seq_along(intensity_cols)) {
    col <- intensity_cols[j]
    v <- df[[col]]
    idx <- !mask[, j] & is.finite(v)
    v[idx] <- v[idx] * factor
    df[[col]] <- v
  }

  assign(nm, df, .GlobalEnv)
}

## --- Post inter-batch plots ---
cat("\n--- Post inter-batch normalization plots ---\n")
for (nm in batch_names) {
  df <- get(nm)
  if (ncol(df) < 10) next
  ggsave(file.path(out_dir_post_out, paste0(nm, "_postInter_sumMedian.png")),
         make_qc_plot(df, paste("Total & Median Intensity —", nm, "(post inter-batch)")),
         width = 10, height = 5, dpi = 150)
  make_heatmap(df, paste("Heatmap —", nm, "(post inter-batch, log10, row-scaled)"),
               file.path(out_dir_post_out, paste0(nm, "_postInter_heatmap.png")))
}

## --- Combine all post-inter plots ---
cat("\n--- Combining all post-inter plots (magick) ---\n")
bar_files <- list.files(out_dir_post_out, pattern = "sumMedian\\.png$", full.names = TRUE)
hm_files  <- list.files(out_dir_post_out, pattern = "heatmap\\.png$",   full.names = TRUE)

if (length(bar_files) > 0) {
  imgs <- lapply(bar_files, function(f) image_resize(image_read(f), "1600x1600"))
  collage <- image_montage(image_join(imgs), tile = "2x2", geometry = "+10+10")
  image_write(collage, file.path(out_dir_combo, "AllBatches_postInter_BarplotCollage.png"), format = "png")
}
if (length(hm_files) > 0) {
  imgs <- lapply(hm_files, function(f) image_resize(image_read(f), "1600x1600"))
  collage <- image_montage(image_join(imgs), tile = "2x2", geometry = "+10+10")
  image_write(collage, file.path(out_dir_combo, "AllBatches_postInter_HeatmapCollage.png"), format = "png")
}

## --- Log2 transformation + export ---
for (nm in batch_names) {
  df <- get(nm)
  if (ncol(df) < 10) next
  intensity_cols <- names(df)[2:10]
  df[, (intensity_cols) := lapply(.SD, function(x) log2(as.numeric(x) + 1)), .SDcols = intensity_cols]
  fwrite(df, file.path(final_out_dir, paste0(nm, "_norm_log2.csv")))
}

cat("\n✅ Hv6 normalization pipeline (Approach 2) completed successfully.\n")


```

#In this chunk we map intensity values to the (hierarchical) orthogroups  

```{r}

library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(data.table)  # for fread
library(tibble)      # for as_tibble

### Annotate the pG normalized ----
### Load Orthogroups and expand them with unassigned proteins
library(tidyverse)

# Load original (OrthoFinder) orthogroup table
OrthoG_full_iso <- read_tsv("Z:/PhD Munchen/Guido personal PhD Project_1/Barley germination proteome/Resources for Data analysis/R/R_import/Hv6/orthos/N0_Hv6.tsv")

# Load Unassigned Proteins
Unassigned_proteins <- read_tsv("../../../R_import/Hv6/orthos/UnassignedGenes_Hv6.tsv")


# Harmonize column names BEFORE combining
# 1. Ensure unassigned table has OG column
if ("Orthogroup" %in% names(Unassigned_proteins)) {
    Unassigned_proteins <- Unassigned_proteins %>% rename(OG = Orthogroup)
}
# 2. Ensure unassigned table has HOG column — create synthetic if missing
if (!"HOG" %in% names(Unassigned_proteins)) {
    Unassigned_proteins <- Unassigned_proteins %>%
        mutate(HOG = paste0("UNASSIGNED_", row_number()))
}
# 3. Standardize species column names
rename_species_cols <- function(df) {
    df %>% rename(
        barley   = matches("Barke|barley", ignore.case = TRUE),
        rye      = matches("SECALE|rye", ignore.case = TRUE),
        sorghum  = matches("Sbicolor|sorghum", ignore.case = TRUE),
        wheat    = matches("wheat", ignore.case = TRUE),
        clade    = matches("Gene Tree Parent Clade|clade", ignore.case = TRUE)
    )
}
OrthoG_full_iso    <- rename_species_cols(OrthoG_full_iso)
Unassigned_proteins <- rename_species_cols(Unassigned_proteins)

# Combine both tables
OrthoG_exp <- bind_rows(OrthoG_full_iso, Unassigned_proteins)

# Finally enforce correct column order
OrthoG_exp <- OrthoG_exp %>%
  select(HOG, barley, rye, sorghum, wheat)

OrthoG_long <- OrthoG_exp %>%
  pivot_longer(cols = c(barley, rye, sorghum, wheat),
               names_to = "Species", values_to = "Proteins") %>%
  separate_rows(Proteins, sep = ",") %>%
  mutate(Proteins = str_trim(Proteins))


### --- Attach normalised intensity values ----

# Load post-normalised Hv6 files (9 channels exported: cols 2:10)
data_dir <- "../../../R_Data/Hv6/pG_iBAQ_postnorm"
file_list <- list.files(path = data_dir, pattern = "_norm_log2\\.csv$", full.names = TRUE)

# Read as tibbles for dplyr
mat_list <- lapply(file_list, function(p) as_tibble(fread(p)))
names(mat_list) <- gsub("^.*/(.*)\\.csv$", "\\1", file_list)

prot_col <- "Protein.IDs_pG"
# Use column NAMES for the 9 channels (2:10)
get_ibaq_names <- function(df) names(df)[2:10]

# Helper to attach intensities for one species
attach_species <- function(species_name, proteome_df) {
  ibaq_names <- get_ibaq_names(proteome_df)
  OrthoG_sub <- OrthoG_long %>% filter(Species == species_name)

  merged <- proteome_df %>%
    right_join(OrthoG_sub, by = setNames("Proteins", prot_col))  # Protein.IDs_pG ↔︎ Proteins

  out <- merged %>%
    filter(if_any(all_of(ibaq_names), ~ !is.na(.))) %>%
    group_by(HOG) %>%
    summarise(
      Proteins = paste(unique(.data[[prot_col]]), collapse = ","),
      across(all_of(ibaq_names), ~ mean(.x, na.rm = TRUE), .names = "{.col}")
    ) %>%
    ungroup()

  # Rename intensity columns to iBAQ_norm_ch1..ch9_species
  old <- ibaq_names
  new <- paste0("iBAQ_norm_ch", seq_along(old), "_", species_name)
  names(out)[match(old, names(out))] <- new
  names(out)[names(out) == "Proteins"] <- paste0("Proteins_", species_name)
  out
}

# Map file names to species (adjust if your file names differ)
barley_df  <- attach_species("barley",   mat_list[["pG_BAR_b1_norm_log2"]])
rye_df     <- attach_species("rye",      mat_list[["pG_RYE_b3_norm_log2"]])
sorghum_df <- attach_species("sorghum",  mat_list[["pG_SOR_b4_norm_log2"]])
wheat_df   <- attach_species("wheat",    mat_list[["pG_WHE_b2_norm_log2"]])

# Merge all into one orthogroup table
ortog_int <- barley_df %>%
  full_join(rye_df,     by = "HOG") %>%
  full_join(sorghum_df, by = "HOG") %>%
  full_join(wheat_df,   by = "HOG")

# Preview & save
print(head(ortog_int, 10))

out_dir <- "../../../R_Data/Hv6/ortog_anno"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)
out_file <- file.path(out_dir, "ortog_int.csv")
data.table::fwrite(ortog_int, out_file)


### --- Attach normalised intensity values (NO MEAN, NO MERGING) ----
### --- This is a second version to make heatmaps later on 

library(data.table)
library(dplyr)
library(tibble)

# --------------------------------------------------
# Load post-normalised Hv6 files
# --------------------------------------------------
data_dir <- "../../../R_Data/Hv6/pG_iBAQ_postnorm"
file_list <- list.files(path = data_dir, pattern = "_norm_log2\\.csv$", full.names = TRUE)

mat_list <- lapply(file_list, function(p) as_tibble(fread(p)))
names(mat_list) <- gsub("^.*/(.*)\\.csv$", "\\1", file_list)

prot_col <- "Protein.IDs_pG"

# Extract intensity column names (channels 2:10)
get_ibaq_names <- function(df) names(df)[2:10]

# --------------------------------------------------
# Helper: attach intensities WITHOUT collapsing
# --------------------------------------------------
attach_species_long <- function(species_name, proteome_df) {

  ibaq_names <- get_ibaq_names(proteome_df)

  # Filter orthogroup table for species
  OrthoG_sub <- OrthoG_long %>% filter(Species == species_name)

  # Join protein intensities
  merged <- proteome_df %>%
    inner_join(
      OrthoG_sub,
      by = setNames("Proteins", prot_col)  # Protein.IDs_pG ↔ Proteins
    )

  # Keep rows with at least one intensity
  merged <- merged %>%
    filter(if_any(all_of(ibaq_names), ~ !is.na(.)))

  # Rename intensity columns to species-aware format
  new_names <- paste0("iBAQ_norm_ch", seq_along(ibaq_names), "_", species_name)
  names(merged)[match(ibaq_names, names(merged))] <- new_names

  # Clean output structure
  merged %>%
    transmute(
      HOG = HOG,
      Species = species_name,
      Protein_ID = .data[[prot_col]],
      across(starts_with("iBAQ_norm_ch"))
    )
}

# --------------------------------------------------
# Attach intensities PER SPECIES (no collapsing)
# --------------------------------------------------
barley_long  <- attach_species_long("barley",   mat_list[["pG_BAR_b1_norm_log2"]])
rye_long     <- attach_species_long("rye",      mat_list[["pG_RYE_b3_norm_log2"]])
sorghum_long <- attach_species_long("sorghum",  mat_list[["pG_SOR_b4_norm_log2"]])
wheat_long   <- attach_species_long("wheat",    mat_list[["pG_WHE_b2_norm_log2"]])

# --------------------------------------------------
# Stack all species together (still long)
# --------------------------------------------------
ortog_long_int <- bind_rows(
  barley_long,
  rye_long,
  sorghum_long,
  wheat_long
)

# --------------------------------------------------
# Preview
# --------------------------------------------------
print(head(ortog_long_int, 20))

# --------------------------------------------------
# Save output
# --------------------------------------------------
out_dir <- "../../../R_Data/Hv6/ortog_anno"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

out_file <- file.path(out_dir, "ortog_long_int_noMean.csv")
data.table::fwrite(ortog_long_int, out_file)

cat("✔ Saved long orthogroup protein-level matrix:\n", out_file, "\n")


```

##In this script we add functional annotations to our proteins

```{r}
                      
#Load the DB matches between the species specific fasta and the uniprot species specific fasta, so that each fasta entry is associated with a uniprot entry 
# set the folder path
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_up"

# Load each file separately
barley_hvup <- read.delim(file.path(path, "barke_upbarley.tsv"), sep = "\t", header = FALSE, stringsAsFactors = FALSE)
rye_scpg <- read.csv(file.path(path, "rye_0225_featuretab_Stein.csv"), header = TRUE, stringsAsFactors = FALSE)
sorghum_sbup <- read.delim(file.path(path, "sorghum_upsorghum.tsv"), sep = "\t", header = FALSE, stringsAsFactors = FALSE)
wheat_taup <- read.delim(file.path(path, "wheat_upwheat.tsv"), sep = "\t", header = FALSE, stringsAsFactors = FALSE)

# Keep only first 2 columns and rename them consistently
barley_hvup  <- barley_hvup[, 1:2]
rye_scpg     <- rye_scpg[, c(1,5)]
sorghum_sbup <- sorghum_sbup[, 1:2]
wheat_taup   <- wheat_taup[, 1:2]

colnames(barley_hvup)  <- c("barleyup_entry", "barley_entry")
colnames(rye_scpg)     <- c("rye_entry",     "rye_proteinname")
colnames(sorghum_sbup) <- c("sbup_entry", "sorghum_entry")
colnames(wheat_taup)   <- c("taup_entry", "wheat_entry")

# Clean annotation files to allow matching 
barley_hvup$barleyup_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_hvup$barleyup_entry)
sorghum_sbup$sbup_entry     <- sub("^[^|]*\\|[^|]*\\|", "", sorghum_sbup$sbup_entry)
wheat_taup$taup_entry       <- sub("^[^|]*\\|[^|]*\\|", "", wheat_taup$taup_entry)

#deduplicate entries 
barley_hvup <- barley_hvup %>%
  distinct(barley_entry, .keep_all = TRUE)
sorghum_sbup <- sorghum_sbup %>%
  distinct(sorghum_entry, .keep_all = TRUE)
wheat_taup <- wheat_taup %>%
  distinct(wheat_entry, .keep_all = TRUE)
  
##Attach functional annotation from uniprot to the uniprot matched entries 
# set the folder path
path <- "../../../R_import/Hv6/anno/masters_up"

# Load each file separately
up_Hv <- read.delim(file.path(path, "unip_Hv_1505.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
up_Sb <- read.delim(file.path(path, "unip_Sb_270825.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
up_Ta <- read.delim(file.path(path, "unip_Ta_1505.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
up_At <- read.delim(file.path(path, "unip_At_1505.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Keep only first 2 columns and rename them consistently
up_Hv  <- up_Hv[, c(3,4,5,7,8,9)]
colnames(up_Hv)  <- c("barleyup_entryname", "barleyup_pname", "barleyup_gname", "barleyup_GO_bp", "barleyup_GO_mf", "barleyup_GO_cc")

up_Sb <- up_Sb[, c(2,3,4,5,7,6)]
colnames(up_Sb)  <- c("sorghumup_entryname", "sorghumup_pname", "sorghumup_gname", "sorghumup_GO_bp", "sorghumup_GO_mf", "sorghumup_GO_cc")

up_Ta   <- up_Ta[, c(3,4,5,7,8,9)]
colnames(up_Ta)  <- c("wheatup_entryname", "wheatup_pname", "wheatup_gname", "wheatup_GO_bp", "wheatup_GO_mf", "wheatup_GO_cc")

up_At     <- up_At[, c(2,3,4,5,7,6)]
colnames(up_At)  <- c("arabidopsisup_entryname", "arabidopsisup_pname", "arabidopsisup_gname", "arabidopsisup_GO_bp", "arabidopsisup_GO_mf", "arabidopsisup_GO_cc")

#Attach the functional annotation to the DB matches and then to the orthology file. I set all.x=TRUE, because i dont want to have annotations for proteins that are not matching 

barley_anno <- merge(
  barley_hvup,
  up_Hv,
  by.x = "barleyup_entry",
  by.y = "barleyup_entryname",
  all.x = TRUE
)
wheat_anno <- merge(
  wheat_taup,
  up_Ta,
  by.x = "taup_entry",
  by.y = "wheatup_entryname",
  all.x = TRUE
)
sorghum_anno <- merge(
  sorghum_sbup,
  up_Sb,
  by.x = "sbup_entry",
  by.y = "sorghumup_entryname",
  all.x = TRUE
)

#Now we want to add Arabidopsis thaliana annotations, especially to be able to run GO enrichment. So we expand the annotations with arabidopsis thaliana annotations and then we will merge everything to the orthology matrix
##Load the matches generated with Diamond BLAST between all the 4 species specific fasta files and the uniprot Arabidopsis thaliana specific fasta file. This allows us to attach Arabidopsis thaliana GO terms, usefull to perform enrichment analysis
# set the folder path
path <- "../../../R_import/Hv6/anno/DB_fasta_vs_upAt"

# Load each file separately
barley_ATup <- read.delim(file.path(path, "araT_barke.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
rye_ATup <- read.delim(file.path(path, "araT_rye.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
sorghum_ATup <- read.delim(file.path(path, "araT_Sbicolor.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)
wheat_ATup <- read.delim(file.path(path, "araT_wheat.tsv"), sep = "\t", header = TRUE, stringsAsFactors = FALSE)

# Keep only first 2 columns and rename them consistently
barley_ATup  <- barley_ATup[, 1:2]
rye_ATup     <- rye_ATup[, 1:2]
sorghum_ATup <- sorghum_ATup[, 1:2]
wheat_ATup   <- wheat_ATup[, 1:2]

colnames(barley_ATup)  <- c("barley_entry",  "araT_entry")
colnames(rye_ATup)     <- c("rye_entry",     "araT_entry")
colnames(sorghum_ATup) <- c("sorghum_entry", "araT_entry")
colnames(wheat_ATup)   <- c("wheat_entry",   "araT_entry")

# Clean annotation files to allow matching 
barley_ATup$araT_entry <- sub("^[^|]*\\|[^|]*\\|", "", barley_ATup$araT_entry)
rye_ATup$araT_entry     <- sub("^[^|]*\\|[^|]*\\|", "", rye_ATup$araT_entry)
sorghum_ATup$araT_entry       <- sub("^[^|]*\\|[^|]*\\|", "", sorghum_ATup$araT_entry)
wheat_ATup$araT_entry       <- sub("^[^|]*\\|[^|]*\\|", "", wheat_ATup$araT_entry)

#deduplicate entries 
barley_ATup <- barley_ATup %>%
  distinct(barley_entry, .keep_all = TRUE)
sorghum_ATup <- sorghum_ATup %>%
  distinct(sorghum_entry, .keep_all = TRUE)
wheat_ATup <- wheat_ATup %>%
  distinct(wheat_entry, .keep_all = TRUE)
rye_ATup <- rye_ATup %>%
  distinct(rye_entry, .keep_all = TRUE)
  
#Merge each annotation file with arabidopsis thaliana entries 
barley_anno_expat <- merge(
  barley_anno,
  barley_ATup,
  by = "barley_entry",
  all.x = TRUE
)

rye_anno_expat <- merge(
  rye_scpg,
  rye_ATup,
  by = "rye_entry",
  all.x = TRUE
)

sorghum_anno_expat <- merge(
  sorghum_anno,
  sorghum_ATup,
  by = "sorghum_entry",
  all.x = TRUE
)

wheat_anno_expat <- merge(
  wheat_anno,
  wheat_ATup,
  by = "wheat_entry",
  all.x = TRUE
)

#Attach functional annotation from Atuniprot
At_UP_annotated <- read_tsv("../../../R_import/Hv6/anno/masters_up/unip_At_1505.tsv")
At_UP_annotated  <- At_UP_annotated[, c(2,3,4,5,6,7,8)]
At_UP_annotated <- At_UP_annotated %>%
  mutate(TAIR = sub(";.*", "", TAIR))

#Finally get the final annotation file 
# Barley
barley_full_anno <- merge(
  barley_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)
barley_full_anno <- barley_full_anno[, c(2, 3, 4, 5, 6, 7, 8, 1, 14, 9, 10, 11, 12, 13)]
colnames(barley_full_anno) <- c("barley_entry","barleyup_entry","barleyup_pname","barleyup_gname",
                                "barleyup_GO_bp","barleyup_GO_mf","barleyup_GO_cc","araT_entry_hv",
                                "TAIR_hv","atup_pname_hv","atup_gname_hv","at_GO_bp_hv",
                                "at_GO_mf_hv","at_GO_cc_hv")

# Rye
rye_full_anno <- merge(
  rye_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)
rye_full_anno <- rye_full_anno[, c(2, 3, 1, 9,4, 5, 6, 7, 8)]
colnames(rye_full_anno) <- c("rye_entry","rye_pname","araT_entry_rye",
                             "TAIR_rye","atup_pname_rye","atup_gname_rye","at_GO_bp_rye",
                             "at_GO_mf_rye","at_GO_cc_rye")

# Sorghum
sorghum_full_anno <- merge(
  sorghum_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)
sorghum_full_anno <- sorghum_full_anno[, c(2, 3, 4, 5, 6, 7, 8, 1, 14, 9, 10, 11, 12, 13)]
colnames(sorghum_full_anno) <- c("sorghum_entry","sorghumup_entry","sorghumup_pname","sorghumup_gname",
                                 "sorghumup_GO_bp","sorghumup_GO_mf","sorghumup_GO_cc","araT_entry_sorghum",
                                "TAIR_sorghum","atup_pname_sorghum","atup_gname_sorghum","at_GO_bp_sorghum",
                                 "at_GO_mf_sorghum","at_GO_cc_sorghum")

# Wheat
wheat_full_anno <- merge(
  wheat_anno_expat,
  At_UP_annotated,
  by.x = "araT_entry",
  by.y = "Entry Name",
  all.x = TRUE
)
wheat_full_anno <- wheat_full_anno[, c(2, 3, 4, 5, 6, 7, 8, 1, 14, 9, 10, 11, 12, 13)]
colnames(wheat_full_anno) <- c("wheat_entry","wheatup_entry","wheatup_pname","wheatup_gname",
                               "wheatup_GO_bp","wheatup_GO_mf","wheatup_GO_cc","araT_entry_wheat",
                               "TAIR_wheat","atup_pname_wheat","atup_gname_wheat","at_GO_bp_wheat",
                               "at_GO_mf_wheat","at_GO_cc_wheat")


#Now the most extraordinary thing, matching all this to the orthologues matrix

#Load orthogroups annotated with intensity values, that needs to be expanded with functional annotations
ortog_int <- read.csv("../../../R_Data/Hv6/ortog_anno/ortog_int.csv",
                      header = TRUE, stringsAsFactors = FALSE)
# Create helper column with only the first protein ID
ortog_int$Proteins_barley_main <- sub(",.*", "", ortog_int$Proteins_barley)
ortog_int$Proteins_rye_main <- sub(",.*", "", ortog_int$Proteins_rye)
ortog_int$Proteins_sorghum_main <- sub(",.*", "", ortog_int$Proteins_sorghum)
ortog_int$Proteins_wheat_main <- sub(",.*", "", ortog_int$Proteins_wheat)

# Merge using the helper column
ortog_int_annoHv <- merge(
  ortog_int,
  barley_full_anno,
  by.x = "Proteins_barley_main",
  by.y = "barley_entry",
  all.x = TRUE
)
ortog_int_annoHvTa <- merge(
  ortog_int_annoHv,
  wheat_full_anno,
  by.x = "Proteins_wheat_main",
  by.y = "wheat_entry",
  all.x = TRUE
)
ortog_int_annoHvTaSb <- merge(
  ortog_int_annoHvTa,
  sorghum_full_anno,
  by.x = "Proteins_sorghum_main",
  by.y = "sorghum_entry",
  all.x = TRUE
)
Hv6_final_matrix <- merge(
  ortog_int_annoHvTaSb,
  rye_full_anno,
  by.x = "Proteins_rye_main",
  by.y = "rye_entry",
  all.x = TRUE
)

#Final adjustment in the order of the columns 
Hv6_final_matrix <- Hv6_final_matrix[
  , c(
    "HOG",
  
    # Barley block
    "Proteins_barley","Proteins_barley_main",
    grep("^iBAQ_norm_ch[0-9]+_barley$", colnames(Hv6_final_matrix), value = TRUE),
    "barleyup_entry","barleyup_pname","barleyup_gname",
    "barleyup_GO_bp","barleyup_GO_mf","barleyup_GO_cc",
    "araT_entry_hv","TAIR_hv","atup_pname_hv","atup_gname_hv",
    "at_GO_bp_hv","at_GO_mf_hv","at_GO_cc_hv",
    
    # Wheat block
    "Proteins_wheat","Proteins_wheat_main",
    grep("^iBAQ_norm_ch[0-9]+_wheat$", colnames(Hv6_final_matrix), value = TRUE),
    "wheatup_entry","wheatup_pname","wheatup_gname",
    "wheatup_GO_bp","wheatup_GO_mf","wheatup_GO_cc",
    "araT_entry_wheat","TAIR_wheat","atup_pname_wheat","atup_gname_wheat",
    "at_GO_bp_wheat","at_GO_mf_wheat","at_GO_cc_wheat",
    
    # Rye block
    "Proteins_rye","Proteins_rye_main",
    grep("^iBAQ_norm_ch[0-9]+_rye$", colnames(Hv6_final_matrix), value = TRUE),
    "rye_pname","araT_entry_rye","TAIR_rye","atup_pname_rye","atup_gname_rye",
    "at_GO_bp_rye","at_GO_mf_rye","at_GO_cc_rye",
    
    # Sorghum block
    "Proteins_sorghum","Proteins_sorghum_main",
    grep("^iBAQ_norm_ch[0-9]+_sorghum$", colnames(Hv6_final_matrix), value = TRUE),
    "sorghumup_entry","sorghumup_pname","sorghumup_gname",
    "sorghumup_GO_bp","sorghumup_GO_mf","sorghumup_GO_cc",
    "araT_entry_sorghum","TAIR_sorghum","atup_pname_sorghum","atup_gname_sorghum",
    "at_GO_bp_sorghum","at_GO_mf_sorghum","at_GO_cc_sorghum"
  )
]

# Save final matrix
outdir <- "../../../R_Data/Hv6"
if (!dir.exists(outdir)) {
  dir.create(outdir, recursive = TRUE)
}

write.csv(
  Hv6_final_matrix,
  file = file.path(outdir, "Hv6_final_matrix.csv"),
  row.names = FALSE
)


# ================================================================
# Attach annotations to LONG orthogroup matrix (NO MEAN VERSION)
# ================================================================

library(dplyr)
library(data.table)

# Load long protein-level orthogroup matrix
ortog_long <- read.csv(
  "../../../R_Data/Hv6/ortog_anno/ortog_long_int_noMean.csv",
  stringsAsFactors = FALSE
)

# --------------------------------------------------
# Split by species for clean joins
# --------------------------------------------------
barley_long  <- ortog_long %>% filter(Species == "barley")
rye_long     <- ortog_long %>% filter(Species == "rye")
sorghum_long <- ortog_long %>% filter(Species == "sorghum")
wheat_long   <- ortog_long %>% filter(Species == "wheat")

# --------------------------------------------------
# Attach annotations directly by protein ID
# --------------------------------------------------

# Barley
barley_annot <- barley_long %>%
  left_join(barley_full_anno, by = c("Protein_ID" = "barley_entry"))

# Rye
rye_annot <- rye_long %>%
  left_join(rye_full_anno, by = c("Protein_ID" = "rye_entry"))

# Sorghum
sorghum_annot <- sorghum_long %>%
  left_join(sorghum_full_anno, by = c("Protein_ID" = "sorghum_entry"))

# Wheat
wheat_annot <- wheat_long %>%
  left_join(wheat_full_anno, by = c("Protein_ID" = "wheat_entry"))

# --------------------------------------------------
# Recombine everything
# --------------------------------------------------
Hv6_final_matrix_no_mean <- bind_rows(
  barley_annot,
  wheat_annot,
  rye_annot,
  sorghum_annot
)

# --------------------------------------------------
# Optional: reorder columns (clean structure)
# --------------------------------------------------
Hv6_final_matrix_no_mean <- Hv6_final_matrix_no_mean %>%
  relocate(HOG, Species, Protein_ID)

# --------------------------------------------------
# Save new matrix
# --------------------------------------------------
outdir <- "../../../R_Data/Hv6"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

write.csv(
  Hv6_final_matrix_no_mean,
  file = file.path(outdir, "Hv6_final_matrix_no_mean.csv"),
  row.names = FALSE
)

cat("✔ Saved protein-resolved annotated matrix:\n",
    file.path(outdir, "Hv6_final_matrix_no_mean.csv"), "\n")

```

###Visualize full dataset 
in this chunk I visualize the full dataset and I also have a script to visualize specific protein families (helpful to quickly check one protein of interest)

```{r}
####See the entire dataset 
library(pheatmap)
library(ggplot2)

# 1. Load data
Hv6_final_matrix <- read.csv("../../../R_Data/Hv6/Hv6_final_matrix.csv",
                             header = TRUE, stringsAsFactors = FALSE)

# 2. Define timepoints (use unique labels)
timepoints <- c("0hs","12hs","24hs","36hs",
                "48hs","60hs","72hs","84hs","96hs")

# 3. Function to rename iBAQ columns
rename_ibaq <- function(df, species){
  old <- grep(paste0("^iBAQ_norm_ch[0-9]+_", species), names(df), value = TRUE)
  new <- paste0("iBAQ_norm_", timepoints, "_", species)
  names(df)[match(old, names(df))] <- new
  return(df)
}

Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, "barley")
Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, "wheat")
Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, "rye")
Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, "sorghum")

## Generate a heatmap of the full dataset
# 1. Define species and timepoints for ordering
species <- c("barley", "wheat", "rye", "sorghum")
timepoints <- c("0hs","12hs","24hs","36hs",
                "48hs","60hs","72hs","84hs","96hs")

# 2. Build the column order explicitly
iBAQ_cols <- unlist(lapply(species, function(sp) {
  paste0("iBAQ_norm_", timepoints, "_", sp)
}))
iBAQ_cols <- iBAQ_cols[iBAQ_cols %in% names(Hv6_final_matrix)]

# 3. Build matrix
iBAQ_matrix <- as.matrix(Hv6_final_matrix[, iBAQ_cols])

# 4. Replace NA with 1
iBAQ_matrix[is.na(iBAQ_matrix)] <- 1

# 5. Clean column names (remove "iBAQ_norm_")
colnames(iBAQ_matrix) <- gsub("^iBAQ_norm_", "", colnames(iBAQ_matrix))

# 6. Prepare output folder
outdir <- "../../../R_export/Hv6/unbiased"
if (!dir.exists(outdir)) {
  dir.create(outdir, recursive = TRUE)
}

# 7. Save heatmap as JPEG
jpeg(file.path(outdir, "Hv6_heatmap_unbiased.jpeg"), width = 12, height = 8, units = "in", res = 300)
pheatmap(
  iBAQ_matrix,
  show_rownames = FALSE,
  show_colnames = TRUE,
  cluster_rows = TRUE,
  cluster_cols = FALSE,  # keep order barley → wheat → rye → sorghum
  scale = "row",
  color = colorRampPalette(c("blue", "white", "firebrick3"))(50),
  legend_labels = "zscore_iBAQ"
)
dev.off()

## --- PCA ANALYSIS ---

# Transpose: samples as rows
pca_res <- prcomp(t(iBAQ_matrix), scale. = TRUE)

# Calculate explained variance
var_expl <- pca_res$sdev^2 / sum(pca_res$sdev^2)
pc1_var <- round(var_expl[1] * 100, 1)
pc2_var <- round(var_expl[2] * 100, 1)

# Build dataframe for plotting
pca_df <- as.data.frame(pca_res$x[, 1:2])
pca_df$Sample <- rownames(pca_df)
pca_df$Species <- sub(".*_", "", pca_df$Sample)
pca_df$Time <- sub("_.*", "", pca_df$Sample)

# Plot PCA
p_pca <- ggplot(pca_df, aes(x = PC1, y = PC2, color = Species, label = Time)) +
  geom_point(size = 4) +
  geom_text(vjust = 1.5, size = 3) +
  theme_classic() +
  ggtitle("Proteomic similarity") +
  theme(plot.title = element_text(hjust = 0.5)) +  # center title
  xlab(paste0("PC1 (", pc1_var, "%)")) +
  ylab(paste0("PC2 (", pc2_var, "%)"))

# Save
ggsave(file.path(outdir, "Hv6_PCA.jpeg"), p_pca, width = 7, height = 6, dpi = 300)


```

###Generate profile plots from candidate protein families 
In this chunk we generate profile plots for candidate protein families. 
```{r}
## ============================================================
## Hv6 — Malting enzyme heatmap (with Hv5 family definitions)
## ============================================================

library(data.table)
library(dplyr)
library(tidyr)
library(pheatmap)
library(RColorBrewer)

# ----------------------------
# 1. Load dataset
# ----------------------------

Hv6_final_matrix <- read.csv(
  "../../../R_Data/Hv6/Hv6_final_matrix.csv",
  header = TRUE,
  stringsAsFactors = FALSE
) %>%
  filter(!grepl("^UNASSIGNED_", .[[1]]))


# ----------------------------
# 2. Rename iBAQ columns
# ----------------------------
timepoints <- c("0hs","12hs","24hs","36hs",
                "48hs","60hs","72hs","84hs","96hs")

rename_ibaq <- function(df, species){
  old <- grep(paste0("^iBAQ_norm_ch[0-9]+_", species), names(df), value = TRUE)
  new <- paste0("iBAQ_norm_", timepoints, "_", species)
  names(df)[match(old, names(df))] <- new
  return(df)
}

species <- c("barley", "wheat", "rye", "sorghum")
for (sp in species) {
  Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, sp)
}

# ----------------------------
# 3. Use Hv5 enzyme family definitions
# ----------------------------

all_families_ID <- list(
    `Alpha-amylase` = c(
        "HORVU.BARKE.PROJ.6HG00559210.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00559200.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00559230.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00559240.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00559250.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00559270.1",  # Amy1-1
        "HORVU.BARKE.PROJ.6HG00561370.1",  # Amy1-2
        "HORVU.BARKE.PROJ.7HG00653200.1",  # Amy2-1
        "HORVU.BARKE.PROJ.7HG00653310.1",  # Amy2-3
        "HORVU.BARKE.PROJ.7HG00653330.1",  # Amy2-2
        "HORVU.BARKE.PROJ.5HG00441970.1",  # Amy3
        "HORVU.BARKE.PROJ.2HG00148940.1",  # Amy4-1
        "HORVU.BARKE.PROJ.3HG00259920.1"   # Amy4-2
    ),
    
    `Beta-amylase` = c(
        "HORVU.BARKE.PROJ.2HG00126310.1",  # Bam5
        "HORVU.BARKE.PROJ.2HG00126320.1",  # Bam5
        "HORVU.BARKE.PROJ.4HG00380960.1",  # Bam5
        "HORVU.BARKE.PROJ.1HG00036700.1",  # Bam1
        "HORVU.BARKE.PROJ.1HG00036740.1",  # Bam1
        "HORVU.BARKE.PROJ.1HG00051030.1",  # Bam3
        "HORVU.BARKE.PROJ.2HG00105510.1",  # Bam9
        "HORVU.BARKE.PROJ.3HG00228930.1",  # Bam1
        "HORVU.BARKE.PROJ.4HG00303690.1",  # Bam2
        "HORVU.BARKE.PROJ.4HG00376360.1",  # Bam1
        "HORVU.BARKE.PROJ.6HG00502080.1"   # Bam8
    ),
    
    `Alpha-glucosidase` = c(
        "HORVU.BARKE.PROJ.7HG00666400.1",  # Agl1
        "HORVU.BARKE.PROJ.3HG00202550.1",  # Agl2
        "HORVU.BARKE.PROJ.1HG00052860.1",  # Agl3
        "HORVU.BARKE.PROJ.4HG00360800.1",  # Agl4
        "HORVU.BARKE.PROJ.2HG00192840.1",  # Agl1
        "HORVU.BARKE.PROJ.2HG00192870.1",  # Agl1
        "HORVU.BARKE.PROJ.2HG00192880.1"   # Agl1
    ),
    
    `Isoamylase` = c(
        "HORVU.BARKE.PROJ.7HG00618070.1",  # Isa1
        "HORVU.BARKE.PROJ.1HG00057560.1",  # Isa2
        "HORVU.BARKE.PROJ.5HG00443610.1",  # Isa3
        "HORVU.BARKE.PROJ.2HG00149340.1",  # Sbe2-2
        "HORVU.BARKE.PROJ.2HG00154890.1",  # Sbe2-1
        "HORVU.BARKE.PROJ.7HG00645630.1",  # Sbe3
        "HORVU.BARKE.PROJ.7HG00681480.1",  # Sbe2-2
        "HORVU.BARKE.PROJ.7HG00822990.1",  # Sbe2-2
        "HORVU.BARKE.PROJ.7HG00823000.1"   # Sbe2-2
    ), 
    
    `Limit dextrinase` = c(
        "HORVU.BARKE.PROJ.7HG00595270.1"    # Lda
    ), 
    
    `Alpha glucan phosphorylase` = c(
        "HORVU.BARKE.PROJ.5HG00465050.1",  # Phs1
        "HORVU.BARKE.PROJ.3HG00275600.1",  # Phs2
        "HORVU.BARKE.PROJ.2HG00708120.1",  # Phs2
        "HORVU.BARKE.PROJ.3HG00275620.1"   # Phs2
    ),
    Cellulase = c(
  "HORVU.BARKE.PROJ.1HG00073020.1",  # T2h3-5
  "HORVU.BARKE.PROJ.1HG00018950.1",  # F9d16-30
  "HORVU.BARKE.PROJ.3HG00226090.1",  # Gh9c3
  "HORVU.BARKE.PROJ.4HG00379800.1",  # Gh9c3
  "HORVU.BARKE.PROJ.2HG00192140.1",  # T21l14-7
  "HORVU.BARKE.PROJ.5HG00427900.1",  # Cel2
  "HORVU.BARKE.PROJ.7HG00596870.1",  # T21l14-7
  "HORVU.BARKE.PROJ.7HG00606250.1",  # T21l14-7
  "HORVU.BARKE.PROJ.5HG00452800.1",  # F19h22-110
  "HORVU.BARKE.PROJ.6HG00554880.1",  # Cel1
  "HORVU.BARKE.PROJ.6HG00561940.1",  # T28a8-150
  "HORVU.BARKE.PROJ.7HG00653770.1",  # F10a5-13
  "HORVU.BARKE.PROJ.6HG00508740.1",  # T21l14-7
  "HORVU.BARKE.PROJ.5HG00453760.1",  # F6f9-1
  "HORVU.BARKE.PROJ.4HG00311880.1",  # Kor1
  "HORVU.BARKE.PROJ.6HG00558010.1",  # T21l14-7
  "HORVU.BARKE.PROJ.5HG00434990.1",  # T21l14-7
  "HORVU.BARKE.PROJ.1HG00078380.1",  # Gh9c3
  "HORVU.BARKE.PROJ.7HG00607220.1",  # Cel1
  "HORVU.BARKE.PROJ.4HG00345900.1",  # Kor1
  "HORVU.BARKE.PROJ.2HG00162350.1",  # Kor1
  "HORVU.BARKE.PROJ.1HG00014980.1",  # F4i1-55
  "HORVU.BARKE.PROJ.2HG00154880.1",  # Cel2
  "HORVU.BARKE.PROJ.7HG00638840.1"   # Cel5
),

Lichenase = c(
  "HORVU.BARKE.PROJ.7HG00679910.1",  # Dl4170c
  "HORVU.BARKE.PROJ.1HG00053050.1"   # Dl4170c
),

`Beta-glucosidase` = c(
  "HORVU.BARKE.PROJ.3HG00279420.1",  # Bglu42
  "HORVU.BARKE.PROJ.3HG00266610.1",  # Bglc1
  "HORVU.BARKE.PROJ.3HG00266590.1"   # Bglc1
),

Arabinofuranosidase = c(
  "HORVU.BARKE.PROJ.4HG00325990.1",  # Asd1
  "HORVU.BARKE.PROJ.4HG00326450.1",  # Asd1
  "HORVU.BARKE.PROJ.5HG00420480.1",  # Asd1
  "HORVU.BARKE.PROJ.4HG00346830.1",  # Asd1
  "HORVU.BARKE.PROJ.2HG00101670.1"   # Asd1
),

Mannanase = c(
  "HORVU.BARKE.PROJ.3HG00253120.1",  # Man7
  "HORVU.BARKE.PROJ.5HG00771660.1",  # Man6
  "HORVU.BARKE.PROJ.6HG00561230.1",  # Man7
  "HORVU.BARKE.PROJ.6HG00561220.1",  # Man7
  "HORVU.BARKE.PROJ.5HG00479880.1",  # Man3
  "HORVU.BARKE.PROJ.3HG00263090.1",  # Man5
  "HORVU.BARKE.PROJ.7HG00612160.1"   # Man2
),

`Alpha-galactosidase` = c(
  "HORVU.BARKE.PROJ.1HG00041210.1",  # Agal2
  "HORVU.BARKE.PROJ.1HG00041240.1",  # Agal1
  "HORVU.BARKE.PROJ.6HG00537080.1",  # Apse
  "HORVU.BARKE.PROJ.6HG00556460.1",  # Apse
  "HORVU.BARKE.PROJ.5HG00772740.1",  # Agal2
  "HORVU.BARKE.PROJ.5HG00772750.1",  # Agal1
  "HORVU.BARKE.PROJ.2HG00102930.1"   # Agal3
),

Polygalacturonase = c(
  "HORVU.BARKE.PROJ.5HG00475020.1",  # Adpg2
  "HORVU.BARKE.PROJ.5HG00475050.1",  # Adpg2
  "HORVU.BARKE.PROJ.7HG00599930.1",  # Adpg2
  "HORVU.BARKE.PROJ.4HG00360780.1",  # F18b13-25
  "HORVU.BARKE.PROJ.6HG00502310.1"   # F24j13-7
),

`Pectin esterase` = c(
  "HORVU.BARKE.PROJ.2HG00100460.1",  # Pme21
  "HORVU.BARKE.PROJ.1HG00054490.1",  # Pme41
  "HORVU.BARKE.PROJ.2HG00157350.1",  # Pme24
  "HORVU.BARKE.PROJ.3HG00268440.1",  # Pme44
  "HORVU.BARKE.PROJ.2HG00188100.1",  # Pme13
  "HORVU.BARKE.PROJ.5HG00473670.1",  # Pme43
  "HORVU.BARKE.PROJ.1HG00058460.1",  # Pme17
  "HORVU.BARKE.PROJ.5HG00437860.1",  # Pme43
  "HORVU.BARKE.PROJ.4HG00349290.1",  # Pme28
  "HORVU.BARKE.PROJ.6HG00563120.1",  # Pme3
  "HORVU.BARKE.PROJ.1HG00085260.1",  # Pme38
  "HORVU.BARKE.PROJ.2HG00107890.1",  # Pme66
  "HORVU.BARKE.PROJ.2HG00103570.1",  # Pme44
  "HORVU.BARKE.PROJ.3HG00230000.1",  # Pme32
  "HORVU.BARKE.PROJ.3HG00232130.1",  # Pme22
  "HORVU.BARKE.PROJ.2HG00105250.1",  # Pme66
  "HORVU.BARKE.PROJ.1HG00045130.1",  # Pme31
  "HORVU.BARKE.PROJ.1HG00069330.1",  # Pme53
  "HORVU.BARKE.PROJ.2HG00102050.1",  # Pme53
  "HORVU.BARKE.PROJ.2HG00115940.1",  # Qrt1
  "HORVU.BARKE.PROJ.2HG00105260.1",  # Pme66
  "HORVU.BARKE.PROJ.2HG00171960.1",  # Pme68
  "HORVU.BARKE.PROJ.2HG00107880.1",  # Pme66
  "HORVU.BARKE.PROJ.4HG00328640.1",  # Pme63
  "HORVU.BARKE.PROJ.2HG00154320.1",  # Pme51
  "HORVU.BARKE.PROJ.3HG00300590.1",  # Pme67
  "HORVU.BARKE.PROJ.4HG00366450.1",  # Pme8
  "HORVU.BARKE.PROJ.5HG00391190.1",  # Pme44
  "HORVU.BARKE.PROJ.6HG00564170.1",  # Pme44
  "HORVU.BARKE.PROJ.5HG00403060.1",  # Qrt1
  "HORVU.BARKE.PROJ.7HG00593290.1",  # Pme29
  "HORVU.BARKE.PROJ.7HG00598930.1",  # Pme2
  "HORVU.BARKE.PROJ.3HG00280980.1",  # Pme51
  "HORVU.BARKE.PROJ.7HG00639410.1",  # Pme15
  "HORVU.BARKE.PROJ.4HG00328630.1"   # Pme63
),
`papain-like cysteine endopeptidase` = c(
  "HORVU.BARKE.PROJ.6HG00551890.1",  # Xcp1
  "HORVU.BARKE.PROJ.6HG00495640.1",  # F23e12.90
  "HORVU.BARKE.PROJ.5HG00420360.1",  # Xcp1
  "HORVU.BARKE.PROJ.1HG00004010.1",  # Xcp2
  "HORVU.BARKE.PROJ.3HG00278240.1",  # Cep1
  "HORVU.BARKE.PROJ.6HG00495650.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495660.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495670.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495680.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495690.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495700.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.6HG00495710.1",  # F28n24.20 
  "HORVU.BARKE.PROJ.2HG00144580.1",  # Rd19c
  "HORVU.BARKE.PROJ.2HG00120520.1",  # Rd19d
  "HORVU.BARKE.PROJ.2HG00186610.1",  # Rd21a
  "HORVU.BARKE.PROJ.2HG00193350.1",  # Rd21b
  "HORVU.BARKE.PROJ.1HG00069020.1",  # Xbcp3
  "HORVU.BARKE.PROJ.5HG00459870.1",  # Cep1
  "HORVU.BARKE.PROJ.7HG00679480.1",  # F28n24.20
  "HORVU.BARKE.PROJ.4HG00311230.1",  # Cathb2
  "HORVU.BARKE.PROJ.2HG00187860.1",  # Cep2
  "HORVU.BARKE.PROJ.1HG00004790.1",  # F28n24.27
  "HORVU.BARKE.PROJ.2HG00192770.1",  # Xbcp3
  "HORVU.BARKE.PROJ.1HG00077400.1",  # Cep1
  "HORVU.BARKE.PROJ.5HG00386780.1"   # Cep3
),

`serine endopeptidase` = c(
  "HORVU.BARKE.PROJ.6HG00555990.1",  # Prss
  "HORVU.BARKE.PROJ.5HG00440510.1",  # Prss
  "HORVU.BARKE.PROJ.5HG00396740.1",  # Prss
  "HORVU.BARKE.PROJ.1HG00044350.1"   # Prss
),

`aspartic endopeptidase` = c(
  "HORVU.BARKE.PROJ.1HG00081950.1",  # Apa1
  "HORVU.BARKE.PROJ.3HG00249950.1",  # Apa2
  "HORVU.BARKE.PROJ.3HG00291880.1",  # Apa1
  "HORVU.BARKE.PROJ.6HG00567790.1",  # Apa1
  "HORVU.BARKE.PROJ.4HG00318350.1",  # F17m5-250
  "HORVU.BARKE.PROJ.7HG00652070.1",  # Apa1
  "HORVU.BARKE.PROJ.1HG00016940.1",  # Apa1
  "HORVU.BARKE.PROJ.5HG00415360.1"   # F17m5-250
),

Aminopeptidase = c(
  "HORVU.BARKE.PROJ.5HG00432120.1",  # Apm1
  "HORVU.BARKE.PROJ.4HG00305500.1",  # Apm1
  "HORVU.BARKE.PROJ.6HG00524520.1"   # Apm1
),

`tripeptidyl peptidase` = c(
  "HORVU.BARKE.PROJ.6HG00545030.1"   # Tpp2
),

`dipeptidyl peptidase` = c(
  "HORVU.BARKE.PROJ.5HG00483050.1",  # F8k7-10
  "HORVU.BARKE.PROJ.5HG00406500.1"   # Mop9-8
),

Carboxypeptidase = c(
  "HORVU.BARKE.PROJ.2HG00171920.1",  # Scpl18
  "HORVU.BARKE.PROJ.5HG00445760.1",  # Scpl9
  "HORVU.BARKE.PROJ.3HG00215140.1",  # Scpl18
  "HORVU.BARKE.PROJ.5HG00469870.1",  # Amp1
  "HORVU.BARKE.PROJ.1HG00008170.1",  # F27d4-19
  "HORVU.BARKE.PROJ.1HG00018290.1",  # Scpl27
  "HORVU.BARKE.PROJ.2HG00107460.1",  # Scpl40
  "HORVU.BARKE.PROJ.1HG00041660.1",  # Scpl51
  "HORVU.BARKE.PROJ.2HG00144730.1",  # Scpl20
  "HORVU.BARKE.PROJ.1HG00070610.1",  # Scpl51
  "HORVU.BARKE.PROJ.1HG00084240.1",  # Scpl50
  "HORVU.BARKE.PROJ.2HG00171840.1",  # Scpl18
  "HORVU.BARKE.PROJ.1HG00048630.1",  # Scpl34
  "HORVU.BARKE.PROJ.2HG00095780.1",  # Scpl50
  "HORVU.BARKE.PROJ.3HG00265950.1",  # Mpa24-11
  "HORVU.BARKE.PROJ.3HG00222540.1",  # Scpl27
  "HORVU.BARKE.PROJ.2HG00148330.1",  # Scpl45
  "HORVU.BARKE.PROJ.1HG00084350.1",  # Scpl50
  "HORVU.BARKE.PROJ.2HG00183040.1",  # Scpl49
  "HORVU.BARKE.PROJ.4HG00316830.1",  # Scpl42
  "HORVU.BARKE.PROJ.2HG00166340.1",  # Scpl35
  "HORVU.BARKE.PROJ.3HG00230350.1",  # Scpl33
  "HORVU.BARKE.PROJ.4HG00304700.1",  # Scpl40
  "HORVU.BARKE.PROJ.3HG00262480.1",  # F10a2-10
  "HORVU.BARKE.PROJ.5HG00442780.1",  # Scpl40
  "HORVU.BARKE.PROJ.4HG00760680.1",  # Scpl51
  "HORVU.BARKE.PROJ.4HG00311110.1",  # Scpl2
  "HORVU.BARKE.PROJ.2HG00722350.1",  # Scpl10
  "HORVU.BARKE.PROJ.5HG00442850.1",  # Scpl28
  "HORVU.BARKE.PROJ.5HG00442860.1",  # Scpl28
  "HORVU.BARKE.PROJ.3HG00230360.1",  # Scpl34
  "HORVU.BARKE.PROJ.6HG00557070.1",  # Scpl49
  "HORVU.BARKE.PROJ.3HG00262490.1",  # F10a2-10
  "HORVU.BARKE.PROJ.5HG00442910.1",  # Scpl40
  "HORVU.BARKE.PROJ.6HG00499600.1",  # Scpl51
  "HORVU.BARKE.PROJ.5HG00442940.1",  # Scpl40
  "HORVU.BARKE.PROJ.4HG00364910.1",  # Scpl27
  "HORVU.BARKE.PROJ.4HG00315620.1",  # Scpl11
  "HORVU.BARKE.PROJ.7HG00625610.1",  # Scpl51
  "HORVU.BARKE.PROJ.5HG00442970.1",  # Scpl40
  "HORVU.BARKE.PROJ.5HG00445740.1",  # Scpl18
  "HORVU.BARKE.PROJ.7HG00593540.1",  # Scpl3
  "HORVU.BARKE.PROJ.7HG00668820.1",  # Scpl51
  "HORVU.BARKE.PROJ.5HG00442890.1",  # Scpl27
  "HORVU.BARKE.PROJ.6HG00540890.1",  # Scpl34
  "HORVU.BARKE.PROJ.6HG00494930.1",  # Scpl48
  "HORVU.BARKE.PROJ.7HG00673840.1",  # F27d4-19
  "HORVU.BARKE.PROJ.2HG00107490.1",  # Scpl40
  "HORVU.BARKE.PROJ.4HG00324700.1",  # Scpl40
  "HORVU.BARKE.PROJ.7HG00653610.1",  # F27d4-19
  "HORVU.BARKE.PROJ.7HG00656000.1",  # Scpl34
  "HORVU.BARKE.PROJ.7HG00673290.1",  # Scpl34
  "HORVU.BARKE.PROJ.7HG00587970.1",  # Sol1
  "HORVU.BARKE.PROJ.7HG00591350.1",  # F1k23-6
  "HORVU.BARKE.PROJ.6HG00565120.1",  # Scpl24
  "HORVU.BARKE.PROJ.3HG00285590.1",  # Scpl20
  "HORVU.BARKE.PROJ.7HG00597910.1",  # Scpl29
  "HORVU.BARKE.PROJ.3HG00277430.1",  # Scpl27
  "HORVU.BARKE.PROJ.6HG00522350.1",  # Scpl45
  "HORVU.BARKE.PROJ.1HG00063380.1"   # F1p2-60
),

Lipoxygenase = c(
  "HORVU.BARKE.PROJ.5HG00383870.1",  # Lox2
  "HORVU.BARKE.PROJ.4HG00307600.1",  # Lox1
  "HORVU.BARKE.PROJ.6HG00519710.1",  # Lox2
  "HORVU.BARKE.PROJ.6HG00490920.1",  # Lox2
  "HORVU.BARKE.PROJ.4HG00307640.1",  # Lox1
  "HORVU.BARKE.PROJ.7HG00579310.1",  # Lox2
  "HORVU.BARKE.PROJ.5HG00462390.1",  # Lox5
  "HORVU.BARKE.PROJ.2HG00188780.1",  # Lox1
  "HORVU.BARKE.PROJ.2HG00156010.1",  # Lox3
  "HORVU.BARKE.PROJ.4HG00369200.1",  # Lox4
  "HORVU.BARKE.PROJ.4HG00369210.1",  # Lox3
  "HORVU.BARKE.PROJ.5HG00481340.1",  # Lox1
  "HORVU.BARKE.PROJ.7HG00616920.1",  # Lox2
  "HORVU.BARKE.PROJ.2HG00188750.1",  # Lox1
  "HORVU.BARKE.PROJ.5HG00462410.1",  # Lox1
  "HORVU.BARKE.PROJ.5HG00462420.1",  # Lox1
  "HORVU.BARKE.PROJ.5HG00462430.1",  # Lox1
  "HORVU.BARKE.PROJ.5HG00462440.1",  # Lox1
  "HORVU.BARKE.PROJ.1HG00050580.1",  # Lox5
  "HORVU.BARKE.PROJ.4HG00760420.1"   # Lox2
),

`allene oxide synthase` = c(
  "HORVU.BARKE.PROJ.4HG00359600.1",  # Aos
  "HORVU.BARKE.PROJ.4HG00359650.1"   # Aos
),

`acid phosphatase` = c(
  "HORVU.BARKE.PROJ.1HG00004770.1",  # Pap22
  "HORVU.BARKE.PROJ.1HG00078980.1",  # Pap27
  "HORVU.BARKE.PROJ.2HG00164150.1",  # Pap3
  "HORVU.BARKE.PROJ.3HG00231330.1",  # Pap15
  "HORVU.BARKE.PROJ.5HG00448690.1",  # Pap27
  "HORVU.BARKE.PROJ.5HG00417950.1",  # Pap20
  "HORVU.BARKE.PROJ.5HG00430080.1",  # Pap23
  "HORVU.BARKE.PROJ.6HG00507050.1",  # Pap18
  "HORVU.BARKE.PROJ.5HG00417910.1",  # Pap21
  "HORVU.BARKE.PROJ.5HG00448710.1",  # Pap27
  "HORVU.BARKE.PROJ.7HG00675360.1",  # Pap26
  "HORVU.BARKE.PROJ.7HG00620380.1",  # Pap27
  "HORVU.BARKE.PROJ.2HG00149170.1",  # Pap22
  "HORVU.BARKE.PROJ.2HG00185580.1",  # Pap22
  "HORVU.BARKE.PROJ.5HG00454460.1",  # Pap15
  "HORVU.BARKE.PROJ.5HG00419620.1",  # Pap23
  "HORVU.BARKE.PROJ.3HG00267210.1",  # Pap10
  "HORVU.BARKE.PROJ.4HG00361290.1",  # Pap27
  "HORVU.BARKE.PROJ.2HG00142220.1",  # Pap2
  "HORVU.BARKE.PROJ.5HG00472100.1",  # Pap9
  "HORVU.BARKE.PROJ.2HG00112820.1",  # Pap3
  "HORVU.BARKE.PROJ.5HG00400860.1",  # Pap27
  "HORVU.BARKE.PROJ.4HG00376810.1",  # Pap27
  "HORVU.BARKE.PROJ.6HG00565200.1",  # Pap10
  "HORVU.BARKE.PROJ.4HG00358100.1",  # Pap3
  "HORVU.BARKE.PROJ.5HG00475530.1",  # Pap10
  "HORVU.BARKE.PROJ.4HG00372650.1",  # Pap10
  "HORVU.BARKE.PROJ.3HG00269070.1",  # Pap27
  "HORVU.BARKE.PROJ.2HG00124710.1",  # Pap
  "HORVU.BARKE.PROJ.1HG00084950.1",  # Pap1
  "HORVU.BARKE.PROJ.6HG00793860.1",  # Pap
  "HORVU.BARKE.PROJ.6HG00797590.1",  # Pap1
  "HORVU.BARKE.PROJ.7HG00629580.1",  # Pap1
  "HORVU.BARKE.PROJ.4HG00755000.1",  # Pap10
  "HORVU.BARKE.PROJ.6HG00533910.1"   # F28i8-17
),

Catalase = c(
  "HORVU.BARKE.PROJ.6HG00495590.1",  # Cat3
  "HORVU.BARKE.PROJ.7HG00681400.1",  # Cat2
  "HORVU.BARKE.PROJ.4HG00374170.1"   # Cat2
),

`superoxide dismutase` = c(
  "HORVU.BARKE.PROJ.2HG00190130.1",  # Msd1
  "HORVU.BARKE.PROJ.2HG00105600.1",  # Csd1
  "HORVU.BARKE.PROJ.4HG00360670.1",  # Csd3
  "HORVU.BARKE.PROJ.5HG00440200.1",  # Csd1
  "HORVU.BARKE.PROJ.5HG00440830.1",  # Csd1
  "HORVU.BARKE.PROJ.7HG00580950.1",  # Fsd2
  "HORVU.BARKE.PROJ.7HG00588580.1",  # Fsd3
  "HORVU.BARKE.PROJ.7HG00625840.1"   # Csd2
),

`polyphenol oxidase` = c(
  "HORVU.BARKE.PROJ.3HG00268620.1",  # Ppo
  "HORVU.BARKE.PROJ.2HG00177380.1",  # Ppo1
  "HORVU.BARKE.PROJ.2HG00177440.1"   # Po0
)


)
barke_gene_map <- c(

  # =========================
  # Alpha-amylase
  # =========================
  "HORVU.BARKE.PROJ.6HG00559210.1" = "Amy1-1.1",
  "HORVU.BARKE.PROJ.6HG00559200.1" = "Amy1-1.2",
  "HORVU.BARKE.PROJ.6HG00559230.1" = "Amy1-1.3",
  "HORVU.BARKE.PROJ.6HG00559240.1" = "Amy1-1.4",
  "HORVU.BARKE.PROJ.6HG00559250.1" = "Amy1-1.5",
  "HORVU.BARKE.PROJ.6HG00559270.1" = "Amy1-1.6",
  "HORVU.BARKE.PROJ.6HG00561370.1" = "Amy1-2",
  "HORVU.BARKE.PROJ.7HG00653200.1" = "Amy2-1",
  "HORVU.BARKE.PROJ.7HG00653330.1" = "Amy2-2",
  "HORVU.BARKE.PROJ.7HG00653310.1" = "Amy2-3",
  "HORVU.BARKE.PROJ.5HG00441970.1" = "Amy3",
  "HORVU.BARKE.PROJ.2HG00148940.1" = "Amy4-1",
  "HORVU.BARKE.PROJ.3HG00259920.1" = "Amy4-2",

# =========================
# Beta-amylase
# =========================
"HORVU.BARKE.PROJ.4HG00380960.1" = "Bmy1",
"HORVU.BARKE.PROJ.2HG00126310.1" = "Bmy2",
"HORVU.BARKE.PROJ.1HG00036700.1" = "Bmy.1",
"HORVU.BARKE.PROJ.1HG00036740.1" = "Bmy.2",
"HORVU.BARKE.PROJ.3HG00228930.1" = "Bmy.3",
"HORVU.BARKE.PROJ.4HG00376360.1" = "Bmy.4",

"HORVU.BARKE.PROJ.4HG00303690.1" = "Bmy.5",

"HORVU.BARKE.PROJ.1HG00051030.1" = "Bmy.6",

"HORVU.BARKE.PROJ.2HG00126320.1" = "Bmy.7",

"HORVU.BARKE.PROJ.6HG00502080.1" = "Bmy.8",

"HORVU.BARKE.PROJ.2HG00105510.1" = "Bmy.9",

# =========================
# Alpha-glucosidase
# =========================
"HORVU.BARKE.PROJ.7HG00666400.1" = "Agl1",
"HORVU.BARKE.PROJ.3HG00202550.1" = "Agl2",

"HORVU.BARKE.PROJ.1HG00052860.1" = "Agl3",

"HORVU.BARKE.PROJ.4HG00360800.1" = "Agl4",
"HORVU.BARKE.PROJ.2HG00192840.1" = "Agl.1",
"HORVU.BARKE.PROJ.2HG00192870.1" = "Agl.2",
"HORVU.BARKE.PROJ.2HG00192880.1" = "Agl.3",
"HORVU.BARKE.PROJ.3HG00234540.1" = "Agl.4",
"HORVU.BARKE.PROJ.6HG00561650.1" = "Agl.5",
"HORVU.BARKE.PROJ.7HG00595360.1" = "Agl.6",


        

# =========================
# Isoamylase / Branching enzymes
# =========================
"HORVU.BARKE.PROJ.7HG00618070.1" = "Iso1",

"HORVU.BARKE.PROJ.1HG00057560.1" = "Iso2",

"HORVU.BARKE.PROJ.5HG00443610.1" = "Iso3",

"HORVU.BARKE.PROJ.2HG00154890.1" = "Iso.1",

"HORVU.BARKE.PROJ.2HG00149340.1" = "Iso.2",
"HORVU.BARKE.PROJ.7HG00681480.1" = "Iso.3",
"HORVU.BARKE.PROJ.7HG00822990.1" = "Iso.4",
"HORVU.BARKE.PROJ.7HG00823000.1" = "Iso.5",

"HORVU.BARKE.PROJ.7HG00645630.1" = "Iso.6",

# =========================
# Limit dextrinase
# =========================
"HORVU.BARKE.PROJ.7HG00595270.1" = "Lda",

# =========================
# Alpha glucan phosphorylase
# =========================
"HORVU.BARKE.PROJ.5HG00465050.1" = "Phs1",

"HORVU.BARKE.PROJ.3HG00275600.1" = "Phs2.1",
"HORVU.BARKE.PROJ.2HG00708120.1" = "Phs2.2",
"HORVU.BARKE.PROJ.3HG00275620.1" = "Phs2.3",

# =========================
# Cellulases / GH9 / KOR
# =========================
"HORVU.BARKE.PROJ.1HG00014980.1" = "CelB2",
"HORVU.BARKE.PROJ.1HG00018950.1" = "CelB6",
"HORVU.BARKE.PROJ.7HG00638840.1" = "Cel13",
"HORVU.BARKE.PROJ.2HG00162350.1" = "Cel5",
"HORVU.BARKE.PROJ.6HG00554880.1" = "Cel16",
"HORVU.BARKE.PROJ.7HG00607220.1" = "Cel.1",

"HORVU.BARKE.PROJ.5HG00427900.1" = "Cel.2",
"HORVU.BARKE.PROJ.2HG00154880.1" = "Cel.3",


"HORVU.BARKE.PROJ.3HG00226090.1" = "Cel.4",
"HORVU.BARKE.PROJ.4HG00379800.1" = "Cel.5",
"HORVU.BARKE.PROJ.1HG00078380.1" = "Cel.6",

"HORVU.BARKE.PROJ.4HG00311880.1" = "Cel.7",
"HORVU.BARKE.PROJ.4HG00345900.1" = "Cel.8",

"HORVU.BARKE.PROJ.1HG00073020.1" = "Cel.9",

"HORVU.BARKE.PROJ.2HG00192140.1" = "Cel.10",
"HORVU.BARKE.PROJ.7HG00596870.1" = "Cel.11",
"HORVU.BARKE.PROJ.7HG00606250.1" = "Cel.12",
"HORVU.BARKE.PROJ.6HG00508740.1" = "Cel.13",
"HORVU.BARKE.PROJ.6HG00558010.1" = "Cel.14",
"HORVU.BARKE.PROJ.5HG00434990.1" = "Cel.15",

"HORVU.BARKE.PROJ.6HG00561940.1" = "Cel.16",

"HORVU.BARKE.PROJ.7HG00653770.1" = "Cel.17",

"HORVU.BARKE.PROJ.5HG00452800.1" = "Cel.18",


"HORVU.BARKE.PROJ.5HG00453760.1" = "Cel.19",



  # =========================
# Lichenase
# =========================
"HORVU.BARKE.PROJ.1HG00053050.1" = "Glb1",
"HORVU.BARKE.PROJ.7HG00679910.1" = "Glb2",

# =========================
# Beta-glucosidase
# =========================
"HORVU.BARKE.PROJ.3HG00266610.1" = "Bglc.1",
"HORVU.BARKE.PROJ.3HG00266590.1" = "Bglc.2",

"HORVU.BARKE.PROJ.3HG00279420.1" = "Bglc.3",

# =========================
# Arabinofuranosidase
# =========================
"HORVU.BARKE.PROJ.2HG00101670.1" = "Asd.1",
"HORVU.BARKE.PROJ.4HG00325990.1" = "Asd.2",
"HORVU.BARKE.PROJ.4HG00326450.1" = "Asd.3",
"HORVU.BARKE.PROJ.4HG00346830.1" = "Asd.4",
"HORVU.BARKE.PROJ.5HG00420480.1" = "Asd.5",

# =========================
# Mannanase
# =========================
"HORVU.BARKE.PROJ.7HG00612160.1" = "Man.1",
"HORVU.BARKE.PROJ.5HG00479880.1" = "Man.2",
"HORVU.BARKE.PROJ.3HG00263090.1" = "Man.3",
"HORVU.BARKE.PROJ.5HG00771660.1" = "Man.4",
"HORVU.BARKE.PROJ.3HG00253120.1" = "Man.5",
"HORVU.BARKE.PROJ.6HG00561220.1" = "Man.6",
"HORVU.BARKE.PROJ.6HG00561230.1" = "Man.7",

# =========================
# Alpha-galactosidase
# =========================
"HORVU.BARKE.PROJ.1HG00041240.1" = "Agal.1",
"HORVU.BARKE.PROJ.5HG00772750.1" = "Agal.2",
"HORVU.BARKE.PROJ.1HG00041210.1" = "Agal.3",
"HORVU.BARKE.PROJ.5HG00772740.1" = "Agal.4",
"HORVU.BARKE.PROJ.2HG00102930.1" = "Agal.5",
"HORVU.BARKE.PROJ.6HG00537080.1" = "Agal.6",
"HORVU.BARKE.PROJ.6HG00556460.1" = "Agal.7",

# =========================
# Polygalacturonase
# =========================
"HORVU.BARKE.PROJ.5HG00475020.1" = "Pg.1",
"HORVU.BARKE.PROJ.5HG00475050.1" = "Pg.2",
"HORVU.BARKE.PROJ.7HG00599930.1" = "Pg.3",
"HORVU.BARKE.PROJ.4HG00360780.1" = "Pg.4",
"HORVU.BARKE.PROJ.6HG00502310.1" = "Pg.5",

# =========================
# Pectin esterase
# =========================
"HORVU.BARKE.PROJ.7HG00598930.1" = "Pme.1",
"HORVU.BARKE.PROJ.6HG00563120.1" = "Pme.2",
"HORVU.BARKE.PROJ.4HG00366450.1" = "Pme.3",
"HORVU.BARKE.PROJ.2HG00188100.1" = "Pme.4",
"HORVU.BARKE.PROJ.7HG00639410.1" = "Pme.5",
"HORVU.BARKE.PROJ.1HG00058460.1" = "Pme.6",
"HORVU.BARKE.PROJ.2HG00100460.1" = "Pme.7",
"HORVU.BARKE.PROJ.3HG00232130.1" = "Pme.8",
"HORVU.BARKE.PROJ.2HG00157350.1" = "Pme.9",
"HORVU.BARKE.PROJ.4HG00349290.1" = "Pme.10",
"HORVU.BARKE.PROJ.7HG00593290.1" = "Pme.11",
"HORVU.BARKE.PROJ.1HG00045130.1" = "Pme.12",
"HORVU.BARKE.PROJ.3HG00230000.1" = "Pme.13",
"HORVU.BARKE.PROJ.1HG00085260.1" = "Pme.14",
"HORVU.BARKE.PROJ.1HG00054490.1" = "Pme.15",
"HORVU.BARKE.PROJ.5HG00473670.1" = "Pme.16",
"HORVU.BARKE.PROJ.5HG00437860.1" = "Pme.17",
"HORVU.BARKE.PROJ.3HG00268440.1" = "Pme.18",
"HORVU.BARKE.PROJ.2HG00103570.1" = "Pme.19",
"HORVU.BARKE.PROJ.5HG00391190.1" = "Pme.20",
"HORVU.BARKE.PROJ.6HG00564170.1" = "Pme.21",
"HORVU.BARKE.PROJ.2HG00154320.1" = "Pme.22",
"HORVU.BARKE.PROJ.3HG00280980.1" = "Pme.23",
"HORVU.BARKE.PROJ.1HG00069330.1" = "Pme.24",
"HORVU.BARKE.PROJ.2HG00102050.1" = "Pme.25",
"HORVU.BARKE.PROJ.4HG00328630.1" = "Pme.26",
"HORVU.BARKE.PROJ.4HG00328640.1" = "Pme.27",
"HORVU.BARKE.PROJ.2HG00107880.1" = "Pme.28",
"HORVU.BARKE.PROJ.2HG00107890.1" = "Pme.29",
"HORVU.BARKE.PROJ.2HG00105250.1" = "Pme.30",
"HORVU.BARKE.PROJ.2HG00105260.1" = "Pme.31",
"HORVU.BARKE.PROJ.3HG00300590.1" = "Pme.32",
"HORVU.BARKE.PROJ.2HG00171960.1" = "Pme.33",
"HORVU.BARKE.PROJ.2HG00115940.1" = "Pme.34",
"HORVU.BARKE.PROJ.5HG00403060.1" = "Pme.35",

# =========================
# Papain-like cysteine endopeptidase
# =========================
"HORVU.BARKE.PROJ.5HG00420360.1" = "Pap4",
"HORVU.BARKE.PROJ.1HG00004010.1" = "Pap5",
"HORVU.BARKE.PROJ.3HG00278240.1" = "Pap14",
"HORVU.BARKE.PROJ.4HG00311230.1" = "Pap20",
"HORVU.BARKE.PROJ.2HG00192770.1" = "Pap24",
# XCPs
"HORVU.BARKE.PROJ.6HG00551890.1" = "Pap.1",

# CEPs

"HORVU.BARKE.PROJ.5HG00459870.1" = "Pap.2",
"HORVU.BARKE.PROJ.1HG00077400.1" = "Pap.3",
"HORVU.BARKE.PROJ.2HG00187860.1" = "Pap.4",
"HORVU.BARKE.PROJ.5HG00386780.1" = "Pap.5",

# RDs
"HORVU.BARKE.PROJ.2HG00144580.1" = "Pap.6",
"HORVU.BARKE.PROJ.2HG00120520.1" = "Pap.7",
"HORVU.BARKE.PROJ.2HG00186610.1" = "Pap.8",
"HORVU.BARKE.PROJ.2HG00193350.1" = "Pap.9",

# XBCPs
"HORVU.BARKE.PROJ.1HG00069020.1" = "Pap.10",

# F23E12 / F28N24 cluster
"HORVU.BARKE.PROJ.6HG00495640.1" = "Pap.11",
"HORVU.BARKE.PROJ.6HG00495650.1" = "Pap.12",
"HORVU.BARKE.PROJ.6HG00495660.1" = "Pap.13",
"HORVU.BARKE.PROJ.6HG00495670.1" = "Pap.14",
"HORVU.BARKE.PROJ.6HG00495680.1" = "Pap.15",
"HORVU.BARKE.PROJ.6HG00495690.1" = "Pap.16",
"HORVU.BARKE.PROJ.6HG00495700.1" = "Pap.17",
"HORVU.BARKE.PROJ.6HG00495710.1" = "Pap.18",
"HORVU.BARKE.PROJ.7HG00679480.1" = "Pap.19",
"HORVU.BARKE.PROJ.1HG00004790.1" = "Pap.20",

# =========================
# Serine endopeptidase
# =========================
"HORVU.BARKE.PROJ.1HG00044350.1" = "Prs.1",
"HORVU.BARKE.PROJ.5HG00396740.1" = "Prs.2",
"HORVU.BARKE.PROJ.5HG00440510.1" = "Prs.3",
"HORVU.BARKE.PROJ.6HG00555990.1" = "Prs.4",

# =========================
# Aspartic endopeptidase
# =========================

# APA family
"HORVU.BARKE.PROJ.1HG00081950.1" = "Apa.1",
"HORVU.BARKE.PROJ.3HG00291880.1" = "Apa.2",
"HORVU.BARKE.PROJ.6HG00567790.1" = "Apa.3",
"HORVU.BARKE.PROJ.7HG00652070.1" = "Apa.4",
"HORVU.BARKE.PROJ.1HG00016940.1" = "Apa.5",
"HORVU.BARKE.PROJ.3HG00249950.1" = "Apa.6",

# F17M5 cluster
"HORVU.BARKE.PROJ.4HG00318350.1" = "Apa.7",
"HORVU.BARKE.PROJ.5HG00415360.1" = "Apa.8",

# =========================
# Aminopeptidase
# =========================
"HORVU.BARKE.PROJ.4HG00305500.1" = "Apm.1",
"HORVU.BARKE.PROJ.5HG00432120.1" = "Apm.2",
"HORVU.BARKE.PROJ.6HG00524520.1" = "Apm.3",

# =========================
# Tripeptidyl peptidase
# =========================
"HORVU.BARKE.PROJ.6HG00545030.1" = "Tpp",

# =========================
# Dipeptidyl peptidase
# =========================
"HORVU.BARKE.PROJ.5HG00406500.1" = "Mop.1",
"HORVU.BARKE.PROJ.5HG00483050.1" = "Mop.2",
# =========================
# Carboxypeptidase (SCPL family etc.)
# =========================

# Non-SCPL
"HORVU.BARKE.PROJ.3HG00285590.1" = "Cbp1",
"HORVU.BARKE.PROJ.7HG00597910.1" = "Cbp2",
"HORVU.BARKE.PROJ.5HG00469870.1" = "Cbp.1",
"HORVU.BARKE.PROJ.3HG00265950.1" = "Cbp.2",
"HORVU.BARKE.PROJ.7HG00587970.1" = "Cbp.3",
"HORVU.BARKE.PROJ.7HG00591350.1" = "Cbp.4",
"HORVU.BARKE.PROJ.1HG00063380.1" = "Cbp.5",

# F27D4 cluster
"HORVU.BARKE.PROJ.1HG00008170.1" = "Cbp.6",
"HORVU.BARKE.PROJ.7HG00673840.1" = "Cbp.7",
"HORVU.BARKE.PROJ.7HG00653610.1" = "Cbp.8",

# F10A2 cluster
"HORVU.BARKE.PROJ.3HG00262480.1" = "Cbp.9",
"HORVU.BARKE.PROJ.3HG00262490.1" = "Cbp.10",

# SCPL family numerically ordered
"HORVU.BARKE.PROJ.4HG00311110.1" = "Cbp.11",
"HORVU.BARKE.PROJ.7HG00593540.1" = "Cbp.12",

"HORVU.BARKE.PROJ.2HG00722350.1" = "Cbp.13",
"HORVU.BARKE.PROJ.4HG00315620.1" = "Cbp.14",

"HORVU.BARKE.PROJ.2HG00171920.1" = "Cbp.15",
"HORVU.BARKE.PROJ.3HG00215140.1" = "Cbp.16",
"HORVU.BARKE.PROJ.2HG00171840.1" = "Cbp.17",
"HORVU.BARKE.PROJ.5HG00445740.1" = "Cbp.18",

"HORVU.BARKE.PROJ.2HG00144730.1" = "Cbp.19",

"HORVU.BARKE.PROJ.6HG00565120.1" = "Cbp.20",

"HORVU.BARKE.PROJ.1HG00018290.1" = "Cbp.21",
"HORVU.BARKE.PROJ.3HG00222540.1" = "Cbp.22",
"HORVU.BARKE.PROJ.4HG00364910.1" = "Cbp.23",
"HORVU.BARKE.PROJ.5HG00442890.1" = "Cbp.24",
"HORVU.BARKE.PROJ.3HG00277430.1" = "Cbp.25",

"HORVU.BARKE.PROJ.5HG00442850.1" = "Cbp.26",
"HORVU.BARKE.PROJ.5HG00442860.1" = "Cbp.27",

"HORVU.BARKE.PROJ.3HG00230350.1" = "Cbp.28",

"HORVU.BARKE.PROJ.1HG00048630.1" = "Cbp.29",
"HORVU.BARKE.PROJ.3HG00230360.1" = "Cbp.30",
"HORVU.BARKE.PROJ.6HG00540890.1" = "Cbp.31",
"HORVU.BARKE.PROJ.7HG00656000.1" = "Cbp.32",
"HORVU.BARKE.PROJ.7HG00673290.1" = "Cbp.33",

"HORVU.BARKE.PROJ.2HG00166340.1" = "Cbp.34",

"HORVU.BARKE.PROJ.2HG00107460.1" = "Cbp.35",
"HORVU.BARKE.PROJ.4HG00304700.1" = "Cbp.36",
"HORVU.BARKE.PROJ.5HG00442780.1" = "Cbp.37",
"HORVU.BARKE.PROJ.5HG00442910.1" = "Cbp.38",
"HORVU.BARKE.PROJ.5HG00442940.1" = "Cbp.39",
"HORVU.BARKE.PROJ.5HG00442970.1" = "Cbp.40",
"HORVU.BARKE.PROJ.2HG00107490.1" = "Cbp.41",
"HORVU.BARKE.PROJ.4HG00324700.1" = "Cbp.42",

"HORVU.BARKE.PROJ.4HG00316830.1" = "Cbp.43",

"HORVU.BARKE.PROJ.2HG00148330.1" = "Cbp.44",
"HORVU.BARKE.PROJ.6HG00522350.1" = "Cbp.45",

"HORVU.BARKE.PROJ.6HG00494930.1" = "Cbp.46",

"HORVU.BARKE.PROJ.2HG00183040.1" = "Cbp.47",
"HORVU.BARKE.PROJ.6HG00557070.1" = "Cbp.48",

"HORVU.BARKE.PROJ.1HG00084240.1" = "Cbp.49",
"HORVU.BARKE.PROJ.2HG00095780.1" = "Cbp.50",
"HORVU.BARKE.PROJ.1HG00084350.1" = "Cbp.51",

"HORVU.BARKE.PROJ.1HG00041660.1" = "Cbp.52",
"HORVU.BARKE.PROJ.1HG00070610.1" = "Cbp.53",
"HORVU.BARKE.PROJ.4HG00760680.1" = "Cbp.54",
"HORVU.BARKE.PROJ.6HG00499600.1" = "Cbp.55",
"HORVU.BARKE.PROJ.7HG00625610.1" = "Cbp.56",
"HORVU.BARKE.PROJ.7HG00668820.1" = "Cbp.57",


# =========================
# Lipoxygenase
# =========================

# Lox1
"HORVU.BARKE.PROJ.4HG00307600.1" = "Lox1",
"HORVU.BARKE.PROJ.4HG00307640.1" = "Lox.1",
"HORVU.BARKE.PROJ.2HG00188780.1" = "Lox.2",
"HORVU.BARKE.PROJ.5HG00481340.1" = "Lox.3",
"HORVU.BARKE.PROJ.2HG00188750.1" = "Lox.4",
"HORVU.BARKE.PROJ.5HG00462410.1" = "Lox.5",
"HORVU.BARKE.PROJ.5HG00462420.1" = "Lox.6",
"HORVU.BARKE.PROJ.5HG00462430.1" = "Lox.7",
"HORVU.BARKE.PROJ.5HG00462440.1" = "Lox.8",

# Lox2
"HORVU.BARKE.PROJ.5HG00383870.1" = "Lox.9",
"HORVU.BARKE.PROJ.6HG00519710.1" = "Lox.10",
"HORVU.BARKE.PROJ.6HG00490920.1" = "Lox.11",
"HORVU.BARKE.PROJ.7HG00579310.1" = "Lox.12",
"HORVU.BARKE.PROJ.7HG00616920.1" = "Lox.13",
"HORVU.BARKE.PROJ.4HG00760420.1" = "Lox.14",

# Lox3
"HORVU.BARKE.PROJ.2HG00156010.1" = "Lox.15",
"HORVU.BARKE.PROJ.4HG00369210.1" = "Lox.16",

# Lox4
"HORVU.BARKE.PROJ.4HG00369200.1" = "Lox.17",

# Lox5
"HORVU.BARKE.PROJ.5HG00462390.1" = "Lox.18",
"HORVU.BARKE.PROJ.1HG00050580.1" = "Lox.19",

# =========================
# Allene oxide synthase
# =========================
"HORVU.BARKE.PROJ.4HG00359600.1" = "Aos.1",
"HORVU.BARKE.PROJ.4HG00359650.1" = "Aos.2",

# =========================
# Acid phosphatase
# =========================

# Pap (unspecified)
"HORVU.BARKE.PROJ.2HG00124710.1" = "Paph.1",
"HORVU.BARKE.PROJ.6HG00793860.1" = "Paph.2",

# Pap1
"HORVU.BARKE.PROJ.1HG00084950.1" = "Paph.3",
"HORVU.BARKE.PROJ.6HG00797590.1" = "Paph.4",
"HORVU.BARKE.PROJ.7HG00629580.1" = "Paph.5",

# Pap2–3
"HORVU.BARKE.PROJ.2HG00142220.1" = "Paph.6",
"HORVU.BARKE.PROJ.2HG00164150.1" = "Paph.7",
"HORVU.BARKE.PROJ.2HG00112820.1" = "Paph.8",
"HORVU.BARKE.PROJ.4HG00358100.1" = "Paph.9",

# Pap9–10
"HORVU.BARKE.PROJ.5HG00472100.1" = "Paph.10",
"HORVU.BARKE.PROJ.3HG00267210.1" = "Paph.11",
"HORVU.BARKE.PROJ.6HG00565200.1" = "Paph.12",
"HORVU.BARKE.PROJ.5HG00475530.1" = "Paph.13",
"HORVU.BARKE.PROJ.4HG00372650.1" = "Paph.14",
"HORVU.BARKE.PROJ.4HG00755000.1" = "Paph.15",

# Pap15
"HORVU.BARKE.PROJ.3HG00231330.1" = "Paph.16",
"HORVU.BARKE.PROJ.5HG00454460.1" = "Paph.17",

# Pap18–23
"HORVU.BARKE.PROJ.6HG00507050.1" = "Paph.18",
"HORVU.BARKE.PROJ.5HG00417950.1" = "Paph.19",
"HORVU.BARKE.PROJ.5HG00417910.1" = "Paph.20",
"HORVU.BARKE.PROJ.1HG00004770.1" = "Paph.21",
"HORVU.BARKE.PROJ.2HG00149170.1" = "Paph.22",
"HORVU.BARKE.PROJ.2HG00185580.1" = "Paph.23",
"HORVU.BARKE.PROJ.5HG00430080.1" = "Paph.24",
"HORVU.BARKE.PROJ.5HG00419620.1" = "Paph.25",

# Pap26–27
"HORVU.BARKE.PROJ.7HG00675360.1" = "Paph.26",
"HORVU.BARKE.PROJ.1HG00078980.1" = "Paph.27",
"HORVU.BARKE.PROJ.5HG00448690.1" = "Paph.28",
"HORVU.BARKE.PROJ.5HG00448710.1" = "Paph.29",
"HORVU.BARKE.PROJ.7HG00620380.1" = "Paph.30",
"HORVU.BARKE.PROJ.4HG00361290.1" = "Paph.31",
"HORVU.BARKE.PROJ.5HG00400860.1" = "Paph.32",
"HORVU.BARKE.PROJ.4HG00376810.1" = "Paph.33",
"HORVU.BARKE.PROJ.3HG00269070.1" = "Paph.34",

# Other
"HORVU.BARKE.PROJ.6HG00533910.1" = "Paph.35",
# =========================
  # Catalase
  # =========================
"HORVU.BARKE.PROJ.7HG00681400.1" = "Cat.1",
"HORVU.BARKE.PROJ.4HG00374170.1" = "Cat.2",
"HORVU.BARKE.PROJ.6HG00495590.1" = "Cat.3",

  # =========================
# Superoxide dismutase
# =========================

# CSDs
"HORVU.BARKE.PROJ.2HG00105600.1" = "Csd.1",
"HORVU.BARKE.PROJ.5HG00440200.1" = "Csd.2",
"HORVU.BARKE.PROJ.5HG00440830.1" = "Csd.3",
"HORVU.BARKE.PROJ.7HG00625840.1" = "Csd.4",
"HORVU.BARKE.PROJ.4HG00360670.1" = "Csd.5",

# FSDs
"HORVU.BARKE.PROJ.7HG00580950.1" = "Csd.6",
"HORVU.BARKE.PROJ.7HG00588580.1" = "Csd.7",

# MSD
"HORVU.BARKE.PROJ.2HG00190130.1" = "Csd.8",

  # =========================
# Polyphenol oxidase
# =========================
"HORVU.BARKE.PROJ.3HG00268620.1" = "Ppo.1",
"HORVU.BARKE.PROJ.2HG00177380.1" = "Ppo.2",
"HORVU.BARKE.PROJ.2HG00177440.1" = "Ppo.3"
)

# ================================================================
# 1) Family → Category mapping (FULL EXPLICIT)
# ================================================================
family_category <- c(
  # ---------------- Starch ----------------
  "Alpha-amylase"              = "Starch degradation",
  "Beta-amylase"               = "Starch degradation",
  "Alpha-glucosidase"          = "Starch degradation",
  "Isoamylase"                 = "Starch degradation",
  "Limit dextrinase"           = "Starch degradation",
  "Alpha glucan phosphorylase" = "Starch degradation",

  # ---------------- Cell wall ----------------
  "Cellulase"           = "Cell wall degradation",
  "Lichenase"           = "Cell wall degradation",
  "Beta-glucosidase"    = "Cell wall degradation",
  "Arabinofuranosidase" = "Cell wall degradation",
  "Mannanase"           = "Cell wall degradation",
  "Alpha-galactosidase" = "Cell wall degradation",
  "Polygalacturonase"   = "Cell wall degradation",
  "Pectin esterase"     = "Cell wall degradation",

  # ---------------- Protein ----------------
  "serine endopeptidase"                = "Protein degradation",
  "papain-like cysteine endopeptidase" = "Protein degradation",
  "aspartic endopeptidase"             = "Protein degradation",
  "Aminopeptidase"                     = "Protein degradation",
  "tripeptidyl peptidase"              = "Protein degradation",
  "dipeptidyl peptidase"               = "Protein degradation",
  "Carboxypeptidase"                   = "Protein degradation",

  # ---------------- Lipid ----------------
  "Lipoxygenase"          = "Lipid degradation",
  "allene oxide synthase" = "Lipid degradation",

  # ---------------- Other ----------------
  "acid phosphatase"     = "Other",
  "Catalase"             = "Other",
  "superoxide dismutase" = "Other",
  "polyphenol oxidase"   = "Other"
)

# ================================================================
# 2) Collect Hv6 proteins
# ================================================================
id_cols <- grep("Protein", names(Hv6_final_matrix), value = TRUE)
if (length(id_cols) == 0) {
  stop("No columns containing protein IDs found in Hv6_final_matrix")
}

malting_rows <- lapply(names(all_families_ID), function(fam) {

  ids <- all_families_ID[[fam]]

  sub <- Hv6_final_matrix %>%
    dplyr::filter(dplyr::if_any(dplyr::all_of(id_cols), ~ . %in% ids))

  if (nrow(sub) > 0) {
    sub$enzyme_family <- fam
    sub$Category <- family_category[[fam]]
    return(sub)
  } else {
    return(NULL)
  }
})

malting_all <- dplyr::bind_rows(malting_rows)

# ================================================================
# FORCE SINGLE BARKE ID PER ROW (handle orthogroups correctly)
# ================================================================

primary_protein_col <- id_cols[1]

# Extract raw values
raw_ids <- malting_all[[primary_protein_col]]

# Split comma-separated orthogroups and keep FIRST ID only
malting_all$primary_barke_id <- vapply(
  strsplit(raw_ids, ","),
  function(x) trimws(x[1]),
  character(1)
)

# Remove empty IDs if any
malting_all <- malting_all[!is.na(malting_all$primary_barke_id) &
                           malting_all$primary_barke_id != "", ]

if (nrow(malting_all) == 0) {
  stop("No matching proteins found for Hv6.")
}

# ================================================================
# 3) Build RAW iBAQ matrix (NO Z-SCORE)
# ================================================================
iBAQ_cols <- grep("^iBAQ_norm_", names(malting_all), value = TRUE)

malting_mat <- as.matrix(malting_all[, iBAQ_cols, drop = FALSE])
mode(malting_mat) <- "numeric"

# Replace non-finite values
malting_mat[!is.finite(malting_mat)] <- NA

# ================================================================
# 4) Rename columns for plotting (FULLY EXPLICIT)
# ================================================================
colnames(malting_mat) <- colnames(malting_mat) %>%
  gsub("^iBAQ_norm_", "", .) %>%
  gsub("hs", " h", .) %>%
  gsub("_", " ", .) %>%
  gsub("barley",  "Barley",  ., ignore.case = TRUE) %>%
  gsub("wheat",   "Wheat",   ., ignore.case = TRUE) %>%
  gsub("rye",     "Rye",     ., ignore.case = TRUE) %>%
  gsub("sorghum", "Sorghum", ., ignore.case = TRUE)

# ================================================================
# 5) Row names using curated barley gene symbols
# ================================================================

# Extract primary protein IDs (first Protein column)
protein_ids <- malting_all$primary_barke_id

# Map to curated barley names
mapped_names <- barke_gene_map[protein_ids]

# Fallback to original ID if missing
mapped_names[is.na(mapped_names)] <- protein_ids[is.na(mapped_names)]

# Ensure uniqueness for pheatmap
rownames(malting_mat) <- make.unique(mapped_names)


# ================================================================
# 6) EXPLICIT FAMILY ORDER (FULL VECTOR, NO SHORTCUTS)
# ================================================================
family_names <- c(
  # ---------------- Starch ----------------
  "Alpha-amylase",
  "Beta-amylase",
  "Alpha-glucosidase",
  "Isoamylase",
  "Limit dextrinase",
  "Alpha glucan phosphorylase",

  # ---------------- Cell wall ----------------
  "Cellulase",
  "Lichenase",
  "Beta-glucosidase",
  "Arabinofuranosidase",
  "Mannanase",
  "Alpha-galactosidase",
  "Polygalacturonase",
  "Pectin esterase",

  # ---------------- Protein ----------------
  "serine endopeptidase",
  "papain-like cysteine endopeptidase",
  "aspartic endopeptidase",
  "Aminopeptidase",
  "tripeptidyl peptidase",
  "dipeptidyl peptidase",
  "Carboxypeptidase",

  # ---------------- Lipid ----------------
  "Lipoxygenase",
  "allene oxide synthase",

  # ---------------- Other ----------------
  "acid phosphatase",
  "Catalase",
  "superoxide dismutase",
  "polyphenol oxidase"
)

category_names <- c(
  "Starch degradation",
  "Cell wall degradation",
  "Lipid degradation",
  "Protein degradation",
  "Other"
)
annotation_row <- data.frame(
  Family   = malting_all$enzyme_family,
  Category = malting_all$Category,
  row.names = rownames(malting_mat)
)



annotation_row$Family   <- factor(annotation_row$Family, levels = family_names)
annotation_row$Category <- factor(annotation_row$Category, levels = category_names)

# ================================================================
# 7) Row ordering
# ================================================================
row_order <- order(annotation_row$Category, annotation_row$Family)

malting_mat   <- malting_mat[row_order, , drop = FALSE]
annotation_row <- annotation_row[row_order, , drop = FALSE]

# ================================================================
# 8) Colors (FULLY EXPLICIT)
# ================================================================
palette <- colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(length(family_names))
family_colors <- setNames(palette, family_names)

category_colors <- c(
  "Starch degradation"    = "#E6C229",
  "Cell wall degradation" = "#7FB685",
  "Protein degradation"   = "#E07A5F",
  "Lipid degradation"     = "#8DA0CB",
  "Other"                 = "#BDBDBD"
)

ann_colors <- list(
  Family   = family_colors,
  Category = category_colors
)

heat_colors <- colorRampPalette(c("#FFFFFF", "#FEE8C8", "#D73000"))(150)

# ================================================================
# 9) Output directory
# ================================================================
out_dir <- "../../../R_export/Hv6/malting_enzymes"
if (!dir.exists(out_dir)) {
  dir.create(out_dir, recursive = TRUE)
}

# ================================================================
# 10) Export RAW heatmaps (NO Z-SCORE)
# ================================================================
for (ext in c("jpeg", "svg", "pdf")) {

  outfile <- file.path(out_dir, paste0("Hv6_allFamilies_RAW.", ext))
  cat("Saving:", outfile, "\n")

  if (ext == "jpeg") jpeg(outfile, width = 14, height = 12, units = "in", res = 350)
  if (ext == "svg")  svg(outfile, width = 14, height = 12)
  if (ext == "pdf")  pdf(outfile, width = 14, height = 12)

  pheatmap::pheatmap(
    malting_mat,
    scale = "none",
    color = heat_colors,
    cluster_rows = FALSE,
    cluster_cols = FALSE,
    annotation_row = annotation_row,
    annotation_colors = ann_colors,
    show_rownames = TRUE,
    show_colnames = TRUE,
    legend = TRUE,
    annotation_legend = TRUE,
    fontsize_row = 5,
    fontsize_col = 7,
    main = "Hv6: Expression of enzyme families (raw normalized iBAQ)"
  )

  dev.off()
}

cat("✔ Hv6 RAW heatmaps saved in:", out_dir, "\n")

## ============================================================
## Hv6 — Species-specific malting enzyme heatmaps (HOG-anchored)
## ============================================================

library(dplyr)
library(pheatmap)
library(RColorBrewer)

# ------------------------------------------------------------
# 1. Load protein-resolved matrix
# ------------------------------------------------------------
Hv6 <- read.csv(
  "../../../R_Data/Hv6/Hv6_final_matrix_no_mean.csv",
  stringsAsFactors = FALSE
)

Hv6 <- Hv6 %>% filter(!grepl("^UNASSIGNED_", HOG))

# ------------------------------------------------------------
# 2. Rename iBAQ channels → timepoints
# ------------------------------------------------------------
timepoints <- c("0hs","12hs","24hs","36hs",
                "48hs","60hs","72hs","84hs","96hs")

rename_ibaq <- function(df, species){
  old <- grep(paste0("^iBAQ_norm_ch[0-9]+_", species), names(df), value = TRUE)
  new <- paste0("iBAQ_norm_", timepoints, "_", species)
  names(df)[match(old, names(df))] <- new
  df
}

for (sp in c("barley","wheat","rye","sorghum")) {
  Hv6 <- rename_ibaq(Hv6, sp)
}

# ------------------------------------------------------------
# 3. Identify malting HOGs FROM BARLEY ONLY
# ------------------------------------------------------------
barley_df <- Hv6 %>% filter(Species == "barley")

malting_barley <- lapply(names(all_families_ID), function(fam){

  ids <- all_families_ID[[fam]]
  sub <- barley_df %>% filter(Protein_ID %in% ids)

  if (nrow(sub) > 0) {
    sub$enzyme_family <- fam
    sub$Category <- family_category[[fam]]
    sub
  }
})

malting_barley <- bind_rows(malting_barley)

if (nrow(malting_barley) == 0)
  stop("No barley malting proteins detected.")

# Core anchor
malting_HOGs <- unique(malting_barley$HOG)

cat("Malting HOGs identified:", length(malting_HOGs), "\n")

# Map HOG → family/category
hog_annotation <- malting_barley %>%
  select(HOG, enzyme_family, Category) %>%
  distinct()

# ------------------------------------------------------------
# 4. Function to build species heatmaps
# ------------------------------------------------------------
make_species_heatmap <- function(species_name){

  cat("Processing:", species_name, "\n")

  df <- Hv6 %>% filter(Species == species_name)

  # Keep ONLY malting HOGs
  malting <- df %>% filter(HOG %in% malting_HOGs)

  if (nrow(malting) == 0) {
    cat("⚠ No proteins for", species_name, "\n")
    return(NULL)
  }

  # Attach family annotation
  malting <- malting %>%
    left_join(hog_annotation, by = "HOG")

      # ----------------------------------------
  # Build intensity matrix (RENAMED columns)
  # ----------------------------------------

  iBAQ_cols <- grep(
    paste0("^iBAQ_norm_.*_", species_name, "$"),
    names(malting),
    value = TRUE
  )

  # Keep only true timepoint columns (exclude GO / annotation garbage)
  iBAQ_cols <- iBAQ_cols[!grepl("GO|entry|TAIR|pname|gname", iBAQ_cols)]

  # Sort by timepoint order
  tp_order <- c("0hs","12hs","24hs","36hs","48hs","60hs","72hs","84hs","96hs")
  iBAQ_cols <- iBAQ_cols[order(match(sub(".*norm_", "", sub(paste0("_", species_name), "", iBAQ_cols)), tp_order))]

  if (length(iBAQ_cols) != 9) {
    stop(paste0(
      "Expected 9 iBAQ columns for ", species_name,
      " but found ", length(iBAQ_cols), ":\n",
      paste(iBAQ_cols, collapse = "\n")
    ))
  }

  mat <- as.matrix(malting[, iBAQ_cols, drop = FALSE])
  mode(mat) <- "numeric"
  mat[!is.finite(mat)] <- NA




  # ----------------------------------------
  # Clean column names
  # ----------------------------------------
  colnames(mat) <- gsub("^iBAQ_norm_", "", colnames(mat))
  colnames(mat) <- gsub("_", " ", colnames(mat))
  colnames(mat) <- gsub("hs", " h", colnames(mat))

  # ----------------------------------------
  # Row names per species
  # ----------------------------------------
  if (species_name == "barley") {

    ids <- malting$Protein_ID
    gene_names <- barke_gene_map[ids]
    gene_names[is.na(gene_names)] <- ids[is.na(gene_names)]

    rownames(mat) <- make.unique(paste0(malting$HOG, " - ", gene_names))

  } else if (species_name == "wheat") {

    rownames(mat) <- make.unique(
      paste0(malting$HOG, " - ", malting$wheatup_entry)
    )

  } else if (species_name == "sorghum") {

    rownames(mat) <- make.unique(
      paste0(malting$HOG, " - ", malting$sorghumup_entry)
    )

  } else if (species_name == "rye") {

    rownames(mat) <- make.unique(malting$HOG)
  }

  # ----------------------------------------
  # Row annotation
  # ----------------------------------------
  annotation_row <- data.frame(
    Family   = malting$enzyme_family,
    Category = malting$Category,
    row.names = rownames(mat)
  )

  family_names <- names(family_category)
  category_names <- unique(family_category)

  annotation_row$Family   <- factor(annotation_row$Family, levels = family_names)
  annotation_row$Category <- factor(annotation_row$Category, levels = category_names)

  ord <- order(annotation_row$Category, annotation_row$Family)
  mat <- mat[ord, , drop = FALSE]
  annotation_row <- annotation_row[ord, , drop = FALSE]

  # ----------------------------------------
  # Colors
  # ----------------------------------------
  palette <- colorRampPalette(RColorBrewer::brewer.pal(12, "Set3"))(length(family_names))
  family_colors <- setNames(palette, family_names)

  category_colors <- c(
    "Starch degradation"    = "#E6C229",
    "Cell wall degradation" = "#7FB685",
    "Protein degradation"   = "#E07A5F",
    "Lipid degradation"     = "#8DA0CB",
    "Other"                 = "#BDBDBD"
  )

  ann_colors <- list(Family = family_colors, Category = category_colors)
  heat_colors <- colorRampPalette(c("#FFFFFF","#FEE8C8","#D73000"))(150)

  # ----------------------------------------
  # Save heatmaps
  # ----------------------------------------
  out_dir <- "../../../R_export/Hv6/malting_enzymes_species"
  if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

  for (ext in c("jpeg","svg","pdf")) {

    outfile <- file.path(out_dir,
                         paste0("Hv6_", species_name, "_RAW_noMean.", ext))

    if (ext == "jpeg") jpeg(outfile, 14, 12, units="in", res=350)
    if (ext == "svg")  svg(outfile, 14, 12)
    if (ext == "pdf")  pdf(outfile, 14, 12)

    pheatmap(
      mat,
      scale = "none",
      color = heat_colors,
      cluster_rows = FALSE,
      cluster_cols = FALSE,
      annotation_row = annotation_row,
      annotation_colors = ann_colors,
      fontsize_row = 5,
      fontsize_col = 7,
      main = paste("Hv6:", species_name, "malting enzymes (HOG anchored)")
    )

    dev.off()
  }

  cat("✔ Done:", species_name, "\n")
}

# ------------------------------------------------------------
# 5. Generate all species
# ------------------------------------------------------------
for (sp in c("barley","wheat","rye","sorghum")) {
  make_species_heatmap(sp)
}

cat("✔ All species heatmaps generated\n")


```


#Unbiased analysis 
We want to identify that part of the proteome that is either much more abundant or uniquely expressed in each species 


```{r}

# --- Load + rename channels to timepoints ---
Hv6_final_matrix <- read.csv("../../../R_Data/Hv6/Hv6_final_matrix.csv",
                             header = TRUE, stringsAsFactors = FALSE)

timepoints <- c("0hs","0hs.1","12hs","24hs","36hs",
                "48hs","60hs","72hs","84hs","96hs","96hs.1")

rename_ibaq <- function(df, species){
  old <- grep(paste0("^iBAQ_norm_ch[0-9]+_", species), names(df), value = TRUE)
  tp_use <- timepoints[seq_len(min(length(old), length(timepoints)))]
  new <- paste0("iBAQ_norm_", tp_use, "_", species)
  names(df)[match(old, names(df))] <- new
  df
}
for (sp in c("barley","wheat","rye","sorghum")) {
  Hv6_final_matrix <- rename_ibaq(Hv6_final_matrix, sp)
}

# --- Parameters ---
fc_threshold <- 10
fc_threshold_hv <- 10
min_tp       <- 11
min_tp_hv       <- 11
fc_threshold_w <- 10
min_tp_w       <- 9
fc_threshold_r <- 10
min_tp_r      <- 9

# --- Long format ---
ibaq_long <- Hv6_final_matrix %>%
  dplyr::select(HOG, starts_with("iBAQ_norm_")) %>%
  pivot_longer(-HOG, names_to = "sample", values_to = "iBAQ") %>%
  separate(sample, into = c("prefix","norm","timepoint","species"),
           sep = "_", extra = "merge") %>%
  mutate(
    iBAQ = as.numeric(iBAQ),
    iBAQ = ifelse(is.na(iBAQ) | iBAQ <= 0, 1, iBAQ),
    timepoint = factor(timepoint, levels = timepoints)
  )

# --- Average per HOG × species × timepoint ---
species_avg <- ibaq_long %>%
  group_by(HOG, species, timepoint) %>%
  summarise(iBAQ = mean(iBAQ, na.rm = TRUE), .groups = "drop")

# --- Compare species at each timepoint ---
comparison_df <- species_avg %>%
  pivot_wider(names_from = species, values_from = iBAQ)

for (sp in c("barley","wheat","rye","sorghum")) {
  if (!sp %in% names(comparison_df)) comparison_df[[sp]] <- 1
}

comparison_df <- comparison_df %>%
  rowwise() %>%
  mutate(
    barley_bias  = (barley  >= fc_threshold_hv * max(c(wheat, rye, sorghum))),
    wheat_bias   = (wheat   >= fc_threshold_w * max(c(barley, rye, sorghum))),
    rye_bias     = (rye     >= fc_threshold_r * max(c(barley, wheat, sorghum))),
    sorghum_bias = (sorghum >= fc_threshold * max(c(barley, wheat, rye)))
  ) %>%
  ungroup()

# --- Count biased timepoints per HOG ---
species_bias_counts <- comparison_df %>%
  group_by(HOG) %>%
  summarise(
    barley_count  = sum(barley_bias,  na.rm = TRUE),
    wheat_count   = sum(wheat_bias,   na.rm = TRUE),
    rye_count     = sum(rye_bias,     na.rm = TRUE),
    sorghum_count = sum(sorghum_bias, na.rm = TRUE),
    .groups = "drop"
  )

# --- Lists of biased OGs ---
species_biased_lists <- list(
  barley  = species_bias_counts %>% filter(barley_count  >= min_tp_hv) %>% pull(HOG),
  wheat   = species_bias_counts %>% filter(wheat_count   >= min_tp_w) %>% pull(HOG),
  rye     = species_bias_counts %>% filter(rye_count     >= min_tp_r) %>% pull(HOG),
  sorghum = species_bias_counts %>% filter(sorghum_count >= min_tp) %>% pull(HOG)
)
biased_union <- unique(unlist(species_biased_lists))

# --- Define the universe (all TAIR IDs across 4 columns) ---
universe <- Hv6_final_matrix %>%
  dplyr::select(TAIR_hv, TAIR_wheat, TAIR_rye, TAIR_sorghum) %>%
  pivot_longer(everything(), values_to = "TAIR") %>%
  filter(!is.na(TAIR)) %>%
  mutate(TAIR = sub(";.*$", "", TAIR)) %>%
  pull(TAIR) %>%
  unique()

# --- Helper to extract TAIR IDs ---
extract_tair <- function(og_list, df) {
  df %>%
    filter(HOG %in% og_list) %>%
    dplyr::select(TAIR_hv, TAIR_wheat, TAIR_rye, TAIR_sorghum) %>%
    unlist() %>%
    na.omit() %>%
    sub(";.*$", "", .) %>%
    unique()
}
barley_tair  <- extract_tair(species_biased_lists$barley,  Hv6_final_matrix)
wheat_tair   <- extract_tair(species_biased_lists$wheat,   Hv6_final_matrix)
rye_tair     <- extract_tair(species_biased_lists$rye,     Hv6_final_matrix)
sorghum_tair <- extract_tair(species_biased_lists$sorghum, Hv6_final_matrix)

# --- Run enrichment analysis ---
enrich_barley  <- enrichGO(gene = barley_tair,  OrgDb = org.At.tair.db, keyType = "TAIR", ont = "ALL", universe = universe)
enrich_barley_2 <- enrich_barley@result %>%
  separate_rows(geneID, sep = "/")
enrich_barley_anno <- merge(
  enrich_barley_2,
  Hv6_final_matrix,
  by.x = "geneID",
  by.y = "TAIR_hv",
  all.x = TRUE
)
enrich_wheat   <- enrichGO(gene = wheat_tair,   OrgDb = org.At.tair.db, keyType = "TAIR", ont = "ALL", universe = universe)
enrich_wheat_2 <- enrich_wheat@result %>%
  separate_rows(geneID, sep = "/")
enrich_wheat_anno <- merge(
  enrich_wheat_2,
  Hv6_final_matrix,
  by.x = "geneID",
  by.y = "TAIR_hv",
  all.x = TRUE
)
enrich_rye     <- enrichGO(gene = rye_tair,     OrgDb = org.At.tair.db, keyType = "TAIR", ont = "ALL", universe = universe)
enrich_rye_2 <- enrich_rye@result %>%
  separate_rows(geneID, sep = "/")
enrich_rye_anno <- merge(
  enrich_rye_2,
  Hv6_final_matrix,
  by.x = "geneID",
  by.y = "TAIR_hv",
  all.x = TRUE
)
enrich_sorghum <- enrichGO(gene = sorghum_tair, OrgDb = org.At.tair.db, keyType = "TAIR", ont = "ALL", universe = universe)
enrich_sorghum_2 <- enrich_sorghum@result %>%
  separate_rows(geneID, sep = "/")
enrich_sorghum_anno <- merge(
  enrich_sorghum_2,
  Hv6_final_matrix,
  by.x = "geneID",
  by.y = "TAIR_hv",
  all.x = TRUE
)

#Generate a unique image with enrichments 
enrich_barley_anno_sel <- enrich_barley@result[c(2,3,4,5), ]
enrich_wheat_anno_sel <- enrich_wheat@result[c(2,3,4), ]
enrich_rye_anno_sel <- enrich_rye@result[c(1,3,4), ]
enrich_sorghum_anno_sel <- enrich_sorghum@result[c(10,17,24,26), ]

# Create a named list of enrichment results
enrich_list <- list(
  "Barley" = enrich_barley_anno_sel,
  "Wheat" = enrich_wheat_anno_sel,
  "Rye" = enrich_rye_anno_sel,
  "Sorghum" = enrich_sorghum_anno_sel
)

# Merge enrichment results into a compareClusterResult object
merged <- merge_result(enrich_list)

# Visualize results
p <- dotplot(merged, showCategory = 18)

# Print the plot (optional in some IDEs)
print(p)

# Save the dotplot
ggsave("../../../R_export/Hv6/unbiased_species_biased/dotplot_biased.png",
       plot = p, width = 9, height = 10, dpi = 300, scale=0.8)

# Save the plot
dev.copy(png, file = "../../../R_export/Hv4/unbiased/enrich_GO/dotplot_ance.png", width = 1500, height = 1500, res = 200)
dev.off()
dotplot(enrich_barley, showCategory = 15)
dotplot(enrich_wheat, showCategory = 15)
dotplot(enrich_rye, showCategory = 15)
dotplot(enrich_sorghum, showCategory = 15)


# --- Heatmap of all biased HOGs ---
species <- c("barley","wheat","rye","sorghum")
col_order <- unlist(lapply(species, function(sp) paste0("iBAQ_norm_", timepoints, "_", sp)))
col_order <- col_order[col_order %in% names(Hv6_final_matrix)]

hm_df <- Hv6_final_matrix[Hv6_final_matrix$HOG %in% biased_union,
                          c("HOG", col_order), drop = FALSE]
hm_mat <- as.matrix(hm_df[, col_order, drop = FALSE])
rownames(hm_mat) <- hm_df$HOG
hm_mat[is.na(hm_mat)] <- 1

outdir <- "../../../R_export/Hv6/unbiased_species_biased"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

jpeg(file.path(outdir, "species_specific_proteomes.jpeg"),
     width = 14, height = 10, units = "in", res = 350)

pheatmap(
  hm_mat,
  cluster_rows = TRUE,
  cluster_cols = FALSE,
  scale = "row",
  show_rownames = FALSE,
  show_colnames = TRUE,
  annotation_row = NULL,
  annotation_col = NULL,
  annotation_legend = FALSE,
  legend = TRUE,
  fontsize_col = 8,
  color = colorRampPalette(c("blue2","white","firebrick3"))(50),
  main = "Species-Biased Proteomes (z-score iBAQ)"
)

dev.off()



```

#Check beta-amy processing

```{r}
############################################################
## HOG peptide extraction & heatmap generation (PDF version)
############################################################

suppressPackageStartupMessages({
    library(data.table)
    library(pheatmap)
})

## ============================================================
## 0) USER INPUTS
## ============================================================

ortho_file <- "../../../R_Data/Hv6/Hv6_final_matrix.csv"
pept_dir   <- "../../../R_import/Hv6/peptides"

out_base   <- "../../../R_export/Hv6/HOG0005408_peptides"
dir.create(out_base, recursive = TRUE, showWarnings = FALSE)

target_hog <- "N0.HOG0005408"

# Thresholds
min_val  <- 0.1
smooth_k <- 3
thr_list <- list(
    barley  = 4,
    wheat   = 4,
    rye     = 4,
    sorghum = 4
)

# Species definitions
species_info <- data.table(
    Species  = c("barley",          "wheat",           "rye",              "sorghum"),
    File     = c("peptides_Hv.txt", "peptides_Ta.txt", "peptides_Sc.txt",  "peptides_Sb.txt"),
    BatchTag = c("Hv6_b1",          "Hv6_b2",          "Hv6_b3_rye",       "Hv6_b4_2"),
    ProtCol  = c("Proteins_barley", "Proteins_wheat",  "Proteins_rye",     "Proteins_sorghum")
)

# Create species folders
for (sp in species_info$Species) {
    dir.create(file.path(out_base, sp), recursive = TRUE, showWarnings = FALSE)
}

## ============================================================
## Helper functions
## ============================================================

edge_ma <- function(x, k = 3){
    if (all(is.na(x))) return(x)
    h <- floor(k / 2)
    x_ext <- c(rep(x[1], h), x, rep(x[length(x)], h))
    y <- stats::filter(x_ext, rep(1/k, k), sides = 2)
    as.numeric(y[(h+1):(length(y)-h)])
}

smooth_matrix <- function(mat, k = 3) {
    t(apply(mat, 1, edge_ma, k = k))
}

get_intensity_cols <- function(dt, batch_tag) {
    patt <- paste0("^Reporter intensity corrected [0-9]+ ", batch_tag, "$")
    grep(patt, names(dt), value = TRUE)
}

protein_in_proteins_column <- function(prot_id, proteins_vec) {
    sapply(
        proteins_vec,
        function(x) {
            ids <- trimws(unlist(strsplit(x, ",")))
            prot_id %in% ids
        }
    )
}

sanitize_prot_id <- function(prot_id) {
    gsub("[^A-Za-z0-9._-]", "_", prot_id)
}

## ============================================================
## 1) Preprocess peptides (impute, correct outliers)
## ============================================================

pep_list      <- list()
mask_list     <- list()
int_cols_list <- list()

for (i in seq_len(nrow(species_info))) {
    
    sp        <- species_info$Species[i]
    pep_file  <- file.path(pept_dir, species_info$File[i])
    batch_tag <- species_info$BatchTag[i]
    
    message("\n-----------------------------------------------------")
    message("📌 Species: ", sp)
    
    if (!file.exists(pep_file)) {
        message("⚠ Missing file, skipping...")
        next
    }
    
    dt <- fread(pep_file, sep = "\t", quote = "", header = TRUE)
    dt_sp <- copy(dt)
    
    int_cols <- get_intensity_cols(dt_sp, batch_tag)
    int_cols <- int_cols[1:9]
    dt_sp[, (int_cols) := lapply(.SD, as.numeric), .SDcols = int_cols]
    
    mat <- as.matrix(dt_sp[, ..int_cols])
    mode(mat) <- "numeric"
    
    thr <- thr_list[[sp]]
    mask <- matrix(FALSE, nrow(mat), ncol(mat))
    
    for (r in seq_len(nrow(mat))) {
        row_vals <- mat[r, ]
        for (j in seq_along(row_vals)) {
            val <- row_vals[j]
            neigh_idx <- intersect(c(j-1, j+1), seq_along(row_vals))
            neigh_vals <- row_vals[neigh_idx]
            neigh_vals <- neigh_vals[is.finite(neigh_vals) & neigh_vals > 0]
            if (!length(neigh_vals)) next
            neigh_mean <- mean(neigh_vals)
            if (is.na(val) || val < min_val) {
                row_vals[j] <- neigh_mean
                mask[r,j] <- TRUE
            } else {
                log2fc <- log2((val+1)/(neigh_mean+1))
                if (abs(log2fc) > thr) {
                    row_vals[j] <- neigh_mean
                    mask[r,j] <- TRUE
                }
            }
        }
        mat[r,] <- row_vals
    }
    
    for (j in seq_along(int_cols)) dt_sp[[int_cols[j]]] <- mat[, j]
    pep_list[[sp]]      <- dt_sp
    mask_list[[sp]]     <- mask
    int_cols_list[[sp]] <- int_cols
}

species_done <- names(pep_list)

## ============================================================
## 2) Intra-species TSN
## ============================================================

for (sp in species_done) {
    dt   <- pep_list[[sp]]
    mask <- mask_list[[sp]]
    int_cols <- int_cols_list[[sp]]
    
    mat <- as.matrix(dt[, ..int_cols])
    col_sums <- sapply(seq_along(int_cols), function(j) {
        sum(mat[!mask[,j], j], na.rm = TRUE)
    })
    
    target <- median(col_sums[col_sums > 0])
    factors <- target / col_sums
    
    for (j in seq_along(int_cols)) {
        idx <- !mask[, j] & is.finite(mat[, j])
        mat[idx, j] <- mat[idx, j] * factors[j]
    }
    
    for (j in seq_along(int_cols)) dt[[int_cols[j]]] <- mat[, j]
    pep_list[[sp]] <- dt
}

## ============================================================
## 3) Inter-species median centering
## ============================================================

species_meds <- sapply(species_done, function(sp) {
    dt <- pep_list[[sp]]
    mask <- mask_list[[sp]]
    int_cols <- int_cols_list[[sp]]
    vals <- unlist(lapply(seq_along(int_cols), function(j) dt[[int_cols[j]]][!mask[,j]]))
    median(vals)
})

grand_median <- median(species_meds)

for (sp in species_done) {
    dt   <- pep_list[[sp]]
    mask <- mask_list[[sp]]
    int_cols <- int_cols_list[[sp]]
    factor <- grand_median / species_meds[sp]
    for (j in seq_along(int_cols)) {
        idx <- !mask[, j] & is.finite(dt[[int_cols[j]]])
        dt[[int_cols[j]]][idx] <- dt[[int_cols[j]]][idx] * factor
    }
    pep_list[[sp]] <- dt
}

## ============================================================
## 4) Smoothing + log2
## ============================================================

for (sp in species_done) {
    dt <- pep_list[[sp]]
    int_cols <- int_cols_list[[sp]]
    mat <- as.matrix(dt[, ..int_cols])
    mat <- log2(smooth_matrix(mat, k = smooth_k) + 1)
    for (j in seq_along(int_cols)) dt[[int_cols[j]]] <- mat[, j]
    pep_list[[sp]] <- dt
}

## ============================================================
## 5) Load HOG table
## ============================================================

ortho <- fread(ortho_file)
row_target <- ortho[HOG == target_hog]

## ============================================================
## 6) Extract peptides per protein, write tsv + PDF heatmap
## ============================================================

for (i in seq_len(nrow(species_info))) {
    
    sp       <- species_info$Species[i]
    prot_col <- species_info$ProtCol[i]
    
    if (!(sp %in% species_done)) next
    
    prot_string <- as.character(row_target[[prot_col]])
    if (is.na(prot_string) || prot_string == "") next
    prot_vec <- trimws(unlist(strsplit(prot_string, ",")))
    prot_vec <- prot_vec[nzchar(prot_vec)]
    
    dt_sp    <- pep_list[[sp]]
    int_cols <- int_cols_list[[sp]]
    
    for (prot_id in prot_vec) {
        
        idx <- protein_in_proteins_column(prot_id, dt_sp$Proteins)
        dt_prot <- dt_sp[idx]
        if (nrow(dt_prot) == 0) next
        
        mat <- as.matrix(dt_prot[, ..int_cols])
        rownames(mat) <- make.unique(dt_prot$Sequence)
        colnames(mat) <- paste0("Ch", seq_len(ncol(mat)))
        
        out_dir <- file.path(out_base, sp, sanitize_prot_id(prot_id))
        dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
        
        fwrite(
            cbind(dt_prot[, .(Sequence, Proteins)], as.data.table(mat)),
            file.path(out_dir, "peptides_normalized.tsv"),
            sep = "\t"
        )
        
        # ---- PDF HEATMAP ----
        row_sd <- apply(mat, 1, sd, na.rm = TRUE)
        mat_hm <- mat[row_sd > 0, , drop = FALSE]
        
        if (nrow(mat_hm) >= 2) {
            pdf(file.path(out_dir, "heatmap.pdf"), width = 12, height = 9)
            pheatmap(
                mat_hm,
                scale="row",
                cluster_rows=TRUE,
                cluster_cols=FALSE,
                main=paste("Peptide map –", sp, prot_id, "\nHOG:", target_hog),
                show_rownames=TRUE,
                fontsize=8
            )
            dev.off()
        }
    }
}

message("\n🎯 DONE (PDF version)!")

```






